### 2024-09-20 manifesto-main.tex
Initial Overleaf Import

% Emerald Publishing - Construction Innovation Submission Template
% by Oleksandr Melnyk
% Ver 0.0.4
% Based on: https://www.emeraldgrouppublishing.com/journal/ci#author-guidelines

\documentclass[11pt]{article}

\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
%\usepackage[a4paper,top=1.5cm,bottom=1.5cm,left=1.5cm,right=1.5cm,marginparwidth=1.75cm]{geometry}
\usepackage[a4paper,top=2cm,bottom=2cm,left=2cm,right=2cm,marginparwidth=2.25cm]{geometry}

% Useful packages
\usepackage{amssymb}
\usepackage{siunitx}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\usepackage{cleveref}
\usepackage[utf8]{inputenc}
\usepackage[right]{lineno}
\usepackage{csquotes}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{adjustbox}
\usepackage{array}
\usepackage{url}
\usepackage{titlesec}
%\usepackage[compatibility=false]{caption}
\usepackage{authblk}
\usepackage{xcolor} % Load the xcolor package for color options
\usepackage{microtype}

%\usepackage{makecell} % multiline table cells

\renewcommand{\thetable}{\Roman{table}}

% Define a new format for \subsection
\titleformat{\subsection}
  {\mdseries\itshape\large} % Medium series, italic shape, and large font size
  {\thesubsection}{1em}{} % Numbering, spacing, and the section title itself

% Emerald Harvard Citation Style

\usepackage[english]{babel}
\usepackage[style=authoryear,backend=biber,natbib=true,maxcitenames=2,uniquelist=false]{biblatex}
\usepackage{natbib}
\addbibresource{manifesto-main.bib} % your .bib file

% Customizing biblatex for Harvard style
% Customizing biblatex for Harvard style
\DeclareNameAlias{sortname}{family-given}
\DeclareNameAlias{default}{family-given}

\renewbibmacro{in:}{}
\DeclareFieldFormat[article]{title}{\mkbibquote{#1}\addcomma}
\DeclareFieldFormat[book]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[bookinbook]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[inbook]{title}{\mkbibquote{#1}\addcomma}
\DeclareFieldFormat[incollection]{title}{\mkbibquote{#1}\addcomma}
\DeclareFieldFormat[inproceedings]{title}{\mkbibquote{#1}\addcomma}
\DeclareFieldFormat[manual]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[misc]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[thesis]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[unpublished]{title}{\mkbibquote{#1}\addcomma}
\DeclareFieldFormat[patent]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[report]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[online]{title}{\mkbibquote{#1}\addcomma}
\DeclareFieldFormat[software]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[booklet]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[periodical]{title}{\mkbibemph{#1}\addcomma}
\DeclareFieldFormat[standard]{title}{\mkbibemph{#1}\addcomma}

\DeclareFieldFormat[article]{journaltitle}{\iffieldundef{shortjournal}{\mkbibemph{#1}\addcomma}{\mkbibemph{\printfield{shortjournal}}\addcomma}}
\DeclareFieldFormat{volume}{\bibstring{volume}~#1}
\DeclareFieldFormat{number}{\bibstring{number}~#1}

% Definitions for "Vol." and "No."
\DefineBibliographyStrings{english}{
  volume = {Vol.},
  number = {No.}
}

\renewbibmacro*{volume+number+eid}{%
  \printfield{volume}%
  \setunit*{\addspace}%
  \printfield{number}%
  \setunit{\addcomma\space}%
  \printfield{eid}}

\renewbibmacro*{journal+issuetitle}{%
  \usebibmacro{journal}%
  \setunit*{\addcomma\space}%
  \usebibmacro{volume+number+eid}%
  \setunit{\addcomma\space}%
  \usebibmacro{issue+date}}

\renewbibmacro*{publisher+location+date}{%
  \printlist{publisher}%
  \iflistundef{location}
    {\setunit*{\addcomma\space}}
    {\setunit*{\addcolon\space}}%
  \printlist{location}%
  \setunit*{\addcomma\space}%
  \usebibmacro{date}}

\renewcommand*{\bibpagespunct}{\addcomma\space}

% Customizing page field format to prevent duplication
% \DeclareFieldFormat{pages}{%
%   \mkfirstpage[{\mkpageprefix[page]{#1}}]{#1}}

% Customizing citations for Harvard style
\DeclareCiteCommand{\cite}[\mkbibparens]
  {\usebibmacro{prenote}}
  {\usebibmacro{citeindex}%
   \usebibmacro{cite}}
  {\multicitedelim}
  {\usebibmacro{postnote}}

\renewbibmacro*{cite:labelyear+extrayear}{%
  \iffieldundef{labelyear}
    {}
    {\printtext[bibhyperref]{%
       \printfield{labelyear}%
       \printfield{extrayear}}}}

\renewbibmacro*{cite:labeldate+extradate}{%
  \iffieldundef{labelyear}
    {}
    {\printtext[bibhyperref]{%
       \printfield{labelyear}%
       \printfield{extradate}}}}

\AtEveryBibitem{
  \clearfield{month}
  \clearfield{day}
  \ifentrytype{book}{
    \clearlist{location}
  }{}
}

% Formatting "et al." in italics followed by a comma
\DefineBibliographyStrings{english}{
  andothers = {\textit{et al.},}
}

\DeclareFieldFormat[article]{volume}{\bibstring{jourvol}\addnbspace #1}
\DeclareFieldFormat[article]{number}{\bibstring{number}\addnbspace #1}
\DeclareFieldFormat[article]{volume}{Vol. #1}
\DeclareFieldFormat[article]{number}{No. #1}
% Customizing DOI field format to lowercase "doi"
%\DeclareFieldFormat{doi}{\bibstring{doi}\addcolon\space\url{#1}}

% Customizing URL field format to "available at:"
\DeclareFieldFormat{url}{\bibstring{available at}\addcolon\space\url{#1}}
\DeclareFieldFormat{urldate}{\mkbibparens{accessed \addspace#1}}

% Customizing urldate to match the required format
\DeclareFieldFormat{urldate}{%
  \mkbibparens{accessed\space%
    \thefield{urlday}\addspace%
    \mkbibmonth{\thefield{urlmonth}}\addspace%
    \thefield{urlyear}}}

% Configure cleveref
\crefformat{figure}{#2Figure~#1#3}
\Crefformat{figure}{#2Figure~#1#3}
\crefformat{table}{#2Table~#1#3}
\Crefformat{table}{#2Table~#1#3}
\crefformat{section}{#2Section~#1#3}
\Crefformat{section}{#2Section~#1#3}

\title{Building a continuous benchmarking ecosystem in bioinformatics}

\author[1,2,*]{Izaskun Mallona}
\author[1,2,3]{Almut Lütge}
\author[2,4]{Charlotte Soneson}
\author[1]{Ben Carrillo}
\author[1,2]{Reto Gerber}
\author[1]{Daniel Incicau}
\author[1,2]{Anthony Sonrel}
\author[1,2,*]{Mark D. Robinson}

\affil[1]{Department of Molecular Life Sciences, University of Zurich, 8057 Zurich, Switzerland}
\affil[2]{SIB Swiss Institute of Bioinformatics, University of Zurich, 8057 Zurich, Switzerland}
\affil[3]{Swiss Data Science Centre, 8092 Zurich, Switzerland}
\affil[4]{Friedrich Miescher Institute for Biomedical Research, 4056 Basel, Switzerland}

\affil[*]{\{mark.robinson,izaskun.mallona\}@mls.uzh.ch}

\linespread{1.4}

\begin{document}
\maketitle

\linenumbers

\begin{abstract}
Benchmarking, which involves collecting reference datasets and demonstrating method performances, is a requirement for the development of new computational tools, but also becomes a domain of its own to achieve neutral comparisons of methods. Although a lot has been written about how to design and conduct benchmark studies, this perspective sheds light on a wish list for a computational platform to orchestrate benchmark studies. We discuss various ideas for organizing reproducible software environments, formally defining benchmarks, orchestrating standardized workflows, and how they interface with computing infrastructure.
\end{abstract}

%add 6 keywords
\textbf{Keywords:} benchmarking; bioinformatics; reproducibility; software environments; licensing; workflows; infrastructure

\pagebreak

\section*{Introduction}
\label{sec:introduction}

In this work, we give our perspective on how to build a high-standard benchmarking ecosystem in the field of bioinformatics, coming primarily from the viewpoint of methodologists (i.e., researchers with computational background working on tools for data). In particular, our case is one of computational scientists embedded in data-rich biology (i.e., computational biology and bioinformatics), where rapid technology development is a distinguishing factor (e.g., single-cell assays coupled to high-throughput sequencing, imaging, etc.); this paucity in turn results in a rapid response from the computational community to develop tools, models, and software for the analysis of such emerging datasets. Although our viewpoint is `bioinformatics', there are analogs in adjacent disciplines where methodologists are creating computational tools in a certain scientific context. One critical step in method development is benchmarking (further defined below), whereby either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; for these, we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, defined in a recent meta-analysis study \cite{Cao2023-jz}. Notably, there are various ways to structure the running of a benchmarking study: it can be run by a single person or a small group on a laptop (or a server), or it can be organized as a `challenge' or within the duration of a hackathon (groups of people that come together to competitively assess their approaches and/or populate benchmark components).

Our goal is to lay out general benchmark parameters and definitions,  while abstracting towards an ideal \textit{benchmarking system}. This ranges from general benchmark conceptualization to various technical specifications. We will offer some remarks on the philosophy around the benchmarking activity (e.g., gatekeeping), as well as some technical recommendations (e.g., standards).

\section{What is a benchmark?}
%\label{sec:main_section}

To begin our discussion of benchmarking (in bioinformatics), a few definitions are needed. For example, what actually is a benchmark, precisely? Here, a benchmark is thought of as a conceptual entity that roughly encompasses all the components of a study to understand the performance of a set of (computational) methods for a given task (here, the primary context is computational methods, but the concept can have a broader scope, e.g., evaluation of data-generating technologies). A benchmark requires a well-defined task, such as inferring what set of genes are differentially expressed from a certain type of data (e.g., gene expression from RNA sequencing counts) or whether a high-dimensional dataset can be clustered into the `correct' cell types (the definition of this correctness, or ground-truth, should be precisely defined in advance). Given a well-defined task, a benchmark consists of benchmark `components' (such as simulations to generate datasets, preprocessing steps, methods, metrics). Ideally, a benchmarking system is used to organize and orchestrate the running of the benchmark, and create `artifacts' (code snapshots, file outputs to be shared, results tables). Ultimately, the downstream results and interpretations, which may involve rankings and various additional analyses, will be derived from these artifacts. The various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny. Such a system should facilitate components to be public, open and allow contributions (e.g. corrections or new components), such that the benchmarks of the future can run more continuously.

At the outset of a benchmark, a benchmark `definition' could be envisioned, which could give a formal specification of the entire set of components (and the pattern of artifacts to be generated). The benchmark definition could even be expressed as a single configuration file that specifies the scope and topology of components to include, details of the repositories (code implementations with versions), the instructions to create software environments (accessible across compute architectures), and the parameters used. A benchmark definition in terms of layers and their corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/bens_layers.pdf}
    \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges. The process of benchmarking is multilayered, with layer-specific factors belonging to different categories (colored boxes) and challenges (in red). \emph{Hardware}: infrastructure and its costs. \emph{Data}: dataset archival, openness, interoperability, archival, and selection.  \emph{Software}: method implementations, reproducibility, workflow execution, continuous integration and delivery (CI/CD), versioning, quality assurance (QA). \emph{Community}: Standardization and impartiality (governance, transparency, building trust, long term maintainability). \emph{Knowledge}: research and meta-research, academic publications.}
    \label{fig:layers}
\end{figure}

\section*{Benchmarks stakeholders}

Benchmarks are useful for a broad set of stakeholders. For a \emph{data analyst}, with an aim to perform a specific type of analysis on a specific dataset, benchmarks are helpful tools to find a suitable method for the task. However, it is important that the benchmarks consider datasets that are similar to the one of interest to the analyst, as method performance often depends on characteristics of the data. Thus, a thorough characterization of the datasets considered in a benchmark is desirable. Given that benchmarks typically evaluate methods using a range of metrics and datasets, not all of which are relevant to all analysts, flexibility in defining the ranking approach and aggregation of metrics is preferable. Such flexibility, which requires access to various artifacts, is typically not part of published benchmarks today \cite{Sonrel2023-te}. However, a well-structured benchmarking system could accommodate flexible filtering and aggregation of metrics and datasets, and additionally provide access to the code and software stack that the analyst would need to apply the selected method to their own dataset.

For a \emph{method developer}, benchmarking serves the purpose of comparing their method to the state of the art in the field. Ideally, this should be done in a way that is as neutral as possible, avoiding selecting datasets and metrics that would unfairly favor the new method \cite{Boulesteix2013-vy, Weber2019-el}. A diverse, easily accessible, collection of benchmarking datasets and metrics for a given task would be one way of addressing this, and additionally has the benefit of lowering the barrier of entry for method developers into a field of application. Moreover, it is important that the new method is compared to the current state of competing methods, to best mimic the choices available to a potential method user. One could, however, imagine other use cases, where the new method would be systematically compared to previous versions of itself or alternative implementations of the `same' method \cite{Rich2024-jh}, to illustrate stability of code changes or improvements made over time. In both cases, the typical approach today would be for the method developer to rerun all the methods and versions to which they want to compare their current method, on all the datasets they wish to use. This can be challenging and time consuming, since not all methods run on the same hardware setup or in the same software environment (e.g., R and Python versions, library dependencies). It also leads to a lot of redundancy across method comparisons (multiple stakeholders implementing similar workflows), which could be reduced by the presence of a benchmarking system where such results would be accessible and extendable. Finally, the results of the benchmark of a new method should be included in the corresponding publication, and thus an easy way of including the new method into the set of results served by the benchmarking system, and generating a snapshot in time for reproducibility, would be of great help to the developer.

Benchmarks can also be of high value to \emph{scientific journals and funding agencies}. Well-executed benchmark studies can be highly utilized, influential and cited, and may point to gaps in the current body of methods, guiding future methodological developments; at the same time, in fast-moving fields like bioinformatics, benchmarks have the tendency to rapidly become stale. Journals and funding bodies also have a vested interest in ensuring that published or funded method evaluations are performed to a high standard, that unnecessary redundancy is reduced, and that results are FAIR (findable, accessible, interoperable and reusable) for maximal benefit to the community \cite{Wilkinson2016-bh}. Thus, quality assurance, neutrality and transparency of benchmarking systems in themselves become highly relevant.

Finally, benchmarks could form a research topic on their own, either for newcomers to the field to get acquainted with the conceptual options that are available, or for experts to consolidate research efforts of a subfield. For this, it is important that there are suitable venues for publishing benchmarks. Regardless, a \emph{benchmarker} (i.e., researcher leading a benchmarking study) would also benefit from using a benchmarking system and curated collections of benchmark artifacts. Such benchmarkers may be good candidates for curating and maintaining such collections by designing and leading benchmark efforts as well as guiding other contributors to adhere to a high standard.

\section*{Benchmarks could be formally defined}

Benchmark artifacts can be seen as a collection of data files and (perhaps diverse) source code implementations assembled in a pipeline topology (e.g., methods are run on some input data, and method outputs act as input for metrics) that were run as a workflow in some computing environment. Current practices in single-cell RNA-seq benchmarking show that code is frequently open, reusable and versioned; input datasets are also typically available \cite{Sonrel2023-te}. However, the \textit{extensibility} of benchmarking studies is generally very low, as well as the proportion of benchmarks using a formal workflow system. Thus, a formal system to organize components and orchestrate the crossing of datasets, methods, and metrics would largely impact the current landscape.

In an ideal situation, running a benchmark includes executing a workflow, a process for which over 350 workflow languages, platforms or systems  \cite{Wratten2021-el, Amstutz2024-qk} and one standard named Common Workflow Language (CWL) (with multiple implementations) exist \cite{Amstutz2016-vo}. Some workflow languages can export workflow definitions to CWL, hence helping computational FAIR guidelines: that is, bundling and processing data in association to their metadata, tracking new metadata generation, recording provenance, and being accessible \cite{Goble2020-ps}. Both the workflow layout and provenance can be formally defined (Figure~\ref{fig:formalization}).

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{figures/bens_formalization.pdf}
    \caption[Benchmark formalization.]{Benchmark formalization. \textbf{A}) Running a benchmark requires running a workflow (execution phase) to generate results to be critically evaluated (analysis phase). \textbf{B}) the execution phase consists on a mapping of methods to specific input files to generate output files, with optional steps to harmonize data and parametrize runs.}
    \label{fig:formalization}
\end{figure}

Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and ideally efficient software stack; managing storage and their access-control lists; rendering results dashboards and summarizing benchmarking results; version controlling code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results. Benchmarking systems are in charge of orchestrating both workflow execution and these benchmarking-specific tasks.

Currently, most benchmarking systems aim to enhance the workflow formalization and execution with strategies to fulfill benchmarking-specific needs. To name a few of these systems, \texttt{ncbench} bundles workflow specification and software management with \texttt{snakemake} to benchmark result visualization with \texttt{Datavzrd} \cite{Hanssen2023-uf}. \texttt{OpenEBench} runs workflows in any workflow language, frequently \texttt{nextflow}, and makes use of openVRE as GUI and means of visualization \cite{Capella-Gutierrez2017-dh}. \texttt{OpenProblems} uses \texttt{Viash} and \texttt{nextflow} to handle software environments and workflows, respectively, and reports updated leaderboards with custom code \cite{Luecken2024-fk}. To our knowledge, no benchmarking platform makes use of a fully expressive benchmark definition formalization language covering both the workflow generation and the benchmarking-specific tasks. Such a language would go a long way toward facilitating benchmark conceptualization, community gathering and result sharing.

\section*{Scope and interpretation of benchmarking results}

Benchmarks are often focused on finding the `top' performing method, or on generating ranked lists of methods. While this is valid in a few fields with clearly defined goals, such as image recognition \cite{Khan2018-ot} or protein structure prediction \cite{Jumper2021-oq}, bioinformatics tasks are typically evaluated by multiple metrics, some of which are not well understood \cite{Lutge2021-mt, Reinke2024-dw}. Additionally, ground truths are not always well-defined (e.g., derived with uncertainty from other experimental data), multiple tasks (e.g., represented by different reference datasets) may be evaluated for a single method and several ways to assess performance are possible. Even when accompanied by dataset- and/or metric-specific performance measures, scaling and aggregation of results makes it difficult to unpack fine-grained nuances of performance from a final summary table reported in most benchmarks (e.g., \footnote{\url{https://www.nature.com/articles/s41592-021-01336-8/figures/11}}). Strobl \textit{et al.} push back against the `one method fits all datasets’ philosophy and advocate for highlighting dataset or parameter properties that are predictive of method performance \cite{Jelizarow2010-kr, Boulesteix2010-zj, Strobl2024-lp}. This aligns with what a data analyst wants: to identify the best performers in settings most similar to their problem at hand. This requires not only a thorough benchmark design, e.g., tracking of meaningful dataset metadata \cite{Strobl2024-lp}, but also can be supported by a benchmarking system that facilitates flexible navigation of results (e.g., via an interactive dashboard).

Besides dashboards for more nuanced result exploration \cite{MariniUnknown-aq}, multi-criteria decision analysis (MCDA) \cite{Taherdoost2023-wd} has been proposed to guide users through benchmark results and multidimensional scaling (MDS) has been used to differentiate the effect of single datasets or metrics and recognize patterns beyond aggregated performance \cite{Niessl2022-fk}. An extensive analysis and presentation of the results not only helps the users in the long term but also the whole scientific field to identify pain points, the complexity of possible scenarios and to isolate the different use-cases. The first advantage of such a strategy is to force the user to an attentive reading of the performance results so that the correct use-case and relevant methods can be identified. The second, more important, advantage to present benchmark results extensively is to identify the challenges in a field of analysis, i.e., the scenarios, types of datasets, or aspects of performance that are systematically associated with poor results for all current methods. Being public and open to contributions from specialists of different fields, a benchmarking system could gather more reference datasets and methods, and allow the current state of the field to be better grasped. %As a result, the performance results will tend to better fit to the current state of the field, as opposed to a static benchmark led by a small panel of experts that leads to a smaller, and sometimes (un)intentionally biased view of the field \cite{Boulesteix2010-zj, Weber2019-el}.

\section*{Parameter sweeps}

Parameters can be included into a benchmark to test method performance across a range of settings, allowing the impact of parameter choices to be assessed  or to find the optimal parameter settings. Notably, when assessing the impact of a certain parameter, the research objective of the benchmark differs from a classic method comparison: methods that do not include the parameter to be explored cannot be evaluated for this question. Parameters can be method-specific, but also dataset-specific, e.g., the true number of clusters when evaluating clustering algorithms. In this case, they can also be seen as (dataset-specific) metadata and the method performances can be compared depending on the availability of this metadata. A direct comparison between methods, with only some utilizing these metadata prevents differentiating between the effect of the parameter and the method itself for performance evaluation. Similarly, optimizing parameters for some but not other methods can skew results. Again, method-specific parameter tuning can be seen as a separate benchmarking task. Alternatively, a method that uses different settings can be interpreted as multiple distinct methods that are evaluated as part of the benchmark. This should be clearly communicated in the results.

Altogether, evaluating parameters as part of a benchmark can give useful insights, but should be linked to a distinct and independent research objective.
For continuous and collaborative benchmarking, parameter tuning and usage should be governed and coordinated between all contributors. Ideally, the benchmarker specifies the parameter scope and related research objectives at the beginning of the benchmark to ensure comparability and interpretability of the results.

\section*{Collaboration, gatekeeping and benchmarking infrastructure go hand in hand}

In designing a benchmarking system, it may be helpful to specify whether the benchmark is intended to be run `centralized' or `decentralized'. These terms can take different meanings depending on the context; we will focus on the organization of the benchmark infrastructure. In most systems, the code to execute will be tracked in a reference git repository (e.g., on GitHub, GitLab). The workflow execution and the storage of derived datasets (as various stages of results) are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery platforms (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., could be a laptop, server or cloud).

Traditionally, benchmarks (whether BOP or MDP) are conducted in a decentralized fashion and typically by a small number of researchers (e.g., for MDP, the new method developer). It is worth mentioning some disadvantages of the decentralized approach. For example, decentralization can be hugely inefficient; researchers studying the same task would collect reference datasets and code snippets to run methods, and implement a bespoke strategy (e.g., shell scripts) to orchestrate the methods run on datasets. Working together would mean sharing not only datasets and code snippets, but also execution environments (software but perhaps also hardware), alignment on parameters to evaluate and eventually sharing, consolidating, and interpreting results. However, because there are no standards of how benchmarks are implemented \cite{Sonrel2023-te}, another large disadvantage of decentralized benchmarks is that, while these components may eventually be openly shared as a product of a research study, the contents of a benchmark are non-interoperable, even across benchmarks of the same task. Furthermore, decentralized benchmarks have been criticized on the basis of `inflating model performance' due to decisions made in preprocessing or parameter tuning \cite{Luecken2024-fk}; such issues could be mitigated by gathering the `wisdom of crowds' that is implicit in running centralized benchmarks.

Arguably, most benchmarks could naturally start out as decentralized, in part as a means to properly define the concept and problem that one wants to work on, but also as a way to generate preliminary results and ideas to solidify the design before `going public' (centralization). Thus, users need a system that can build benchmark components in both a centralized and decentralized fashion, i.e., starting decentralized and promoting a benchmark to centralized when it becomes sufficiently mature. Accordingly, the process of centralization is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to resources) and adding measures to build trust and transparency.

\section*{Building trust and bringing together communities}

An active community is vital to the success of a benchmarking system. As mentioned, different roles and ways that people can contribute to or interact with a benchmarking system have been defined. The benchmarkers (which may include multiple scientists) are responsible for planning and coordinating a benchmark, defining the task, possibly splitting it into subtasks or processing stages, and defining the data formats across the stages; the benchmarker also brings an authority role of how to review and approve contributions. The contributors (which may include method developers) curate and add content to the benchmark, which could be new datasets, analysis methods or evaluation metrics, adhering to the guidelines set up by the benchmarker. Finally, the viewers (which may include benchmarkers and contributors) of benchmark results are users who retrieve one or more of the artifacts (dataset files, intermediate results, or metric scores). This could span a range of use cases, including the data analyst choosing which method to use for a specific application, an instructor retrieving a curated dataset for teaching purposes, or a methods researcher prototyping their new method.

Given the importance of such a community around a benchmarking system, the question arises on how it can best be assembled and sustained. Some strategies include organizing events, including seminars or workshops (typically on the educational side) as well as hackathons and challenges (more aimed at practical onboarding or solicitation of contributions). Career-related or financial incentives, such as the prospect of a scientific publication, or compensation for the allocated time, are also common motivators. For a method developer, contributing to a public benchmark provides a way to test their method without the need to set up all the infrastructure themselves, and also provides an opportunity to advertise their work. Ultimately, the likelihood that someone will contribute to a benchmark will likely be strongly driven by whether the results are of interest to them and their research, as well as how easy it is to contribute. Community engagement and contributions can further be made easier by providing clear guidelines (and organization), a suggestion of `good first issues' to tackle, a clear set of contributor guidelines and a code of conduct that covers all interactions related to the project.

The public solicitation of contributions, however, is not without difficulties, and it is important to organize the setup transparently to ensure trust from all parties. A benchmarker should be able to trace and test contributions, to detect misbehaviour and suspicious activities, but also to avoid that non-functional or suboptimal contributions enter the benchmark. This can be addressed by a `quarantine zone', where new content is automatically tested before being added to the system (i.e., executed in a sandbox environment). Similarly, a contributor should be able to trace their contribution, to ensure that no unintended modifications are made after submission. Ultimately, challenges arise when anyone can add a new method. Benchmarkers are typically not equally familiar with all included methods, and one would also need to be aware of potential malicious intent (e.g., submitting a suboptimal application of a competing method). On the other hand, many risks can be mitigated by the inherent transparency \cite{Greenstein2016-os}, and feedback loops with (all) original developers could be established to ensure optimal method usage.

A related challenge concerns which datasets and metrics to include in a benchmark. On the one hand, diverse sets of datasets and metrics are important to ensure generalizability and avoid overfitting. On the other hand, inclusion of low-quality or redundant datasets or metrics does not benefit the benchmark, and may provide misleading (aggregate) results. Possible mitigations may include sensitivity analyses, evaluating the impact of single datasets or metrics on the overall, aggregated performance of a method, or reducing redundancy by downweighting similar metrics to avoid a specific aspect dominating the evaluation. From a wider perspective, these questions are all related to the governance of a benchmark, a topic that has not received much attention. Ultimately benchmarks may want to, depending on their scale and complexity, implement different governance models. In one model, the benchmarkers have the ultimate authority to decide on what gets included in the benchmark, simplifying the decision making but creating a risk of gatekeeping. In another model, decisions are made collectively by a group of contributors, which may mitigate the risk of gatekeeping at least partially, but creates a more complex decision structure and introduces the need for a conflict resolution strategy. The governance model of a benchmark could also cover aspects related to computing infrastructure and storage, as well as guidelines to manage authorship of any resulting publications or resources. Having the governance model spelled out when launching a benchmark is likely to reduce friction in later stages of the project. In all cases, an advantage of doing things in public is that contributions are logged.

Neutrality of a benchmark is a highly desirable property and it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods). Although highly desirable, neutrality is nonetheless extremely tough to assess given only the downstream artifacts of a benchmark; this is an area of governance that deserves further attention to increase trust. One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a method. Similarly, ground truth generation and simulations in MDP benchmarks can unconsciously mirror the method capabilities, mechanisms, or strengths, thus biasing evaluation. Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmarks) on analyses that involve their own tools, where neutrality is then hard to establish (e.g., perceived conflict of interest to evaluate one's own method) or where deeper knowledge of their own tools might favor the implementation of the latter compared to other competitors. Here, some recent work to mitigate this conflict involves pre-registration, which typically involves specifying the parameters of your study in advance and posting them on a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial formalization of a benchmark could semi-automatically provide the basis for the formal preregistration. In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, because the goal is not simply to advertise the virtues of a single new method; nonetheless, the challenge to assess neutrality directly remains difficult. Other benchmarking models are possible, from challenges to hackathons and here, there is hope that a consensus of the masses can help to establish neutrality. One potential strategy, in the context of the systematic design of benchmarks, is establishing a system where a benchmark definition is parsed into a pre-registration template. Templates exist for simulation studies that could be adapted more generally for benchmarking projects \cite{Siepe2023-tz}. How to reconcile continuity and pre-registration of benchmarks needs further attention.

Finally, it is worth noting that there is a tension between preregistration and the concept of `continuous' benchmarking, where new datasets, methods and new metrics are added over time. The logic of continuous benchmarking is that as the problem becomes more understood in the field, it becomes increasingly clear what are good ways to evaluate the performance of methods.

\section*{Long-term software reproducibility and data preservation}

Software plays an important role in the current scientific landscape \cite{Howison2015-rc}, since it is widely used throughout the scientific process, including data collection, simulation, analysis and reporting. Researchers are encouraged to publish their data and code to enhance transparency and reproducibility in their work. Ideally, this would increase the adoption rate by the wider scientific community. In practice, however, it is sometimes easier to develop new code than to reuse existing software \cite{Trisovic2022-ol}. This leads to the phenomenon of `academic abandonware', where projects are forgotten in code repositories, exacerbating the reproducibility crisis in research; one of the noted challenges in benchmarking was the lack of extensible frameworks \cite{Sonrel2023-te}. A common reason for abandoned projects is their failure to adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) principles \cite{Wilkinson2016-bh}. Another common cause is the life cycle inherent to academic practice: after delivering a research output, there are often no resources to maintain a software package in the longer term; left alone, the likelihood of a particular library compiling or being able to run after 5-10 years will drastically decrease.

Operational costs are given by the expected compute and storage requirements. In the context of benchmarking, predicting CPU and memory usage is challenging. The methods under evaluation typically have variable computational demands. Benchmark designers could impose resource limits, thereby evaluating the methods under different constraints. Storage costs, on the other hand, are more predictable and can be directly influenced by design decisions of the benchmark (e.g., low retention for re-generatable artifacts, `cold storage' for code archives and software, external data stores such as Zenodo for method performance artifacts, CernVM-FS for reusable software).

\subsection*{Design properties}

A good benchmarking system needs to package its components in a way that ensures that their execution and artifacts can be consistently replicated within practical limits.

The UNIX design philosophy \cite{Pike1984-we} is worth embracing to produce benchmarking systems software that is lightweight, robust, and easily maintainable; ease of use and extensibility are also desirable. Furthermore, to reduce the entry barrier for contributors, the benchmarking system should be capable of running locally with minimal setup. This approach ensures that the system remains relevant and functional even in the absence of centralization. Furthermore, intuitive APIs and sensible defaults \cite{Proctor2018-do} are factors that contribute to highly usable systems. Structured approaches to documentation, including adopting frameworks like diataxis \cite{ProcidaUnknown-vf}, can facilitate adoption and new contributions to a benchmark.

\subsection*{Reproducible software environments}

Software reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}. Computationally, a benchmark executes code that transforms input data into outputs: this execution takes place and is affected by the execution environment; the environment encompasses the base system (OS, compiler toolchains, libraries) and a set of configurations. The benchmark definition should control as much as possible of the execution environment.

Leaving data aside, it can be useful to divide the codebase used into three distinct categories:

\begin{enumerate}
\item The benchmarking system itself has an impact on operational costs, such as storage needs, execution platform, etc. For example, it imposes a certain software stack and base dependencies, and orchestrates the execution of a benchmark plan; usually, the system mandates the choice of a particular workflow manager. At this level, any requirements for input/output formats and shapes are defined; different systems can impose different constraints.

\item Contributions related to individual benchmark components (e.g., datasets, methods, metrics) are expected to be small, typically short scripts that process data and wrap functionality by importing external libraries (where the component implementation itself is developed). Even so, imposing good practices at this level (output validation, testing for abnormal termination, ability to run with a subset of input data) can be beneficial to increase the quality and maintainability of the contributions. The toolset can also enforce validation of metadata annotations (authorship, versioning, licences) for each contributed module.

\item The software dependencies have easily the biggest impact on replicability and maintainability; archiving all combinations of a large dependency tree, across an arbitrary number of base OS images, will exponentially increase retention and maintenance costs. Turning the long-term replicability of a benchmark into a tractable problem is linked to choosing a suitable software management system. Several software management systems have emerged to address these problems, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automating reproducible and efficient software management, from \texttt{EasyBuild} \cite{Hoste2012-gg} and \texttt{Spack} \cite{Gamblin2015-ll} to distributing software installations such as CernVM-FS \cite{Blomer2013-jk} and combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.

\end{enumerate}

\section*{Software licenses and attribution}

Even the simplest benchmark designed and run by a single person aiming to compare their method against various `gold standard' methods incorporates the work of many (dataset generators, method developers, operating systems with system dependencies and their developers). Benchmark artifacts are generated by running a workflow \cite{Jackson2021-ke}. Hence, data, code and operating system dependencies are all accessible during the benchmarking process. For reproducibility and transparency and to build trust \cite{Laine2007-py}, this freedom should be extended for others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how another person is implementing my own method), change (bug fix, contribute) and improve (extend) benchmarks. Such freedoms are granted (or restricted) by software and content licenses, as are copyright and authorship and attribution \cite{Kreutzer2014-ua}. Without explicit licenses, even publicly available code is copyrighted and reuse is restricted (to reproduce, distribute, or create derivative works) \cite{Kreutzer2014-ua}.

In practical terms, a benchmarking system needs to be aware of software and data licenses on import, track further changes and attribute them to the (academic) benchmark contributor. A concise summary of license rights and restrictions, and compatibility across commonly-used open-source licenses can be found in Table~\ref{tab:licenses}. Benchmarking components should be contributed with a clear FOSS (free and open source software) license, perhaps by assigning the copyright holder to contributors describing their name, email and  ORCID (as the canonical scholarly identifier), hence allowing seamless reuse, transparency and easing attribution.

\begin{table}[]
    \caption[Overview of commonly-used open-source licenses in benchmarking studies.]{Overview of commonly used open-source licenses in benchmarking studies. Code or data with restrictions to redistribute or modify are not suitable for collaborative benchmarking. \emph{License}: license name. \emph{Redistribution}: right to redistribute, e.g. reuse in another benchmark/somewhere else. \emph{Modification}: right to modify the code to produce derivative works, e.g. in another benchmark/somewhere else. \emph{Sublicensing}: for derivative works, obligation to keep the same license or freedom to be licensed differently (even copyright). Relevant if the code contributor does not want derivative works to be `closed source' or copyrighted. \emph{Commercial use}: for derivative works, whether commercial usage is restricted or not. Relevant to code contributors not willing commercial reuse of their code (by others). \emph{Attribution}: for derivative works, obligation to attribute original author(s). Relevant to code contributors willing their names/authorship to be credited in derivative works. \emph{Linking}: dynamic or static linking / use as libraries. Particularly important for metrics, as these code can be frequently incorporated (=executed) across benchmarks. }
    \vspace{0.5cm}
    \centering
    \begin{adjustbox}{max width=\textwidth}
    \begin{tabular}{|l|l|l|l|l|l|l|}
    \hline
    License       &  Redistribution       &   Modification   & Sublicensing  & Commercial use & Attribution    & Linking   \\ \hline
    no license	& not allowed	& not allowed	& no sublicensing	& not allowed	& required	& not allowed	 \\
    GPLv3 &   	allowed	 & with restrictions	& GPLv3-compatible	& allowed	& required	& GPLv3-compatible \\
    Apache &	allowed &	allowed	& minimal restrictions	& allowed	& required	& allowed \\
    MIT	& allowed	& allowed	& full freedom	& allowed	& required	& allowed	 \\
    Public Domain	& allowed	& allowed	& full freedom	& allowed	& not required	& allowed  \\
    \textsc{CC-BY-NC-ND}	& allowed	& not allowed&	(no derivatives)	& not allowed	& required &	(not a code license)	 \\
    \textsc{CC-BY-SA-NC}	& allowed	& allowed	& not more restrictive& 	not allowed &	required	& (not a code license)	 \\
    \textsc{CC-BY-SA}	& allowed	& allowed	& not more restrictive &	allowed	& required&	(not a code license)  \\
    \textsc{CC-BY} &	allowed	& allowed	& full freedom 	& allowed& 	required&	(not a code license) \\ \hline

    \end{tabular}
    \end{adjustbox}
    \label{tab:licenses}
\end{table}

%https://docs.google.com/spreadsheets/d/1HvhiDHkY5BSvIJ4607krXIiymk9uGWKsEg3cwaZVfBI/edit?gid=0#gid=0

Similarly, even if all benchmark code were free software, running them could be hindered by (non-free) proprietary interpreters (e.g., MATLAB). Benchmarking systems should be explicit in allowing or disallowing proprietary software requirements, and be aware and track EULAs (end user license agreements). This conundrum extends to the operating system layer and to system dependencies (e.g., compilers). In terms of benchmark re-runnability, fully FOSS system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{Droge2023-ax}.

Finally, data (inputs and outputs) are frequently stored in centralized data repositories \cite{Potter2015-fs, Sicilia2017-ko, van-de-Sandt2019-wd} under content licenses, which might enforce attribution or copyleft. A benchmarking system should automate license metadata handling during data import and export, crediting the original authors if applicable; it also represents an opportunity to enforce that academics use appropriate licenses for data and code.

Scientific benchmarking is rarely a single-person endeavour and has to deal with diverse content and contributions. Software and content licenses standardize the (legally binding) usage rights and restrictions. Benchmarking systems can leverage licenses and version control to track, store and monitor the reuse of data and code in an automated way. On the other hand, contribution attribution in the academic sense (e.g., authorship) is as important and beyond automation. Collaborative benchmarking initiatives should explicitly list guidelines on using and distributing code and results (i.e., licenses) as well as clearly communicate the authorship strategies. These guidelines, complemented with a code of conduct, could be bundled into the benchmark description.

\section*{There are various design tradeoffs}

So far, no single benchmarking system can accommodate all use-cases in bioinformatics. However, documentation of design choices for a given system can help to align expectations. Some important design trade-offs are:

\begin{itemize}
    \item Flexibility vs. complexity: unrestricted freedom for method contributors can mean displacing the burden of maintainability towards the benchmarker. Fewer restrictions (e.g., simpler entry points) means that more methods can be included in a single benchmark; but, it can also imply a higher complexity for the initial setup by the benchmarker. Constraints can also enable validation, which can greatly help with the development process and quality assurance.
    \item  Perfect software reproducibility is possible \cite{Lamb2021-tb}, but it comes with a high engineering cost. It is thus reasonable for a benchmarker to state the desired degree of replicability and the corresponding constraints; different tasks might demand different levels of compliance. 
    \item  Security concerns: decentralized runs lower the participation barrier, but it also increases the potential for attacks to their computing environment. Sandboxed environments (restricting network access, filesystem access etc) can be provided as mitigations. Additionally, static and dynamic code analysis in the continuous integration can help approve potentially risky submissions.
\end{itemize}

\section*{Open data formats and standards}

Benchmarking involves applying methods to data. The types of data can be diverse and even a single dataset can be stored in multiple formats. A broad distinction can be made according to general vs. specific formats and open vs. proprietary formats. Programming language-specific file formats should be avoided since those formats may change over time, security risks can arise \cite{Huynh2023-pq, Bleih2024-lv}, and in general they are not interoperable. Likewise, proprietary formats should be avoided, since they may change and could require specific (proprietary) software to read. General language-agnostic formats often do not change over time and are thus more interoperable such that these datasets can be made FAIR \cite{Wilkinson2016-bh}. Generally, using established standards is recommended (e.g., SAM, BED, FASTQ, etc for genomics), since related methods are developed around these (typically storage-efficient) formats.

If raw input data consists of multiple file formats, an option could be to convert the data into a common format with the caveat that this `intermediate' raw data could be either stored (increased storage usage, easier benchmark definition) or created `on the fly' (increased compute). Furthermore, how raw input data is retrieved (e.g., automatic retrieval once and store locally, automatic retrieval every time, or manual retrieval and store locally) needs to be decided.

Additionally, which level and scope of data validation needs to be decided. Validations include content integrity (e.g., md5sum), data syntax (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data). Intermediate data and results should be made FAIR, leading to additional decisions, including the scope of data sharing (e.g., only the most important subset), the handling of the metadata (automatic versus manual annotation) and technical aspects (where and how to make available, e.g., GEO or Zenodo). Automatic generation of metadata might require predefined standards and an ontology to fully track how the data was generated \cite{LeboUnknown-om}.

\section*{Conclusion}
\label{sec:conclusion}

Benchmarking is a cornerstone in bioinformatics, impacting how methods are conceived, developed, published and chosen. Benchmarking is driven by the effort of communities of people with the expertise to address similar (biological and technical) problems. This expertise is leveraged by current practices of community benchmarking, so scientific advance is eased by joint efforts to catalog open problems, generate data to depict (or simulate) them, develop methods to address them, and evaluate their performances. Community benchmarking has been proved successful and very productive \cite{Moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.

Here we discuss the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting technologies to potentially ease the technical, scientific and contribution challenges.

\hfill

\section*{Acknowledgments}

We thank the Robinsonlab members for their constructive feedback.

\section*{Author contributions}

Conceptualization: all. Visualization: BC. Original draft preparation: all. Writing, reviewing and editing: all. Supervision: IM, MDR. Funding acquisition: MDR.

\section*{Funding}

MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36_14). CS is supported by the Novartis Research Foundation.

\section*{Disclosure statement}

No potential conflict of interest is reported by the authors

\pagebreak
\linespread{1.1} 
\printbibliography

\end{document}

### 2024-09-23 manifesto-main.tex
more

@@
 Benchmarking, which involves collecting reference datasets and demonstrating method performances, is a requirement for the development of new computational tools, but also becomes a domain of its own to achieve neutral comparisons of methods.
-Although a lot has been written about how to design and conduct benchmark studies, this perspective sheds light on a wish list for a computational platform to orchestrate benchmark studies.
+Although a lot has been written about how to design and conduct benchmark studies, this Perspective sheds light on a wish list for a computational platform to orchestrate benchmark studies.
 We discuss various ideas for organizing reproducible software environments, formally defining benchmarks, orchestrating standardized workflows, and how they interface with computing infrastructure.
@@
 In this work, we give our perspective on how to build a high-standard benchmarking ecosystem in the field of bioinformatics, coming primarily from the viewpoint of methodologists (i.e., researchers with computational background working on tools for data).
-In particular, our case is one of computational scientists embedded in data-rich biology (i.e., computational biology and bioinformatics), where rapid technology development is a distinguishing factor (e.g., single-cell assays coupled to high-throughput sequencing, imaging, etc.); this paucity in turn results in a rapid response from the computational community to develop tools, models, and software for the analysis of such emerging datasets.
+In particular, our case is one of computational scientists embedded in data-rich biology (i.e., computational biology and bioinformatics), where one of the distinguishing factors is rapid technology development (e.g., single-cell assays coupled to high-throughput sequencing, imaging, etc.), which in turn results in a rapid response from the computational community to develop tools, models, and software for the analysis of such emerging datasets.
 Although our viewpoint is `bioinformatics', there are analogs in adjacent disciplines where methodologists are creating computational tools in a certain scientific context.
@@
 
-Our goal is to lay out general benchmark parameters and definitions,  while abstracting towards an ideal \textit{benchmarking system}.
+Our goal here is to lay out general benchmark parameters and definitions, while abstracting towards an ideal \textit{benchmarking system}.
 This ranges from general benchmark conceptualization to various technical specifications.
@@
 
-Benchmark artifacts can be seen as a collection of data files and (perhaps diverse) source code implementations assembled in a pipeline topology (e.g., methods are run on some input data, and method outputs act as input for metrics) that were run as a workflow in some computing environment.
+Benchmark artifacts can be seen as a collection of data files and (perhaps diverse) source code implementations with a pipeline topology (e.g., methods are run on some input data, and method outputs act as input for metrics) that were ideally run as a workflow in some computing environment.
 Current practices in single-cell RNA-seq benchmarking show that code is frequently open, reusable and versioned; input datasets are also typically available \cite{Sonrel2023-te}.
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and ideally efficient software stack; managing storage and their access-control lists; rendering results dashboards and summarizing benchmarking results; version controlling code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and efficient software stack; managing storage and their access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results.
 Benchmarking systems are in charge of orchestrating both workflow execution and these benchmarking-specific tasks.
@@
 Additionally, ground truths are not always well-defined (e.g., derived with uncertainty from other experimental data), multiple tasks (e.g., represented by different reference datasets) may be evaluated for a single method and several ways to assess performance are possible.
-Even when accompanied by dataset- and/or metric-specific performance measures, scaling and aggregation of results makes it difficult to unpack fine-grained nuances of performance from a final summary table reported in most benchmarks (e.g., \footnote{\url{https://www.nature.com/articles/s41592-021-01336-8/figures/11}}).
+Even when accompanied by dataset- and/or metric-specific performance measures, scaling and aggregation of results make it difficult to unpack fine-grained nuances of performance from a final summary table reported in most benchmarks (e.g., \footnote{\url{https://www.nature.com/articles/s41592-021-01336-8/figures/11}}).
 Strobl \textit{et al.} push back against the `one method fits all datasets’ philosophy and advocate for highlighting dataset or parameter properties that are predictive of method performance \cite{Jelizarow2010-kr, Boulesteix2010-zj, Strobl2024-lp}.
@@
 Altogether, evaluating parameters as part of a benchmark can give useful insights, but should be linked to a distinct and independent research objective.
+
 For continuous and collaborative benchmarking, parameter tuning and usage should be governed and coordinated between all contributors.
@@
 In designing a benchmarking system, it may be helpful to specify whether the benchmark is intended to be run `centralized' or `decentralized'.
-These terms can take different meanings depending on the context; we will focus on the organization of the benchmark infrastructure.
-In most systems, the code to execute will be tracked in a reference git repository (e.g., on GitHub, GitLab).
-The workflow execution and the storage of derived datasets (as various stages of results) are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery platforms (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., could be a laptop, server or cloud).
+These terms can take different meanings depending on the context, but here the benchmark infrastructure location is intended.
+In most systems, the code to execute will be tracked in a reference \texttt{git} repository (e.g., on GitHub, GitLab).
+The workflow execution and the storage of derived datasets, and various stages of results are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery platforms (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., could be a laptop, server or cloud).
 
@@
 Thus, users need a system that can build benchmark components in both a centralized and decentralized fashion, i.e., starting decentralized and promoting a benchmark to centralized when it becomes sufficiently mature.
-Accordingly, the process of centralization is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to resources) and adding measures to build trust and transparency.
+Accordingly, the process of centralization is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to computational resources) and adding measures to build trust and transparency.
 
@@
 The governance model of a benchmark could also cover aspects related to computing infrastructure and storage, as well as guidelines to manage authorship of any resulting publications or resources.
-Having the governance model spelled out when launching a benchmark is likely to reduce friction in later stages of the project.
+Having the governance model spelled out when launching a benchmark is likely to reduce friction at later stages of the project.
 In all cases, an advantage of doing things in public is that contributions are logged.
 
-Neutrality of a benchmark is a highly desirable property and it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods).
+Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods).
 Although highly desirable, neutrality is nonetheless extremely tough to assess given only the downstream artifacts of a benchmark; this is an area of governance that deserves further attention to increase trust.
@@
 Similarly, ground truth generation and simulations in MDP benchmarks can unconsciously mirror the method capabilities, mechanisms, or strengths, thus biasing evaluation.
-Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmarks) on analyses that involve their own tools, where neutrality is then hard to establish (e.g., perceived conflict of interest to evaluate one's own method) or where deeper knowledge of their own tools might favor the implementation of the latter compared to other competitors.
+Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then hard to establish (e.g., a perceived conflict of interest to evaluate one's own method) or where a deeper knowledge of their own tools might favor the implementation of the latter compared to other competitors.
 Here, some recent work to mitigate this conflict involves pre-registration, which typically involves specifying the parameters of your study in advance and posting them on a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial formalization of a benchmark could semi-automatically provide the basis for the formal preregistration.
@@
 Researchers are encouraged to publish their data and code to enhance transparency and reproducibility in their work.
-Ideally, this would increase the adoption rate by the wider scientific community.
+Ideally, this would increase benchmark (software) reuse by the wider scientific community.
 In practice, however, it is sometimes easier to develop new code than to reuse existing software \cite{Trisovic2022-ol}.
@@
 
-The UNIX design philosophy \cite{Pike1984-we} is worth embracing to produce benchmarking systems software that is lightweight, robust, and easily maintainable; ease of use and extensibility are also desirable.
+The UNIX design philosophy \cite{Pike1984-we} is worth embracing to produce software for benchmarking systems that are lightweight, robust, and easily maintainable; ease of use and extensibility are also desirable.
 Furthermore, to reduce the entry barrier for contributors, the benchmarking system should be capable of running locally with minimal setup.
@@
 Even so, imposing good practices at this level (output validation, testing for abnormal termination, ability to run with a subset of input data) can be beneficial to increase the quality and maintainability of the contributions.
-The toolset can also enforce validation of metadata annotations (authorship, versioning, licences) for each contributed module.
+The toolset can also enforce the validation of metadata annotations (authorship, versioning, licences) for each contributed module.
 
@@
 Turning the long-term replicability of a benchmark into a tractable problem is linked to choosing a suitable software management system.
-Several software management systems have emerged to address these problems, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automating reproducible and efficient software management, from \texttt{EasyBuild} \cite{Hoste2012-gg} and \texttt{Spack} \cite{Gamblin2015-ll} to distributing software installations such as CernVM-FS \cite{Blomer2013-jk} and combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.
+Several software management systems have emerged to address these problems, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.
 
@@
 Such freedoms are granted (or restricted) by software and content licenses, as are copyright and authorship and attribution \cite{Kreutzer2014-ua}.
-Without explicit licenses, even publicly available code is copyrighted and reuse is restricted (to reproduce, distribute, or create derivative works) \cite{Kreutzer2014-ua}.
+Without explicit licenses, even publicly available code is copyrighted by default, and reuse is restricted (to reproduce, distribute, or create derivative works) \cite{Kreutzer2014-ua}.
 
-In practical terms, a benchmarking system needs to be aware of software and data licenses on import, track further changes and attribute them to the (academic) benchmark contributor.
+In practical terms, a benchmarking system needs to be aware of software and data licenses, track further changes and attribute them to the (academic) benchmark contributor.
 A concise summary of license rights and restrictions, and compatibility across commonly-used open-source licenses can be found in Table~\ref{tab:licenses}.
-Benchmarking components should be contributed with a clear FOSS (free and open source software) license, perhaps by assigning the copyright holder to contributors describing their name, email and  ORCID (as the canonical scholarly identifier), hence allowing seamless reuse, transparency and easing attribution.
+Benchmark contributions should bring a clear FOSS (free and open source software) license, perhaps by assigning the copyright holder to contributors describing their name, email and  ORCID (as the canonical scholarly identifier), hence allowing seamless reuse, transparency and easing attribution.
 
@@
 
-Similarly, even if all benchmark code were free software, running them could be hindered by (non-free) proprietary interpreters (e.g., MATLAB).
+Similarly, even if all benchmark code were free software, running it could be hindered by (non-free) proprietary interpreters (e.g., MATLAB).
 Benchmarking systems should be explicit in allowing or disallowing proprietary software requirements, and be aware and track EULAs (end user license agreements).
@@
 
-\section*{There are various design tradeoffs}
+\section*{There are various hidden design tradeoffs}
 
@@
 It is thus reasonable for a benchmarker to state the desired degree of replicability and the corresponding constraints; different tasks might demand different levels of compliance.
-    \item  Security concerns: decentralized runs lower the participation barrier, but it also increases the potential for attacks to their computing environment.
-Sandboxed environments (restricting network access, filesystem access etc) can be provided as mitigations.
-Additionally, static and dynamic code analysis in the continuous integration can help approve potentially risky submissions.
+    \item  Security concerns: decentralized runs lower the participation barrier, but it also increases the potential for attacks on their computing environment.
+Sandboxed environments (restricting network access, filesystem access etc) can be provided as mitigation.
+Additionally, static and dynamic code analysis in the continuous integration process can aid in approving potentially risky submissions.
 \end{itemize}
@@
 proprietary formats.
-Programming language-specific file formats should be avoided since those formats may change over time, security risks can arise \cite{Huynh2023-pq, Bleih2024-lv}, and in general they are not interoperable.
+File formats specific to a programming language should be avoided since those formats may change over time, security risks can arise \cite{Huynh2023-pq, Bleih2024-lv}, and in general they are not interoperable.
 Likewise, proprietary formats should be avoided, since they may change and could require specific (proprietary) software to read.
@@
 Validations include content integrity (e.g., md5sum), data syntax (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data).
-Intermediate data and results should be made FAIR, leading to additional decisions, including the scope of data sharing (e.g., only the most important subset), the handling of the metadata (automatic versus manual annotation) and technical aspects (where and how to make available, e.g., GEO or Zenodo).
+Intermediate data and results should be made FAIR, leading to additional decisions, including the scope of data sharing (e.g., only the most important subset), the handling of the metadata (automatic versus manual annotation) and the provider (where and how to make available, e.g., GEO or Zenodo).
 Automatic generation of metadata might require predefined standards and an ontology to fully track how the data was generated \cite{LeboUnknown-om}.
@@
 
-Benchmarking is a cornerstone in bioinformatics, impacting how methods are conceived, developed, published and chosen.
-Benchmarking is driven by the effort of communities of people with the expertise to address similar (biological and technical) problems.
-This expertise is leveraged by current practices of community benchmarking, so scientific advance is eased by joint efforts to catalog open problems, generate data to depict (or simulate) them, develop methods to address them, and evaluate their performances.
+Benchmarking is a cornerstone practice in bioinformatics, impacting how methods are conceived, developed, published and chosen.
+Benchmarking is driven by the effort of people with the expertise to address (biological and technical) problems.
+This expertise can be leveraged by current practices of community benchmarking, so scientific advance is eased by joint efforts to catalog open problems, generate (or simulate) appropriate input data, develop methods to address them, and evaluate their performances.
 Community benchmarking has been proved successful and very productive \cite{Moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.
 
-Here we discuss the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting technologies to potentially ease the technical, scientific and contribution challenges.
+Here we discussed the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting technologies to potentially ease the technical, scientific and contribution challenges.
 
@@
 
-MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36_14).
+MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36\_14).
 CS is supported by the Novartis Research Foundation.

### 2024-09-23 manifesto-main.tex
funkyheatmaps

@@
 Additionally, ground truths are not always well-defined (e.g., derived with uncertainty from other experimental data), multiple tasks (e.g., represented by different reference datasets) may be evaluated for a single method and several ways to assess performance are possible.
-Even when accompanied by dataset- and/or metric-specific performance measures, scaling and aggregation of results make it difficult to unpack fine-grained nuances of performance from a final summary table reported in most benchmarks (e.g., \footnote{\url{https://www.nature.com/articles/s41592-021-01336-8/figures/11}}).
+Even when accompanied by dataset- and/or metric-specific performance measures, scaling and aggregation of results make it difficult to unpack fine-grained nuances of performance from a final summary table reported as a FunkyHeatmap \cite{funkyheatmap} as in most benchmarks (e.g., \cite{Saelens2019-jy}).
 Strobl \textit{et al.} push back against the `one method fits all datasets’ philosophy and advocate for highlighting dataset or parameter properties that are predictive of method performance \cite{Jelizarow2010-kr, Boulesteix2010-zj, Strobl2024-lp}.
@@
 Being public and open to contributions from specialists of different fields, a benchmarking system could gather more reference datasets and methods, and allow the current state of the field to be better grasped.
+
 %As a result, the performance results will tend to better fit to the current state of the field, as opposed to a static benchmark led by a small panel of experts that leads to a smaller, and sometimes (un)intentionally biased view of the field \cite{Boulesteix2010-zj, Weber2019-el}.
@@
 On the other hand, inclusion of low-quality or redundant datasets or metrics does not benefit the benchmark, and may provide misleading (aggregate) results.
-Possible mitigations may include sensitivity analyses, evaluating the impact of single datasets or metrics on the overall, aggregated performance of a method, or reducing redundancy by downweighting similar metrics to avoid a specific aspect dominating the evaluation.
+Possible mitigations may include sensitivity analyses, evaluating the impact of single datasets or metrics on the overall, aggregated performance of a method, or reducing redundancy by down-weighting similar metrics to avoid a specific aspect dominating the evaluation.
 From a wider perspective, these questions are all related to the governance of a benchmark, a topic that has not received much attention.
@@
 Additionally, which level and scope of data validation needs to be decided.
-Validations include content integrity (e.g., md5sum), data syntax (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data).
-Intermediate data and results should be made FAIR, leading to additional decisions, including the scope of data sharing (e.g., only the most important subset), the handling of the metadata (automatic versus manual annotation) and the provider (where and how to make available, e.g., GEO or Zenodo).
+Validations include content integrity (e.g., md5sum), adherence to a data specification (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data).
+Intermediate data and results should be made FAIR, leading to additional decisions, including the scope of data sharing (e.g., only the most important subset), the handling of the metadata (automatic versus manual annotation) and the data storage provider and policies (where and how to make available, e.g., GEO or Zenodo).
 Automatic generation of metadata might require predefined standards and an ontology to fully track how the data was generated \cite{LeboUnknown-om}.

### 2024-09-23 manifesto-main.tex
sync
Comment out declarefieldformats as arxiv doesn't like them; add biber bbl
Vancouver style
lost refs

@@
 \usepackage{microtype}
+%\usepackage{url}
+\PassOptionsToPackage{hyphens}{url}
+\usepackage{hyperref}
 
@@
 \usepackage[english]{babel}
-\usepackage[style=authoryear,backend=biber,natbib=true,maxcitenames=2,uniquelist=false]{biblatex}
-\usepackage{natbib}
-\addbibresource{manifesto-main.bib} % your .bib file
-
-% Customizing biblatex for Harvard style
-% Customizing biblatex for Harvard style
-\DeclareNameAlias{sortname}{family-given}
-\DeclareNameAlias{default}{family-given}
-
-\renewbibmacro{in:}{}
-\DeclareFieldFormat[article]{title}{\mkbibquote{#1}\addcomma}
-\DeclareFieldFormat[book]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[bookinbook]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[inbook]{title}{\mkbibquote{#1}\addcomma}
-\DeclareFieldFormat[incollection]{title}{\mkbibquote{#1}\addcomma}
-\DeclareFieldFormat[inproceedings]{title}{\mkbibquote{#1}\addcomma}
-\DeclareFieldFormat[manual]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[misc]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[thesis]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[unpublished]{title}{\mkbibquote{#1}\addcomma}
-\DeclareFieldFormat[patent]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[report]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[online]{title}{\mkbibquote{#1}\addcomma}
-\DeclareFieldFormat[software]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[booklet]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[periodical]{title}{\mkbibemph{#1}\addcomma}
-\DeclareFieldFormat[standard]{title}{\mkbibemph{#1}\addcomma}
-
-\DeclareFieldFormat[article]{journaltitle}{\iffieldundef{shortjournal}{\mkbibemph{#1}\addcomma}{\mkbibemph{\printfield{shortjournal}}\addcomma}}
-\DeclareFieldFormat{volume}{\bibstring{volume}~#1}
-\DeclareFieldFormat{number}{\bibstring{number}~#1}
-
-% Definitions for "Vol." and "No."
-\DefineBibliographyStrings{english}{
-  volume = {Vol.},
-  number = {No.}
-}
-
-\renewbibmacro*{volume+number+eid}{%
-  \printfield{volume}%
-  \setunit*{\addspace}%
-  \printfield{number}%
-  \setunit{\addcomma\space}%
-  \printfield{eid}}
-
-\renewbibmacro*{journal+issuetitle}{%
-  \usebibmacro{journal}%
-  \setunit*{\addcomma\space}%
-  \usebibmacro{volume+number+eid}%
-  \setunit{\addcomma\space}%
-  \usebibmacro{issue+date}}
-
-\renewbibmacro*{publisher+location+date}{%
-  \printlist{publisher}%
-  \iflistundef{location}
-    {\setunit*{\addcomma\space}}
-    {\setunit*{\addcolon\space}}%
-  \printlist{location}%
-  \setunit*{\addcomma\space}%
-  \usebibmacro{date}}
-
-\renewcommand*{\bibpagespunct}{\addcomma\space}
-
-% Customizing page field format to prevent duplication
-% \DeclareFieldFormat{pages}{%
-%   \mkfirstpage[{\mkpageprefix[page]{#1}}]{#1}}
-
-% Customizing citations for Harvard style
-\DeclareCiteCommand{\cite}[\mkbibparens]
-  {\usebibmacro{prenote}}
-  {\usebibmacro{citeindex}%
-   \usebibmacro{cite}}
-  {\multicitedelim}
-  {\usebibmacro{postnote}}
-
-\renewbibmacro*{cite:labelyear+extrayear}{%
-  \iffieldundef{labelyear}
-    {}
-    {\printtext[bibhyperref]{%
-       \printfield{labelyear}%
-       \printfield{extrayear}}}}
-
-\renewbibmacro*{cite:labeldate+extradate}{%
-  \iffieldundef{labelyear}
-    {}
-    {\printtext[bibhyperref]{%
-       \printfield{labelyear}%
-       \printfield{extradate}}}}
-
-\AtEveryBibitem{
-  \clearfield{month}
-  \clearfield{day}
-  \ifentrytype{book}{
-    \clearlist{location}
-  }{}
-}
-
-% Formatting "et al." in italics followed by a comma
-\DefineBibliographyStrings{english}{
-  andothers = {\textit{et al.},}
-}
-
-\DeclareFieldFormat[article]{volume}{\bibstring{jourvol}\addnbspace #1}
-\DeclareFieldFormat[article]{number}{\bibstring{number}\addnbspace #1}
-\DeclareFieldFormat[article]{volume}{Vol. #1}
-\DeclareFieldFormat[article]{number}{No. #1}
-% Customizing DOI field format to lowercase "doi"
-%\DeclareFieldFormat{doi}{\bibstring{doi}\addcolon\space\url{#1}}
-
-% Customizing URL field format to "available at:"
-\DeclareFieldFormat{url}{\bibstring{available at}\addcolon\space\url{#1}}
-\DeclareFieldFormat{urldate}{\mkbibparens{accessed \addspace#1}}
-
-% Customizing urldate to match the required format
-\DeclareFieldFormat{urldate}{%
-  \mkbibparens{accessed\space%
-    \thefield{urlday}\addspace%
-    \mkbibmonth{\thefield{urlmonth}}\addspace%
-    \thefield{urlyear}}}
-
-% Configure cleveref
-\crefformat{figure}{#2Figure~#1#3}
-\Crefformat{figure}{#2Figure~#1#3}
-\crefformat{table}{#2Table~#1#3}
-\Crefformat{table}{#2Table~#1#3}
-\crefformat{section}{#2Section~#1#3}
-\Crefformat{section}{#2Section~#1#3}
 
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{figures/bens_formalization.pdf}
+    \includegraphics[width=0.75\linewidth]{figures/bens_formalization.pdf}
     \caption[Benchmark formalization.]{Benchmark formalization.
@@
 
-Besides dashboards for more nuanced result exploration \cite{MariniUnknown-aq}, multi-criteria decision analysis (MCDA) \cite{Taherdoost2023-wd} has been proposed to guide users through benchmark results and multidimensional scaling (MDS) has been used to differentiate the effect of single datasets or metrics and recognize patterns beyond aggregated performance \cite{Niessl2022-fk}.
+Besides dashboards for more nuanced result exploration \cite{bettr}, multi-criteria decision analysis (MCDA) \cite{Taherdoost2023-wd} has been proposed to guide users through benchmark results and multidimensional scaling (MDS) has been used to differentiate the effect of single datasets or metrics and recognize patterns beyond aggregated performance \cite{Niessl2022-fk}.
 An extensive analysis and presentation of the results not only helps the users in the long term but also the whole scientific field to identify pain points, the complexity of possible scenarios and to isolate the different use-cases.
@@
 \linespread{1.1} 
-\printbibliography
+%\printbibliography
+
+\bibliographystyle{vancouver}  
+\bibliography{clean} % your .bib file
 

### 2024-10-21 manifesto-main.tex
fix
No lineno
Shortening

@@
 
-\linenumbers
+%\linenumbers
 
@@
 In this work, we give our perspective on how to build a high-standard benchmarking ecosystem in the field of bioinformatics, coming primarily from the viewpoint of methodologists (i.e., researchers with computational background working on tools for data).
-In particular, our case is one of computational scientists embedded in data-rich biology (i.e., computational biology and bioinformatics), where one of the distinguishing factors is rapid technology development (e.g., single-cell assays coupled to high-throughput sequencing, imaging, etc.), which in turn results in a rapid response from the computational community to develop tools, models, and software for the analysis of such emerging datasets.
 Although our viewpoint is `bioinformatics', there are analogs in adjacent disciplines where methodologists are creating computational tools in a certain scientific context.
-One critical step in method development is benchmarking (further defined below), whereby either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; for these, we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, defined in a recent meta-analysis study \cite{Cao2023-jz}.
-Notably, there are various ways to structure the running of a benchmarking study: it can be run by a single person or a small group on a laptop (or a server), or it can be organized as a `challenge' or within the duration of a hackathon (groups of people that come together to competitively assess their approaches and/or populate benchmark components).
+One critical step in method development is benchmarking (further defined below), whereby either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; for these, we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, as defined in \cite{Cao2023-jz}.
+Notably, a benchmarking study can be run by a single person or a small group on a laptop (or a server), or it can be organized as a `challenge' or within the duration of a hackathon.
 
-Our goal here is to lay out general benchmark parameters and definitions, while abstracting towards an ideal \textit{benchmarking system}.
-This ranges from general benchmark conceptualization to various technical specifications.
-We will offer some remarks on the philosophy around the benchmarking activity (e.g., gatekeeping), as well as some technical recommendations (e.g., standards).
+Our goal here is to lay out an ideal \textit{benchmarking system}, from general benchmark conceptualization to various technical specifications.
+We offer some remarks on the philosophy around benchmarking (e.g., gatekeeping), as well as some technical recommendations (e.g., standards).
 
-\section{What is a benchmark?}
+\section*{What is a benchmarking system?}
 %\label{sec:main_section}
 
-To begin our discussion of benchmarking (in bioinformatics), a few definitions are needed.
-For example, what actually is a benchmark, precisely?
+What actually is a benchmark, precisely?
 Here, a benchmark is thought of as a conceptual entity that roughly encompasses all the components of a study to understand the performance of a set of (computational) methods for a given task (here, the primary context is computational methods, but the concept can have a broader scope, e.g., evaluation of data-generating technologies).
-A benchmark requires a well-defined task, such as inferring what set of genes are differentially expressed from a certain type of data (e.g., gene expression from RNA sequencing counts) or whether a high-dimensional dataset can be clustered into the `correct' cell types (the definition of this correctness, or ground-truth, should be precisely defined in advance).
-Given a well-defined task, a benchmark consists of benchmark `components' (such as simulations to generate datasets, preprocessing steps, methods, metrics).
-Ideally, a benchmarking system is used to organize and orchestrate the running of the benchmark, and create `artifacts' (code snapshots, file outputs to be shared, results tables).
-Ultimately, the downstream results and interpretations, which may involve rankings and various additional analyses, will be derived from these artifacts.
+A benchmark requires a well-defined task and a definition of correctness, or ground-truth, precisely defined in advance.
+Then, a benchmark consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, metrics.
+
+Can benchmarking be systematized?
+Benchmarks should be orchestrated and run by `benchmarking system's to create `artifacts' (code snapshots, file outputs to be shared, performance outputs).
 The various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
@@
 corrections or new components), such that the benchmarks of the future can run more continuously.
-
 At the outset of a benchmark, a benchmark `definition' could be envisioned, which could give a formal specification of the entire set of components (and the pattern of artifacts to be generated).
@@
 For a \emph{data analyst}, with an aim to perform a specific type of analysis on a specific dataset, benchmarks are helpful tools to find a suitable method for the task.
-However, it is important that the benchmarks consider datasets that are similar to the one of interest to the analyst, as method performance often depends on characteristics of the data.
+Hence, it is important that benchmarks include datasets similar to the one of interest to the analyst, as method performance often depends on characteristics of the data.
 Thus, a thorough characterization of the datasets considered in a benchmark is desirable.
@@
 
-Benchmark artifacts can be seen as a collection of data files and (perhaps diverse) source code implementations with a pipeline topology (e.g., methods are run on some input data, and method outputs act as input for metrics) that were ideally run as a workflow in some computing environment.
+Benchmarks can be seen as a collection of data files and (perhaps diverse) source code ideally run as a workflow in some computing environment.
 Current practices in single-cell RNA-seq benchmarking show that code is frequently open, reusable and versioned; input datasets are also typically available \cite{Sonrel2023-te}.
 However, the \textit{extensibility} of benchmarking studies is generally very low, as well as the proportion of benchmarks using a formal workflow system.
-Thus, a formal system to organize components and orchestrate the crossing of datasets, methods, and metrics would largely impact the current landscape.
+Thus, a formal system to orchestrate benchmarking workflows would largely impact the current landscape.
 
-In an ideal situation, running a benchmark includes executing a workflow, a process for which over 350 workflow languages, platforms or systems  \cite{Wratten2021-el, Amstutz2024-qk} and one standard named Common Workflow Language (CWL) (with multiple implementations) exist \cite{Amstutz2016-vo}.
+Over 350 workflow languages, platforms or systems  \cite{Wratten2021-el, Amstutz2024-qk} and one standard named Common Workflow Language (CWL) (with multiple implementations) exist \cite{Amstutz2016-vo}.
 Some workflow languages can export workflow definitions to CWL, hence helping computational FAIR guidelines: that is, bundling and processing data in association to their metadata, tracking new metadata generation, recording provenance, and being accessible \cite{Goble2020-ps}.
-Both the workflow layout and provenance can be formally defined (Figure~\ref{fig:formalization}).
+For benchmarking, we believe both the workflow layout and provenance can be formally defined (Figure~\ref{fig:formalization}).
 
@@
 While this is valid in a few fields with clearly defined goals, such as image recognition \cite{Khan2018-ot} or protein structure prediction \cite{Jumper2021-oq}, bioinformatics tasks are typically evaluated by multiple metrics, some of which are not well understood \cite{Lutge2021-mt, Reinke2024-dw}.
-Additionally, ground truths are not always well-defined (e.g., derived with uncertainty from other experimental data), multiple tasks (e.g., represented by different reference datasets) may be evaluated for a single method and several ways to assess performance are possible.
+Additionally, ground truths are not always well-defined (e.g., derived with uncertainty from other experimental data) and multiple tasks (e.g., represented by different reference datasets) may be evaluated for a single method and several ways to assess performance are possible.
 Even when accompanied by dataset- and/or metric-specific performance measures, scaling and aggregation of results make it difficult to unpack fine-grained nuances of performance from a final summary table reported as a FunkyHeatmap \cite{funkyheatmap} as in most benchmarks (e.g., \cite{Saelens2019-jy}).
@@
 This aligns with what a data analyst wants: to identify the best performers in settings most similar to their problem at hand.
-This requires not only a thorough benchmark design, e.g., tracking of meaningful dataset metadata \cite{Strobl2024-lp}, but also can be supported by a benchmarking system that facilitates flexible navigation of results (e.g., via an interactive dashboard).
+This requires not only a thorough benchmark design, e.g., tracking of meaningful dataset metadata \cite{Strobl2024-lp}, but also can be supported by a benchmarking system that facilitates flexible navigation of results.
 
-Besides dashboards for more nuanced result exploration \cite{bettr}, multi-criteria decision analysis (MCDA) \cite{Taherdoost2023-wd} has been proposed to guide users through benchmark results and multidimensional scaling (MDS) has been used to differentiate the effect of single datasets or metrics and recognize patterns beyond aggregated performance \cite{Niessl2022-fk}.
+Besides interactive dashboards for more nuanced result exploration \cite{bettr}, multi-criteria decision analysis (MCDA) \cite{Taherdoost2023-wd} has been proposed to guide users through benchmark results and multidimensional scaling (MDS) has been used to differentiate the effect of single datasets or metrics and recognize patterns beyond aggregated performance \cite{Niessl2022-fk}.
 An extensive analysis and presentation of the results not only helps the users in the long term but also the whole scientific field to identify pain points, the complexity of possible scenarios and to isolate the different use-cases.
@@
 Being public and open to contributions from specialists of different fields, a benchmarking system could gather more reference datasets and methods, and allow the current state of the field to be better grasped.
-
-%As a result, the performance results will tend to better fit to the current state of the field, as opposed to a static benchmark led by a small panel of experts that leads to a smaller, and sometimes (un)intentionally biased view of the field \cite{Boulesteix2010-zj, Weber2019-el}.
 
@@
 In designing a benchmarking system, it may be helpful to specify whether the benchmark is intended to be run `centralized' or `decentralized'.
-These terms can take different meanings depending on the context, but here the benchmark infrastructure location is intended.
+These terms can take different meanings depending on the context, but here the benchmark infrastructure location is considered.
 In most systems, the code to execute will be tracked in a reference \texttt{git} repository (e.g., on GitHub, GitLab).
-The workflow execution and the storage of derived datasets, and various stages of results are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery platforms (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., could be a laptop, server or cloud).
+The workflow execution and the storage of derived datasets, and various stages of results are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery platforms (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., could be a laptop, server or cloud; e.g.
+omnibenchmark \cite{omnibenchmark}).
 
 Traditionally, benchmarks (whether BOP or MDP) are conducted in a decentralized fashion and typically by a small number of researchers (e.g., for MDP, the new method developer).
-It is worth mentioning some disadvantages of the decentralized approach.
+This brings some disadvantages.
 For example, decentralization can be hugely inefficient; researchers studying the same task would collect reference datasets and code snippets to run methods, and implement a bespoke strategy (e.g., shell scripts) to orchestrate the methods run on datasets.
@@
 
-Arguably, most benchmarks could naturally start out as decentralized, in part as a means to properly define the concept and problem that one wants to work on, but also as a way to generate preliminary results and ideas to solidify the design before `going public' (centralization).
+Arguably, most benchmarks could start out as decentralized, in part as a means to properly define the concept and problem, but also as a way to generate preliminary results and ideas to solidify the design before `going public' (centralization).
 Thus, users need a system that can build benchmark components in both a centralized and decentralized fashion, i.e., starting decentralized and promoting a benchmark to centralized when it becomes sufficiently mature.
-Accordingly, the process of centralization is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to computational resources) and adding measures to build trust and transparency.
+Accordingly, the process of centralization is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to computational resources) and adding measures to build trust \textit{via} transparency.
 
@@
 
-\bibliographystyle{vancouver}  
+\bibliographystyle{vancouver} % elsarticle-num} % vancouver}  
 \bibliography{clean} % your .bib file

### 2024-10-22 manifesto-main.tex
first round of trimming

@@
 
-In this work, we give our perspective on how to build a high-standard benchmarking ecosystem in the field of bioinformatics, coming primarily from the viewpoint of methodologists (i.e., researchers with computational background working on tools for data).
-Although our viewpoint is `bioinformatics', there are analogs in adjacent disciplines where methodologists are creating computational tools in a certain scientific context.
-One critical step in method development is benchmarking (further defined below), whereby either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; for these, we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, as defined in \cite{Cao2023-jz}.
+In this work,  we present a perspective on building a robust benchmarking ecosystem in bioinformatics, coming primarily from the methodologist's viewpoint (researchers developing computational tools).
+While our focus is bioinformatics, we aim to abstract principles and connect to similar challenges in adjacent  fields.
+Benchmarking  (further defined below) is a critical step in method development: either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, as defined in \cite{Cao2023-jz}.
 Notably, a benchmarking study can be run by a single person or a small group on a laptop (or a server), or it can be organized as a `challenge' or within the duration of a hackathon.
 
-Our goal here is to lay out an ideal \textit{benchmarking system}, from general benchmark conceptualization to various technical specifications.
-We offer some remarks on the philosophy around benchmarking (e.g., gatekeeping), as well as some technical recommendations (e.g., standards).
+In our discussion, we outline concepts and specifications for an ideal \textit{benchmarking system}; we examine philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization practices.
 
@@
 
-What actually is a benchmark, precisely?
-Here, a benchmark is thought of as a conceptual entity that roughly encompasses all the components of a study to understand the performance of a set of (computational) methods for a given task (here, the primary context is computational methods, but the concept can have a broader scope, e.g., evaluation of data-generating technologies).
-A benchmark requires a well-defined task and a definition of correctness, or ground-truth, precisely defined in advance.
+We define a benchmark as a conceptual framework to evaluate the performance of computational methods for a given task, though it can extend to assessing other aspects, like data-generating technologies.
+A benchmark thus requires a well-defined task and a definition of correctness, or ground-truth, to be precisely defined in advance.
 Then, a benchmark consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, metrics.
@@
 
-For a \emph{method developer}, benchmarking serves the purpose of comparing their method to the state of the art in the field.
-Ideally, this should be done in a way that is as neutral as possible, avoiding selecting datasets and metrics that would unfairly favor the new method \cite{Boulesteix2013-vy, Weber2019-el}.
-A diverse, easily accessible, collection of benchmarking datasets and metrics for a given task would be one way of addressing this, and additionally has the benefit of lowering the barrier of entry for method developers into a field of application.
+For \emph{method developers}, benchmarking compares their method to the state of the art, ideally using neutral datasets and metrics to avoid intrinsic bias \cite{Boulesteix2013-vy, Weber2019-el}.
+A diverse, accessible set of benchmarking datasets and metrics can address this and lower the entry barrier for method developers.
 Moreover, it is important that the new method is compared to the current state of competing methods, to best mimic the choices available to a potential method user.
-One could, however, imagine other use cases, where the new method would be systematically compared to previous versions of itself or alternative implementations of the `same' method \cite{Rich2024-jh}, to illustrate stability of code changes or improvements made over time.
+Alternatively, new methods could be compared to previous versions or different implementations of the `same' method \cite{Rich2024-jh}, to illustrate stability of code changes or improvements made over time.
 In both cases, the typical approach today would be for the method developer to rerun all the methods and versions to which they want to compare their current method, on all the datasets they wish to use.
@@
 
-Benchmarks can be seen as a collection of data files and (perhaps diverse) source code ideally run as a workflow in some computing environment.
+Benchmarks are collections of data and source code, ideally executed as workflows within a computing environment.
 Current practices in single-cell RNA-seq benchmarking show that code is frequently open, reusable and versioned; input datasets are also typically available \cite{Sonrel2023-te}.
@@
 
-Over 350 workflow languages, platforms or systems  \cite{Wratten2021-el, Amstutz2024-qk} and one standard named Common Workflow Language (CWL) (with multiple implementations) exist \cite{Amstutz2016-vo}.
+Over 350 workflow languages, platforms or systems  \cite{Wratten2021-el, Amstutz2024-qk} and one standard (Common Workflow Language, CWL) exist \cite{Amstutz2016-vo}.
 Some workflow languages can export workflow definitions to CWL, hence helping computational FAIR guidelines: that is, bundling and processing data in association to their metadata, tracking new metadata generation, recording provenance, and being accessible \cite{Goble2020-ps}.
@@
 
-Parameters can be included into a benchmark to test method performance across a range of settings, allowing the impact of parameter choices to be assessed  or to find the optimal parameter settings.
+Parameters can be included in a benchmark to test method performance across a range of settings, allowing the impact of parameter choices to be assessed  or to find the optimal parameter settings.
 Notably, when assessing the impact of a certain parameter, the research objective of the benchmark differs from a classic method comparison: methods that do not include the parameter to be explored cannot be evaluated for this question.
@@
 These terms can take different meanings depending on the context, but here the benchmark infrastructure location is considered.
-In most systems, the code to execute will be tracked in a reference \texttt{git} repository (e.g., on GitHub, GitLab).
-The workflow execution and the storage of derived datasets, and various stages of results are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery platforms (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., could be a laptop, server or cloud; e.g.
+In most systems, code will be tracked in a reference \texttt{git} repository (e.g., on GitHub, GitLab).
+The workflow execution and the storage of derived files are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., laptop, server or cloud; e.g.
 omnibenchmark \cite{omnibenchmark}).
@@
 
-Arguably, most benchmarks could start out as decentralized, in part as a means to properly define the concept and problem, but also as a way to generate preliminary results and ideas to solidify the design before `going public' (centralization).
-Thus, users need a system that can build benchmark components in both a centralized and decentralized fashion, i.e., starting decentralized and promoting a benchmark to centralized when it becomes sufficiently mature.
-Accordingly, the process of centralization is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to computational resources) and adding measures to build trust \textit{via} transparency.
+Most benchmarks can begin decentralized, allowing for concept definition and preliminary results before `going public' (centralizing phase).
+Users need a system that supports building benchmarks both ways, transitioning to centralization as the design matures.
+However, becoming centralized is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to computational resources) and adding measures to build trust \textit{via} transparency.
 
@@
 An active community is vital to the success of a benchmarking system.
-As mentioned, different roles and ways that people can contribute to or interact with a benchmarking system have been defined.
+As mentioned, different roles and people can contribute to or interact with a benchmarking system.
 The benchmarkers (which may include multiple scientists) are responsible for planning and coordinating a benchmark, defining the task, possibly splitting it into subtasks or processing stages, and defining the data formats across the stages; the benchmarker also brings an authority role of how to review and approve contributions.
-The contributors (which may include method developers) curate and add content to the benchmark, which could be new datasets, analysis methods or evaluation metrics, adhering to the guidelines set up by the benchmarker.
+The contributors (which may include method developers) curate and add content to the benchmark, which could be new datasets, methods or metrics, adhering to the guidelines set up by the benchmarker.
 Finally, the viewers (which may include benchmarkers and contributors) of benchmark results are users who retrieve one or more of the artifacts (dataset files, intermediate results, or metric scores).
-This could span a range of use cases, including the data analyst choosing which method to use for a specific application, an instructor retrieving a curated dataset for teaching purposes, or a methods researcher prototyping their new method.
+These could span a range of use cases, including the data analyst choosing which method to use for a specific application, an instructor retrieving a curated dataset for teaching purposes, or a methods researcher prototyping their new method.
 
-Given the importance of such a community around a benchmarking system, the question arises on how it can best be assembled and sustained.
-Some strategies include organizing events, including seminars or workshops (typically on the educational side) as well as hackathons and challenges (more aimed at practical onboarding or solicitation of contributions).
+As for assembling and sustaining communities, some strategies include organizing events, including seminars or workshops (typically on the educational side) as well as hackathons and challenges (more aimed at practical onboarding or solicitation of contributions).
 Career-related or financial incentives, such as the prospect of a scientific publication, or compensation for the allocated time, are also common motivators.
 For a method developer, contributing to a public benchmark provides a way to test their method without the need to set up all the infrastructure themselves, and also provides an opportunity to advertise their work.
-Ultimately, the likelihood that someone will contribute to a benchmark will likely be strongly driven by whether the results are of interest to them and their research, as well as how easy it is to contribute.
-Community engagement and contributions can further be made easier by providing clear guidelines (and organization), a suggestion of `good first issues' to tackle, a clear set of contributor guidelines and a code of conduct that covers all interactions related to the project.
+Ultimately, the likelihood that someone will contribute to a benchmark will be associated with whether the results are of interest to them, as well as how easy it is to contribute.
+Contributions can be made easier by providing good organization, suggested `good first issues' to tackle, contributor guidelines and a code of conduct.
 
-The public solicitation of contributions, however, is not without difficulties, and it is important to organize the setup transparently to ensure trust from all parties.
-A benchmarker should be able to trace and test contributions, to detect misbehaviour and suspicious activities, but also to avoid that non-functional or suboptimal contributions enter the benchmark.
-This can be addressed by a `quarantine zone', where new content is automatically tested before being added to the system (i.e., executed in a sandbox environment).
+Publicly soliciting contributions poses challenges, requiring a transparent setup to build trust.
+Benchmarkers must trace and test inputs to detect misuse and prevent non-functional or suboptimal entries.
+This can be addressed by a `quarantine zone', where new content is automatically tested before being added to the system.
 Similarly, a contributor should be able to trace their contribution, to ensure that no unintended modifications are made after submission.
-Ultimately, challenges arise when anyone can add a new method.
-Benchmarkers are typically not equally familiar with all included methods, and one would also need to be aware of potential malicious intent (e.g., submitting a suboptimal application of a competing method).
-On the other hand, many risks can be mitigated by the inherent transparency \cite{Greenstein2016-os}, and feedback loops with (all) original developers could be established to ensure optimal method usage.
+Ultimately, many risks can be mitigated by the inherent transparency \cite{Greenstein2016-os}, and feedback loops with (all) original developers.
 
-A related challenge concerns which datasets and metrics to include in a benchmark.
-On the one hand, diverse sets of datasets and metrics are important to ensure generalizability and avoid overfitting.
-On the other hand, inclusion of low-quality or redundant datasets or metrics does not benefit the benchmark, and may provide misleading (aggregate) results.
-Possible mitigations may include sensitivity analyses, evaluating the impact of single datasets or metrics on the overall, aggregated performance of a method, or reducing redundancy by down-weighting similar metrics to avoid a specific aspect dominating the evaluation.
+A key challenge is selecting datasets and metrics for a benchmark.
+Diversity ensures generalizability and prevents overfitting, but low-quality or redundant data can lead to misleading results.
+Possible mitigations include sensitivity analyses, evaluating the impact of single datasets or metrics on the overall performance metrics, or reducing redundancy by down-weighting similar metrics to avoid a specific aspect dominating the evaluation.
 From a wider perspective, these questions are all related to the governance of a benchmark, a topic that has not received much attention.
-Ultimately benchmarks may want to, depending on their scale and complexity, implement different governance models.
-In one model, the benchmarkers have the ultimate authority to decide on what gets included in the benchmark, simplifying the decision making but creating a risk of gatekeeping.
-In another model, decisions are made collectively by a group of contributors, which may mitigate the risk of gatekeeping at least partially, but creates a more complex decision structure and introduces the need for a conflict resolution strategy.
+As for governance models, benchmarkers could have the ultimate authority to decide on what gets included in the benchmark, simplifying the decision making but creating a risk of gatekeeping; or decisions could be made collectively by a group of contributors, which may mitigate the risk of gatekeeping at least partially, but creates a more complex decision structure and introduces the need for a conflict resolution strategy.
 The governance model of a benchmark could also cover aspects related to computing infrastructure and storage, as well as guidelines to manage authorship of any resulting publications or resources.
 Having the governance model spelled out when launching a benchmark is likely to reduce friction at later stages of the project.
-In all cases, an advantage of doing things in public is that contributions are logged.
 
 Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods).
-Although highly desirable, neutrality is nonetheless extremely tough to assess given only the downstream artifacts of a benchmark; this is an area of governance that deserves further attention to increase trust.
 One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a method.
@@
 Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then hard to establish (e.g., a perceived conflict of interest to evaluate one's own method) or where a deeper knowledge of their own tools might favor the implementation of the latter compared to other competitors.
-Here, some recent work to mitigate this conflict involves pre-registration, which typically involves specifying the parameters of your study in advance and posting them on a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial formalization of a benchmark could semi-automatically provide the basis for the formal preregistration.
+Here, some recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance on a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial formalization of a benchmark could semi-automatically provide the basis for the formal preregistration.
 In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, because the goal is not simply to advertise the virtues of a single new method; nonetheless, the challenge to assess neutrality directly remains difficult.
@@
 Templates exist for simulation studies that could be adapted more generally for benchmarking projects \cite{Siepe2023-tz}.
-How to reconcile continuity and pre-registration of benchmarks needs further attention.
 
-Finally, it is worth noting that there is a tension between preregistration and the concept of `continuous' benchmarking, where new datasets, methods and new metrics are added over time.
+Finally, there is a tension between preregistration and the concept of `continuous' benchmarking, where new datasets, methods and new metrics are added over time.
 The logic of continuous benchmarking is that as the problem becomes more understood in the field, it becomes increasingly clear what are good ways to evaluate the performance of methods.
@@
 In practice, however, it is sometimes easier to develop new code than to reuse existing software \cite{Trisovic2022-ol}.
-This leads to the phenomenon of `academic abandonware', where projects are forgotten in code repositories, exacerbating the reproducibility crisis in research; one of the noted challenges in benchmarking was the lack of extensible frameworks \cite{Sonrel2023-te}.
-A common reason for abandoned projects is their failure to adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) principles \cite{Wilkinson2016-bh}.
-Another common cause is the life cycle inherent to academic practice: after delivering a research output, there are often no resources to maintain a software package in the longer term; left alone, the likelihood of a particular library compiling or being able to run after 5-10 years will drastically decrease.
+This leads to the phenomenon of `academic abandonware', where projects are forgotten in code repositories, exacerbating the reproducibility crisis in research; one of the noted challenges in benchmarking was the lack of extensibility \cite{Sonrel2023-te}.
+A common cause is the life cycle inherent to academic practice: after delivering a research output, there are often no resources to maintain a software package in the longer term.
 
 Operational costs are given by the expected compute and storage requirements.
-In the context of benchmarking, predicting CPU and memory usage is challenging.
-The methods under evaluation typically have variable computational demands.
+In the context of benchmarking, methods under evaluation typically have variable computational demands which are hard to predict.
 Benchmark designers could impose resource limits, thereby evaluating the methods under different constraints.
@@
 
-Software reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}.
-Computationally, a benchmark executes code that transforms input data into outputs: this execution takes place and is affected by the execution environment; the environment encompasses the base system (OS, compiler toolchains, libraries) and a set of configurations.
-The benchmark definition should control as much as possible of the execution environment.
+Computational reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}.
+During benchmarking, a workflow runs code transforming inputs into outputs: this execution is affected by the base system (OS, compiler toolchains, libraries) and its configuration.
+The benchmark definition should control the execution environment.
 
@@
 \begin{enumerate}
-\item The benchmarking system itself has an impact on operational costs, such as storage needs, execution platform, etc.
-For example, it imposes a certain software stack and base dependencies, and orchestrates the execution of a benchmark plan; usually, the system mandates the choice of a particular workflow manager.
-At this level, any requirements for input/output formats and shapes are defined; different systems can impose different constraints.
+\item The benchmarking system has an impact on operational costs, such as storage needs, execution platform, etc.
+The system mandates the choice of a particular stack and workflow manager.
+At this level, any requirements for input/output formats and shapes are defined.
 
-\item Contributions related to individual benchmark components (e.g., datasets, methods, metrics) are expected to be small, typically short scripts that process data and wrap functionality by importing external libraries (where the component implementation itself is developed).
-Even so, imposing good practices at this level (output validation, testing for abnormal termination, ability to run with a subset of input data) can be beneficial to increase the quality and maintainability of the contributions.
-The toolset can also enforce the validation of metadata annotations (authorship, versioning, licences) for each contributed module.
+\item Benchmark components, like dataset-generating simulations, methods, or metrics, are typically small scripts that process data and use external libraries.
+Even if method development happens elsewhere, good practices should be embraced (output validation, abnormal termination checks, running on subsets of input data).
+Metadata validation should be enforced for each module, including authorship, versioning, and licenses.
 
-\item The software dependencies have easily the biggest impact on replicability and maintainability; archiving all combinations of a large dependency tree, across an arbitrary number of base OS images, will exponentially increase retention and maintenance costs.
-Turning the long-term replicability of a benchmark into a tractable problem is linked to choosing a suitable software management system.
-Several software management systems have emerged to address these problems, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.
+\item Software dependencies cause the biggest impact on replicability and maintainability; archiving all combinations of a large dependency tree, across several OS images, exponentially increases retention and maintenance costs.
+Making the long-term replicability of a benchmark tractable implies adopting a sane software management system.
+Several solutions exist, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.
 
@@
 
-Even the simplest benchmark designed and run by a single person aiming to compare their method against various `gold standard' methods incorporates the work of many (dataset generators, method developers, operating systems with system dependencies and their developers).
+Even a simple, single-person benchmark comparing a method to gold standards relies on multiple contributions, from dataset creators and method developers to system software authors.
 Benchmark artifacts are generated by running a workflow \cite{Jackson2021-ke}.
 Hence, data, code and operating system dependencies are all accessible during the benchmarking process.
-For reproducibility and transparency and to build trust \cite{Laine2007-py}, this freedom should be extended for others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how another person is implementing my own method), change (bug fix, contribute) and improve (extend) benchmarks.
+For reproducibility, transparency and to build trust \cite{Laine2007-py}, this accessibility and freedom should be extended for others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how another person is implementing my own method), change (bug fix, contribute) and improve (extend) benchmarks.
 Such freedoms are granted (or restricted) by software and content licenses, as are copyright and authorship and attribution \cite{Kreutzer2014-ua}.
 Without explicit licenses, even publicly available code is copyrighted by default, and reuse is restricted (to reproduce, distribute, or create derivative works) \cite{Kreutzer2014-ua}.
-
-In practical terms, a benchmarking system needs to be aware of software and data licenses, track further changes and attribute them to the (academic) benchmark contributor.
 A concise summary of license rights and restrictions, and compatibility across commonly-used open-source licenses can be found in Table~\ref{tab:licenses}.
-Benchmark contributions should bring a clear FOSS (free and open source software) license, perhaps by assigning the copyright holder to contributors describing their name, email and ORCID (as the canonical scholarly identifier), hence allowing seamless reuse, transparency and easing attribution.
+Benchmark contributions should bundle licenses to scholarly names including email and ORCID (as the canonical identifier), hence allowing seamless reuse, transparency and attribution.
 
@@
 Similarly, even if all benchmark code were free software, running it could be hindered by (non-free) proprietary interpreters (e.g., MATLAB).
-Benchmarking systems should be explicit in allowing or disallowing proprietary software requirements, and be aware and track EULAs (end user license agreements).
+Benchmarking systems should be explicit in allowing or disallowing proprietary software, and be aware and track EULAs (end user license agreements).
 This conundrum extends to the operating system layer and to system dependencies (e.g., compilers).
-In terms of benchmark re-runnability, fully FOSS system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{Droge2023-ax}.
+In terms of benchmark re-runnability, fully free software system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{Droge2023-ax}.
 
-Finally, data (inputs and outputs) are frequently stored in centralized data repositories \cite{Potter2015-fs, Sicilia2017-ko, van-de-Sandt2019-wd} under content licenses, which might enforce attribution or copyleft.
+Finally, data files (inputs and outputs) are frequently stored in centralized repositories \cite{Potter2015-fs, Sicilia2017-ko, van-de-Sandt2019-wd} under content licenses, which might enforce attribution or copyleft.
 A benchmarking system should automate license metadata handling during data import and export, crediting the original authors if applicable; it also represents an opportunity to enforce that academics use appropriate licenses for data and code.
-
-Scientific benchmarking is rarely a single-person endeavour and has to deal with diverse content and contributions.
-Software and content licenses standardize the (legally binding) usage rights and restrictions.
-Benchmarking systems can leverage licenses and version control to track, store and monitor the reuse of data and code in an automated way.
-On the other hand, contribution attribution in the academic sense (e.g., authorship) is as important and beyond automation.
-Collaborative benchmarking initiatives should explicitly list guidelines on using and distributing code and results (i.e., licenses) as well as clearly communicate the authorship strategies.
-These guidelines, complemented with a code of conduct, could be bundled into the benchmark description.
 
@@
 Constraints can also enable validation, which can greatly help with the development process and quality assurance.
-    \item  Perfect software reproducibility is possible \cite{Lamb2021-tb}, but it comes with a high engineering cost.
-It is thus reasonable for a benchmarker to state the desired degree of replicability and the corresponding constraints; different tasks might demand different levels of compliance.
-    \item  Security concerns: decentralized runs lower the participation barrier, but it also increases the potential for attacks on their computing environment.
+    \item  Achieving software reproducibility needs discipline and commitment \cite{Lamb2021-tb}.
+A benchmarker can state the desired degree of replicability and constraints; different tasks require different levels of compliance.
+    \item  Security concerns: decentralized runs lower the participation barrier, but also increase the attack  surface.
 Sandboxed environments (restricting network access, filesystem access etc) can be provided as mitigation.
@@
 Benchmarking involves applying methods to data.
-The types of data can be diverse and even a single dataset can be stored in multiple formats.
+Data types can be diverse and even a single dataset can be stored in multiple formats.
 A broad distinction can be made according to general vs.
@@
 Likewise, proprietary formats should be avoided, since they may change and could require specific (proprietary) software to read.
-General language-agnostic formats often do not change over time and are thus more interoperable such that these datasets can be made FAIR \cite{Wilkinson2016-bh}.
-Generally, using established standards is recommended (e.g., SAM, BED, FASTQ, etc for genomics), since related methods are developed around these (typically storage-efficient) formats.
+General language-agnostic formats often do not change over time and are thus align to FAIR principles \cite{Wilkinson2016-bh}.
+Generally, using established standards is recommended (e.g., SAM, BED, FASTQ, etc for genomics), since related methods are developed around these formats.
 
 If raw input data consists of multiple file formats, an option could be to convert the data into a common format with the caveat that this `intermediate' raw data could be either stored (increased storage usage, easier benchmark definition) or created `on the fly' (increased compute).
-Furthermore, how raw input data is retrieved (e.g., automatic retrieval once and store locally, automatic retrieval every time, or manual retrieval and store locally) needs to be decided.
+Furthermore, how raw input data are retrieved (e.g., automatic retrieval once and store locally, automatic retrieval every time, or manual retrieval and store locally) needs to be decided.
+These intermediate results could also be made FAIR.
 
-Additionally, which level and scope of data validation needs to be decided.
-Validations include content integrity (e.g., md5sum), adherence to a data specification (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data).
-Intermediate data and results should be made FAIR, leading to additional decisions, including the scope of data sharing (e.g., only the most important subset), the handling of the metadata (automatic versus manual annotation) and the data storage provider and policies (where and how to make available, e.g., GEO or Zenodo).
+Additionally, validation are recommended. Validations include content integrity (e.g., md5sum), adherence to a data specification (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data).
+
+Finally, metadata (automatic versus manual annotation) should not be neglected.
 Automatic generation of metadata might require predefined standards and an ontology to fully track how the data was generated \cite{LeboUnknown-om}.

### 2024-10-23 manifesto-main.tex
Updates from Overleaf

@@
 \usepackage[utf8]{inputenc}
-\usepackage[right]{lineno}
+%\usepackage[right]{lineno}
 \usepackage{csquotes}
@@
 
-\renewcommand{\thetable}{\Roman{table}}
-
-% Define a new format for \subsection
-\titleformat{\subsection}
-  {\mdseries\itshape\large} % Medium series, italic shape, and large font size
-  {\thesubsection}{1em}{} % Numbering, spacing, and the section title itself
+%\renewcommand{\thetable}{\Roman{table}}
 
@@
 
-\section*{What is a benchmarking system?}
+\section*{Why benchmarking systems?}
 %\label{sec:main_section}
@@
 Can benchmarking be systematized?
-Benchmarks should be orchestrated and run by `benchmarking system's to create `artifacts' (code snapshots, file outputs to be shared, performance outputs).
+Benchmarks should be orchestrated and run by \emph{benchmarking systems} to create `artifacts' (code snapshots, file outputs to be shared, performance outputs).
 The various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
@@
 \pagebreak
-\linespread{1.1} 
+\linespread{1} 
 %\printbibliography

### 2024-10-29 manifesto-main.tex
Add Mark's edits

@@
 Benchmarking (further defined below) is a critical step in method development: either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, as defined in \cite{Cao2023-jz}.
-Notably, a benchmarking study can be run by a single person or a small group on a laptop (or a server), or it can be organized as a `challenge' or within the duration of a hackathon.
+A benchmarking study can be run by a single person or a small group, or it can be organized as a `challenge' or within a hackathon.
 
@@
 
-\section*{Why benchmarking systems?}
+\section*{Why a benchmarking system?}
 %\label{sec:main_section}
@@
 A benchmark thus requires a well-defined task and a definition of correctness, or ground-truth, to be precisely defined in advance.
-Then, a benchmark consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, metrics.
+A benchmark then consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, and metrics.
 
@@
 The various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
-Such a system should facilitate components to be public, open and allow contributions (e.g.
-corrections or new components), such that the benchmarks of the future can run more continuously.
+Such a system should facilitate components to be public, open and allow contributions (e.g., corrections or new components), such that the benchmarks of the future can run more continuously.
 At the outset of a benchmark, a benchmark `definition' could be envisioned, which could give a formal specification of the entire set of components (and the pattern of artifacts to be generated).
-The benchmark definition could even be expressed as a single configuration file that specifies the scope and topology of components to include, details of the repositories (code implementations with versions), the instructions to create software environments (accessible across compute architectures), and the parameters used.
+The benchmark definition could even be expressed as a single configuration file that specifies the scope and topology of components to include, details of the repositories (code implementations with versions), the instructions to create software environments (accessible across compute architectures), the parameters used, and what components to snapshot for a release.
 A benchmark definition in terms of layers and their corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.
@@
 
-\section*{Benchmarks stakeholders}
+\section*{Benchmark stakeholders}
 
@@
 
-For \emph{method developers}, benchmarking compares their method to the state of the art, ideally using neutral datasets and metrics to avoid intrinsic bias \cite{Boulesteix2013-vy, Weber2019-el}.
+For \emph{method developers}, benchmarking allows their method to be compared to the state of the art, ideally using neutral datasets and metrics to avoid intrinsic bias \cite{Boulesteix2013-vy, Weber2019-el}.
 A diverse, accessible set of benchmarking datasets and metrics can address this and lower the entry barrier for method developers.
 Moreover, it is important that the new method is compared to the current state of competing methods, to best mimic the choices available to a potential method user.
-Alternatively, new methods could be compared to previous versions or different implementations of the `same' method \cite{Rich2024-jh}, to illustrate stability of code changes or improvements made over time.
+Alternatively, new methods could be compared to previous versions or even different implementations of the `same' method \cite{Rich2024-jh}, to illustrate stability of code changes or improvements made over time.
 In both cases, the typical approach today would be for the method developer to rerun all the methods and versions to which they want to compare their current method, on all the datasets they wish to use.
@@
 
-Benchmarks can also be of high value to \emph{scientific journals and funding agencies}.
+Benchmarks can also be of high value to \emph{scientific journals} and \emph{funding agencies}.
 Well-executed benchmark studies can be highly utilized, influential and cited, and may point to gaps in the current body of methods, guiding future methodological developments; at the same time, in fast-moving fields like bioinformatics, benchmarks have the tendency to rapidly become stale.
-Journals and funding bodies also have a vested interest in ensuring that published or funded method evaluations are performed to a high standard, that unnecessary redundancy is reduced, and that results are FAIR (findable, accessible, interoperable and reusable) for maximal benefit to the community \cite{Wilkinson2016-bh}.
-Thus, quality assurance, neutrality and transparency of benchmarking systems in themselves become highly relevant.
+Journals and funding bodies also have a vested interest in ensuring that published or funded method developments are performed to a high standard, that unnecessary redundancy is reduced, and that results are FAIR (findable, accessible, interoperable and reusable) for maximal benefit to the community \cite{Wilkinson2016-bh}.
+Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
 
-Finally, benchmarks could form a research topic on their own, either for newcomers to the field to get acquainted with the conceptual options that are available, or for experts to consolidate research efforts of a subfield.
+Finally, benchmarks could form a research topic on their own, either for newcomers to the field to get acquainted with the conceptual options available, or for experts to consolidate research efforts of a subfield.
 For this, it is important that there are suitable venues for publishing benchmarks.
@@
 
-\section*{Benchmarks could be formally defined}
+\section*{Benchmarks can be formally defined}
 
 Benchmarks are collections of data and source code, ideally executed as workflows within a computing environment.
-Current practices in single-cell RNA-seq benchmarking show that code is frequently open, reusable and versioned; input datasets are also typically available \cite{Sonrel2023-te}.
-However, the \textit{extensibility} of benchmarking studies is generally very low, as well as the proportion of benchmarks using a formal workflow system.
+Indeed, recent studies highlight that code is frequently open, reusable and versioned, as are input datasets \cite{Sonrel2023-te, cao2023-jz}.
+However, \textit{extensibility} of benchmarking studies is generally very low, as is the proportion of benchmarks using a formal workflow system.
 Thus, a formal system to orchestrate benchmarking workflows would largely impact the current landscape.
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and efficient software stack; managing storage and their access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and efficient software stack; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results.
 Benchmarking systems are in charge of orchestrating both workflow execution and these benchmarking-specific tasks.
@@
 \texttt{OpenProblems} uses \texttt{Viash} and \texttt{nextflow} to handle software environments and workflows, respectively, and reports updated leaderboards with custom code \cite{Luecken2024-fk}.
-To our knowledge, no benchmarking platform makes use of a fully expressive benchmark definition formalization language covering both the workflow generation and the benchmarking-specific tasks.
+To our knowledge, no benchmarking platform makes use of a fully expressive benchmark definition formalization language covering workflow generation and benchmarking-specific tasks.
 Such a language would go a long way toward facilitating benchmark conceptualization, community gathering and result sharing.
@@
 
-Benchmarks are often focused on finding the `top' performing method, or on generating ranked lists of methods.
+Benchmarks are often focused on finding the `top' performing method(s), or on generating rankings.
 While this is valid in a few fields with clearly defined goals, such as image recognition \cite{Khan2018-ot} or protein structure prediction \cite{Jumper2021-oq}, bioinformatics tasks are typically evaluated by multiple metrics, some of which are not well understood \cite{Lutge2021-mt, Reinke2024-dw}.
@@
 Besides interactive dashboards for more nuanced result exploration \cite{bettr}, multi-criteria decision analysis (MCDA) \cite{Taherdoost2023-wd} has been proposed to guide users through benchmark results and multidimensional scaling (MDS) has been used to differentiate the effect of single datasets or metrics and recognize patterns beyond aggregated performance \cite{Niessl2022-fk}.
-An extensive analysis and presentation of the results not only helps the users in the long term but also the whole scientific field to identify pain points, the complexity of possible scenarios and to isolate the different use-cases.
+An extensive analysis and presentation of the results not only helps users in the long term but also the whole scientific field to identify limitation across various use cases.
 The first advantage of such a strategy is to force the user to an attentive reading of the performance results so that the correct use-case and relevant methods can be identified.
-The second, more important, advantage to present benchmark results extensively is to identify the challenges in a field of analysis, i.e., the scenarios, types of datasets, or aspects of performance that are systematically associated with poor results for all current methods.
+The second, more important, advantage to present benchmark results extensively is to identify the nuances, e.g., the scenarios, types of datasets, or aspects of performance that are systematically associated with poor results.
 Being public and open to contributions from specialists of different fields, a benchmarking system could gather more reference datasets and methods, and allow the current state of the field to be better grasped.
@@
 
-Parameters can be included in a benchmark to test method performance across a range of settings, allowing the impact of parameter choices to be assessed  or to find the optimal parameter settings.
-Notably, when assessing the impact of a certain parameter, the research objective of the benchmark differs from a classic method comparison: methods that do not include the parameter to be explored cannot be evaluated for this question.
+The impact of parameter choices can be included in a benchmark to test performance across a wider range of settings, or to find the optimal parameter settings.
 Parameters can be method-specific, but also dataset-specific, e.g., the true number of clusters when evaluating clustering algorithms.
@@
 In designing a benchmarking system, it may be helpful to specify whether the benchmark is intended to be run `centralized' or `decentralized'.
-These terms can take different meanings depending on the context, but here the benchmark infrastructure location is considered.
-In most systems, code will be tracked in a reference \texttt{git} repository (e.g., on GitHub, GitLab).
-The workflow execution and the storage of derived files are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., laptop, server or cloud; e.g.
-omnibenchmark \cite{omnibenchmark}).
+These terms can take different meanings depending on the context, but here the benchmark infrastructure location is intended.
+In most systems, code will be tracked in a \texttt{git} repository (e.g., on GitHub, GitLab).
+The workflow execution and the storage of derived files are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery (e.g., GitHub actions for \texttt{ncbench}) or on arbitrary infrastructure (i.e., laptop, server or cloud; e.g., omnibenchmark \cite{omnibenchmark}).
 
@@
 Working together would mean sharing not only datasets and code snippets, but also execution environments (software but perhaps also hardware), alignment on parameters to evaluate and eventually sharing, consolidating, and interpreting results.
-However, because there are no standards of how benchmarks are implemented \cite{Sonrel2023-te}, another large disadvantage of decentralized benchmarks is that, while these components may eventually be openly shared as a product of a research study, the contents of a benchmark are non-interoperable, even across benchmarks of the same task.
+However, because there are no standards of how benchmarks are implemented \cite{Sonrel2023-te}, another disadvantage of decentralized benchmarks is that, while these components may eventually be openly shared as a product of a research study, the contents of a benchmark are non-interoperable, even across benchmarks of the same task.
 Furthermore, decentralized benchmarks have been criticized on the basis of `inflating model performance' due to decisions made in preprocessing or parameter tuning \cite{Luecken2024-fk}; such issues could be mitigated by gathering the `wisdom of crowds' that is implicit in running centralized benchmarks.
 
-Most benchmarks can begin decentralized, allowing for concept definition and preliminary results before `going public' (centralizing phase).
+Most benchmarks could begin decentralized, allowing for concept definition and preliminary results before `going public' (centralization phase).
 Users need a system that supports building benchmarks both ways, transitioning to centralization as the design matures.
-However, becoming centralized is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to computational resources) and adding measures to build trust \textit{via} transparency.
+However, becoming centralized is not cost-free: it implies gatekeeping (contributors authenticate to the system and get access to computational resources) and adding measures to build trust \textit{via} transparency.
 
@@
 As mentioned, different roles and people can contribute to or interact with a benchmarking system.
-The benchmarkers (which may include multiple scientists) are responsible for planning and coordinating a benchmark, defining the task, possibly splitting it into subtasks or processing stages, and defining the data formats across the stages; the benchmarker also brings an authority role of how to review and approve contributions.
-The contributors (which may include method developers) curate and add content to the benchmark, which could be new datasets, methods or metrics, adhering to the guidelines set up by the benchmarker.
-Finally, the viewers (which may include benchmarkers and contributors) of benchmark results are users who retrieve one or more of the artifacts (dataset files, intermediate results, or metric scores).
+Benchmarkers (which may include multiple scientists) are responsible for planning and coordinating a benchmark, defining the task, possibly splitting it into subtasks or processing stages, and defining the data formats across the stages; the benchmarker also brings an authority role of how to review and approve contributions.
+Contributors (which may include method developers) curate and add content to the benchmark, which could be new datasets, methods or metrics, adhering to the guidelines set up by the benchmarker.
+Finally, the viewers (which may include benchmarkers and contributors) of benchmark results are users who retrieve one or more of the artifacts (code snippets, dataset files, intermediate results, or metric scores).
 These could span a range of use cases, including the data analyst choosing which method to use for a specific application, an instructor retrieving a curated dataset for teaching purposes, or a methods researcher prototyping their new method.
 
-As for assembling and sustaining communities, some strategies include organizing events, including seminars or workshops (typically on the educational side) as well as hackathons and challenges (more aimed at practical onboarding or solicitation of contributions).
+For assembling and sustaining communities, strategies include organizing seminars or workshops (on the educational side) or  hackathons and challenges (aimed at practical onboarding or contributions).
 Career-related or financial incentives, such as the prospect of a scientific publication, or compensation for the allocated time, are also common motivators.
@@
 Publicly soliciting contributions poses challenges, requiring a transparent setup to build trust.
-Benchmarkers must trace and test inputs to detect misuse and prevent non-functional or suboptimal entries.
+Benchmarkers must trace and test contributions to detect misuse and prevent non-functional or suboptimal implementations.
 This can be addressed by a `quarantine zone', where new content is automatically tested before being added to the system.
 Similarly, a contributor should be able to trace their contribution, to ensure that no unintended modifications are made after submission.
-Ultimately, many risks can be mitigated by the inherent transparency \cite{Greenstein2016-os}, and feedback loops with (all) original developers.
+Ultimately, many risks can be mitigated by transparency \cite{Greenstein2016-os}, and feedback loops with (all) developers.
 
 A key challenge is selecting datasets and metrics for a benchmark.
-Diversity ensures generalizability and prevents overfitting, but low-quality or redundant data can lead to misleading results.
-Possible mitigations include sensitivity analyses, evaluating the impact of single datasets or metrics on the overall performance metrics, or reducing redundancy by down-weighting similar metrics to avoid a specific aspect dominating the evaluation.
+Diversity of these ensures generalizability and prevents overfitting, but low-quality or redundant datasets can lead to misleading results.
+Possible mitigations include sensitivity analyses, evaluating the impact of single datasets or metrics on overall performance metrics, or reducing redundancy by down-weighting similar metrics to avoid a specific aspect dominating the evaluation.
 From a wider perspective, these questions are all related to the governance of a benchmark, a topic that has not received much attention.
-As for governance models, benchmarkers could have the ultimate authority to decide on what gets included in the benchmark, simplifying the decision making but creating a risk of gatekeeping; or decisions could be made collectively by a group of contributors, which may mitigate the risk of gatekeeping at least partially, but creates a more complex decision structure and introduces the need for a conflict resolution strategy.
+For governance, benchmarkers could choose to have the ultimate authority to decide on what gets included in a benchmark, simplifying the decision making but creating a risk of gatekeeping; or, decisions could be made collectively by a council of contributors, which may mitigate the risk of gatekeeping at least partially, but creates a more complex decision structure and introduces the need for a conflict resolution strategy.
 The governance model of a benchmark could also cover aspects related to computing infrastructure and storage, as well as guidelines to manage authorship of any resulting publications or resources.
-Having the governance model spelled out when launching a benchmark is likely to reduce friction at later stages of the project.
+Having the governance model spelled out early in a project is likely to reduce friction later.
 
 Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods).
-One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a method.
+One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a new method.
 Similarly, ground truth generation and simulations in MDP benchmarks can unconsciously mirror the method capabilities, mechanisms, or strengths, thus biasing evaluation.
-Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then hard to establish (e.g., a perceived conflict of interest to evaluate one's own method) or where a deeper knowledge of their own tools might favor the implementation of the latter compared to other competitors.
-Here, some recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance on a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial formalization of a benchmark could semi-automatically provide the basis for the formal preregistration.
-In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, because the goal is not simply to advertise the virtues of a single new method; nonetheless, the challenge to assess neutrality directly remains difficult.
-Other benchmarking models are possible, from challenges to hackathons and here, there is hope that a consensus of the masses can help to establish neutrality.
+Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then harder to establish.
+Here,  recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance at a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial benchmark formalization could even provide the basis for the formal preregistration.
+In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, but the challenge to assess neutrality directly remains difficult.
+Other benchmarking models are possible, from challenges to hackathons, with the hope that a consensus of the masses can further help to establish neutrality.
 One potential strategy, in the context of the systematic design of benchmarks, is establishing a system where a benchmark definition is parsed into a pre-registration template.
@@
 Operational costs are given by the expected compute and storage requirements.
-In the context of benchmarking, methods under evaluation typically have variable computational demands which are hard to predict.
-Benchmark designers could impose resource limits, thereby evaluating the methods under different constraints.
+In the context of benchmarking, methods under evaluation typically have unpredictable computational demands.
+Benchmark designers could impose resource limits, thereby evaluating the methods under constraints.
 Storage costs, on the other hand, are more predictable and can be directly influenced by design decisions of the benchmark (e.g., low retention for re-generatable artifacts, `cold storage' for code archives and software, external data stores such as Zenodo for method performance artifacts, CernVM-FS for reusable software).
@@
 
-A good benchmarking system needs to package its components in a way that ensures that their execution and artifacts can be consistently replicated within practical limits.
+A good benchmarking system needs to package its components in a way that ensures that their execution and artifacts can be consistently replicated (within practical limits).
 
@@
 
-Even a simple, single-person benchmark comparing a method to gold standards relies on multiple contributions, from dataset creators and method developers to system software authors.
-Benchmark artifacts are generated by running a workflow \cite{Jackson2021-ke}.
+Even a simple, single-person benchmark comparing methods relies on multiple contributions, from dataset creators and method developers to system software and workflow authors.
 Hence, data, code and operating system dependencies are all accessible during the benchmarking process.
-For reproducibility, transparency and to build trust \cite{Laine2007-py}, this accessibility and freedom should be extended for others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how another person is implementing my own method), change (bug fix, contribute) and improve (extend) benchmarks.
+For reproducibility, transparency and to build trust \cite{Laine2007-py}, this accessibility and freedom should be extended for others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how someone is implementing a method), change (bug fix, contribute) and improve (extend) benchmarks.
 Such freedoms are granted (or restricted) by software and content licenses, as are copyright and authorship and attribution \cite{Kreutzer2014-ua}.
@@
 \emph{License}: license name.
-\emph{Redistribution}: right to redistribute, e.g.
-reuse in another benchmark/somewhere else.
-\emph{Modification}: right to modify the code to produce derivative works, e.g.
-in another benchmark/somewhere else.
+\emph{Redistribution}: right to redistribute, e.g., reuse in another benchmark/somewhere else.
+\emph{Modification}: right to modify the code to produce derivative works, e.g., in another benchmark/somewhere else.
 \emph{Sublicensing}: for derivative works, obligation to keep the same license or freedom to be licensed differently (even copyright).
@@
 
-Additionally, validation are recommended. Validations include content integrity (e.g., md5sum), adherence to a data specification (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data).
+Additionally, validation is recommended. Validations include content integrity (e.g., md5sum), adherence to a data specification (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data).
 

### 2024-10-29 manifesto-main.tex
Add edits following Mark's remarks

@@
 For \emph{method developers}, benchmarking allows their method to be compared to the state of the art, ideally using neutral datasets and metrics to avoid intrinsic bias \cite{Boulesteix2013-vy, Weber2019-el}.
-A diverse, accessible set of benchmarking datasets and metrics can address this and lower the entry barrier for method developers.
+A diverse, accessible set of benchmarking datasets and metrics can address this bias and lower the entry barrier for method developers.
 Moreover, it is important that the new method is compared to the current state of competing methods, to best mimic the choices available to a potential method user.
@@
 Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and efficient software stack; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results.
-Benchmarking systems are in charge of orchestrating both workflow execution and these benchmarking-specific tasks.
 
@@
 The impact of parameter choices can be included in a benchmark to test performance across a wider range of settings, or to find the optimal parameter settings.
+
 Parameters can be method-specific, but also dataset-specific, e.g., the true number of clusters when evaluating clustering algorithms.
-In this case, they can also be seen as (dataset-specific) metadata and the method performances can be compared depending on the availability of this metadata.
-A direct comparison between methods, with only some utilizing these metadata prevents differentiating between the effect of the parameter and the method itself for performance evaluation.
-Similarly, optimizing parameters for some but not other methods can skew results.
-Again, method-specific parameter tuning can be seen as a separate benchmarking task.
-Alternatively, a method that uses different settings can be interpreted as multiple distinct methods that are evaluated as part of the benchmark.
+In this case, they can also be seen as (dataset-specific) metadata.
+To differentiate between the effect of the parameter and the methods performances themselves, data-specific parameters should be shared across methods during evaluation.
+
+Similarly, optimizing parameters for some but not other methods can skew results and should be avoided.
+
+Aside of optimal and method-specific parameter searches, comparing a method using different settings against other methods can be viewed as a comparing fully distinct methods.
 This should be clearly communicated in the results.
@@
 
-For continuous and collaborative benchmarking, parameter tuning and usage should be governed and coordinated between all contributors.
-Ideally, the benchmarker specifies the parameter scope and related research objectives at the beginning of the benchmark to ensure comparability and interpretability of the results.
+For continuous and collaborative benchmarking, parameter optimization and usage should be governed and coordinated between all contributors, ideally ensuring comparability and interpretability of the results.
 
@@
 An active community is vital to the success of a benchmarking system.
-As mentioned, different roles and people can contribute to or interact with a benchmarking system.
 Benchmarkers (which may include multiple scientists) are responsible for planning and coordinating a benchmark, defining the task, possibly splitting it into subtasks or processing stages, and defining the data formats across the stages; the benchmarker also brings an authority role of how to review and approve contributions.
@@
 Software plays an important role in the current scientific landscape \cite{Howison2015-rc}, since it is widely used throughout the scientific process, including data collection, simulation, analysis and reporting.
-Researchers are encouraged to publish their data and code to enhance transparency and reproducibility in their work.
+Researchers are encouraged to follow FAIR guidelines and reproducibility in their work.
 Ideally, this would increase benchmark (software) reuse by the wider scientific community.
 In practice, however, it is sometimes easier to develop new code than to reuse existing software \cite{Trisovic2022-ol}.
-This leads to the phenomenon of `academic abandonware', where projects are forgotten in code repositories, exacerbating the reproducibility crisis in research; one of the noted challenges in benchmarking was the lack of extensibility \cite{Sonrel2023-te}.
-A common cause is the life cycle inherent to academic practice: after delivering a research output, there are often no resources to maintain a software package in the longer term.
+This leads to the phenomenon of `academic abandonware', where projects are forgotten in code repositories and not maintained after delivering a research output.
 
 Operational costs are given by the expected compute and storage requirements.
-In the context of benchmarking, methods under evaluation typically have unpredictable computational demands.
+In the context of benchmarking, methods under evaluation typically have variable computational demands.
 Benchmark designers could impose resource limits, thereby evaluating the methods under constraints.
@@
 A concise summary of license rights and restrictions, and compatibility across commonly-used open-source licenses can be found in Table~\ref{tab:licenses}.
-Benchmark contributions should bundle licenses to scholarly names including email and ORCID (as the canonical identifier), hence allowing seamless reuse, transparency and attribution.
+Benchmark contributions should be licensed and associated to scholarly names including email and ORCID (as the canonical identifier), hence allowing seamless reuse, transparency and attribution.
 

### 2024-10-29 manifesto-main.tex
More edits following Mark's comments

@@
 
-\section*{Parameter sweeps}
+\section*{(Hyper)parameters}
 
@@
 
-Parameters can be method-specific, but also dataset-specific, e.g., the true number of clusters when evaluating clustering algorithms.
+Dataset-specific parameters, e.g., the true number of clusters when evaluating clustering algorithms, are not method-specific.
 In this case, they can also be seen as (dataset-specific) metadata.
@@
 
-Similarly, optimizing parameters for some but not other methods can skew results and should be avoided.
+Similarly, optimizing method-specific parameters for some but not other methods can skew results and should be avoided.
 
@@
 This should be clearly communicated in the results.
-
-Altogether, evaluating parameters as part of a benchmark can give useful insights, but should be linked to a distinct and independent research objective.
 

### 2024-10-30 manifesto-main.tex
Ben's

@@
 Dataset-specific parameters, e.g., the true number of clusters when evaluating clustering algorithms, are not method-specific.
-In this case, they can also be seen as (dataset-specific) metadata.
+They can also be seen as (dataset-specific) metadata.
 To differentiate between the effect of the parameter and the methods performances themselves, data-specific parameters should be shared across methods during evaluation.
@@
 
-Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods).
+Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods)\cite{weber2019-el,jelizarow2010-kr}.
 One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a new method.
@@
     \item Flexibility vs.
-complexity: unrestricted freedom for method contributors can mean displacing the burden of maintainability towards the benchmarker.
-Fewer restrictions (e.g., simpler entry points) means that more methods can be included in a single benchmark; but, it can also imply a higher complexity for the initial setup by the benchmarker.
-Constraints can also enable validation, which can greatly help with the development process and quality assurance.
+constraints:  An unconstrained benchmark grants method contributors greater flexibility but requires more effort from the benchmark maintainer (maintenance, version support).
+Introducing constraints (such as software dependencies or common data formats) adds initial setup work but can boost maintainability through built-in validation and quality checks.
     \item  Achieving software reproducibility needs discipline and commitment \cite{Lamb2021-tb}.

### 2024-10-31 letter_to_editor.tex
cover letter sketched

\documentclass[a4paper]{article}

\usepackage[margin=1in]{geometry}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\hypersetup{
	unicode,
}

\newcommand{\submission}{X}
\newcommand{\journal}{Y}

\begin{document}

\noindent Dear Editor,\\

We would like to submit our draft ``\textbf{\submission}'' to \textbf{\journal}.\\

Benchmarking is an integral part of computational biology and bioinformatics. Yet benchmarks are far from being standardized, reusable or extensible, and their implementation is cumbersome in terms of repetitive tasks, such as managing reproducible software environments or formalizing computational workflows.\\

In this manuscript we describe the desired properties of benchmarking systems, e.g., software solutions to facilitate, standardize, coordinate and help distributing benchmarks.\\

We think our paper matches \journal's call for .\\

Thank you in advance for processing our submission. \\

Best regards,\\

Izaskun Mallona\\
\indent University of Zurich\\

\vspace{1cm}

\textit{Suggested reviewers}

\begin{itemize}
    \item Dr Salvador Capella Gutierrez, \texttt{salvador.capella@bsc.es}, leader of the Spanish National Bioinformatics (INB) Coordination Node at the Barcelona Supercomputing Center (BSC).
    \item Dr Malte Luecken, \texttt{malte.luecken@helmholtz-muenchen.de}, group leader of the Computational Cellular Health Group, Helmholtz Munich.
\end{itemize}

\end{document}

### 2024-10-31 letter_to_editor.tex
more changes

@@
     \item Dr Malte Luecken, \texttt{malte.luecken@helmholtz-muenchen.de}, group leader of the Computational Cellular Health Group, Helmholtz Munich.
+    \item Prof Dr Carole Goble , \texttt{carole.goble@manchester.ac.uk}, full professor and co-leader of the Information Management Group, School of Computer Science in the University of Manchester.
+    \item Prof Dr Johannes Koester, \texttt{johannes.koester@uni-due.de}, professor for Bioinformatics and Computational Oncology, Institute for AI in Medicine (IKIM), University Medicine Essen.
 \end{itemize}

### 2024-10-31 letter_to_editor.tex
Add Boulesteix

@@
     \item Dr Salvador Capella Gutierrez, \texttt{salvador.capella@bsc.es}, leader of the Spanish National Bioinformatics (INB) Coordination Node at the Barcelona Supercomputing Center (BSC).
-    \item Dr Malte Luecken, \texttt{malte.luecken@helmholtz-muenchen.de}, group leader of the Computational Cellular Health Group, Helmholtz Munich.
+    \item Dr Malte Luecken, \texttt{malte.luecken@helmholtz-muenchen.de}, leader of the Computational Cellular Health Group, Helmholtz Munich.
     \item Prof Dr Carole Goble , \texttt{carole.goble@manchester.ac.uk}, full professor and co-leader of the Information Management Group, School of Computer Science in the University of Manchester.
     \item Prof Dr Johannes Koester, \texttt{johannes.koester@uni-due.de}, professor for Bioinformatics and Computational Oncology, Institute for AI in Medicine (IKIM), University Medicine Essen.
+    \item Prof Dr Anne-Laure Boulesteix, \texttt{boulesteix@ibe.med.uni-muenchen.de}, leader of the Biometry in Molecular Medicine group, LMU Munich
 \end{itemize}

### 2024-11-04 manifesto-main.tex
charlotte's edits

@@
 
+Benchmarking (further defined below) is a critical step in method development and published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{Cao2023-jz}.
+A benchmarking study can be run by a single person or a small group, or it can be organized as a `challenge' or within a hackathon.
 In this work, we present a perspective on building a robust benchmarking ecosystem in bioinformatics, coming primarily from the methodologist's viewpoint (researchers developing computational tools).
 While our focus is bioinformatics, we aim to abstract principles and connect to similar challenges in adjacent  fields.
-Benchmarking (further defined below) is a critical step in method development: either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, as defined in \cite{Cao2023-jz}.
-A benchmarking study can be run by a single person or a small group, or it can be organized as a `challenge' or within a hackathon.
 
@@
 However, \textit{extensibility} of benchmarking studies is generally very low, as is the proportion of benchmarks using a formal workflow system.
-Thus, a formal system to orchestrate benchmarking workflows would largely impact the current landscape.
 
 Over 350 workflow languages, platforms or systems  \cite{Wratten2021-el, Amstutz2024-qk} and one standard (Common Workflow Language, CWL) exist \cite{Amstutz2016-vo}.
-Some workflow languages can export workflow definitions to CWL, hence helping computational FAIR guidelines: that is, bundling and processing data in association to their metadata, tracking new metadata generation, recording provenance, and being accessible \cite{Goble2020-ps}.
+Some workflow languages can export workflow definitions to CWL, hence helping computational FAIR guidelines: that is, bundling and processing data together with their metadata, tracking new metadata generation, recording provenance, and being accessible \cite{Goble2020-ps}.
 For benchmarking, we believe both the workflow layout and provenance can be formally defined (Figure~\ref{fig:formalization}).
@@
 To name a few of these systems, \texttt{ncbench} bundles workflow specification and software management with \texttt{snakemake} to benchmark result visualization with \texttt{Datavzrd} \cite{Hanssen2023-uf}.
-\texttt{OpenEBench} runs workflows in any workflow language, frequently \texttt{nextflow}, and makes use of openVRE as GUI and means of visualization \cite{Capella-Gutierrez2017-dh}.
+\texttt{OpenEBench} runs workflows in any workflow language, frequently \texttt{nextflow}, and makes use of \texttt{openVRE} as GUI and means of visualization \cite{Capella-Gutierrez2017-dh}.
 \texttt{OpenProblems} uses \texttt{Viash} and \texttt{nextflow} to handle software environments and workflows, respectively, and reports updated leaderboards with custom code \cite{Luecken2024-fk}.
@@
 Besides interactive dashboards for more nuanced result exploration \cite{bettr}, multi-criteria decision analysis (MCDA) \cite{Taherdoost2023-wd} has been proposed to guide users through benchmark results and multidimensional scaling (MDS) has been used to differentiate the effect of single datasets or metrics and recognize patterns beyond aggregated performance \cite{Niessl2022-fk}.
-An extensive analysis and presentation of the results not only helps users in the long term but also the whole scientific field to identify limitation across various use cases.
-The first advantage of such a strategy is to force the user to an attentive reading of the performance results so that the correct use-case and relevant methods can be identified.
-The second, more important, advantage to present benchmark results extensively is to identify the nuances, e.g., the scenarios, types of datasets, or aspects of performance that are systematically associated with poor results.
+An extensive analysis and presentation of the results helps not only users in the long term but also the whole scientific field to identify limitations across various use cases.
+The first advantage of such a strategy is to force the user to an attentive reading of the performance results so that the correct use case and relevant methods can be identified.
+The second, more important, advantage of presenting benchmark results extensively is to identify the nuances, e.g., the scenarios, types of datasets, or aspects of performance that are systematically associated with poor results.
 Being public and open to contributions from specialists of different fields, a benchmarking system could gather more reference datasets and methods, and allow the current state of the field to be better grasped.
@@
 Here,  recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance at a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial benchmark formalization could even provide the basis for the formal preregistration.
-In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, but the challenge to assess neutrality directly remains difficult.
+In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, but the challenge to assess neutrality directly remains.
 Other benchmarking models are possible, from challenges to hackathons, with the hope that a consensus of the masses can further help to establish neutrality.
@@
 
-So far, no single benchmarking system can accommodate all use-cases in bioinformatics.
+So far, no single benchmarking system can accommodate all use cases in bioinformatics.
 However, documentation of design choices for a given system can help to align expectations.
@@
 Likewise, proprietary formats should be avoided, since they may change and could require specific (proprietary) software to read.
-General language-agnostic formats often do not change over time and are thus align to FAIR principles \cite{Wilkinson2016-bh}.
+General language-agnostic formats often do not change over time and thus align to FAIR principles \cite{Wilkinson2016-bh}.
 Generally, using established standards is recommended (e.g., SAM, BED, FASTQ, etc for genomics), since related methods are developed around these formats.

### 2024-11-06 manifesto-main.tex
Someone's (Mark's?) changes

@@
 
-Benchmarking (further defined below) is a critical step in method development and published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{Cao2023-jz}.
+Benchmarking (further defined below) is a critical step in method development and is typically published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{Cao2023-jz}.
 A benchmarking study can be run by a single person or a small group, or it can be organized as a `challenge' or within a hackathon.
@@
 
-In our discussion, we outline concepts and specifications for an ideal \textit{benchmarking system}; we examine philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization practices.
+In our discussion, we outline concepts and specifications for an ideal benchmarking system; we examine philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization practices.
 
@@
 
-\section*{(Hyper)parameters}
-
-The impact of parameter choices can be included in a benchmark to test performance across a wider range of settings, or to find the optimal parameter settings.
-
+Parameter choices influence results interpretation, as they can be included in a benchmark to test performance across a wider range of settings, or to find the optimal parameter settings. 
 Dataset-specific parameters, e.g., the true number of clusters when evaluating clustering algorithms, are not method-specific.
@@
 To differentiate between the effect of the parameter and the methods performances themselves, data-specific parameters should be shared across methods during evaluation.
-
 Similarly, optimizing method-specific parameters for some but not other methods can skew results and should be avoided. 
-
 Aside of optimal and method-specific parameter searches, comparing a method using different settings against other methods can be viewed as a comparing fully distinct methods.
 This should be clearly communicated in the results.
-
 For continuous and collaborative benchmarking, parameter optimization and usage should be governed and coordinated between all contributors, ideally ensuring comparability and interpretability of the results.

### 2024-11-07 manifesto-main.tex
Mark's changes

@@
 
-Benchmarking (further defined below) is a critical step in method development and is typically published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{Cao2023-jz}.
+Benchmarking (further defined below) is a critical step in method development.
+Such studies are typically published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{cao2023-jz}.
 A benchmarking study can be run by a single person or a small group, or it can be organized as a `challenge' or within a hackathon.
@@
 
-In our discussion, we outline concepts and specifications for an ideal benchmarking system; we examine philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization practices.
+In our discussion, we outline concepts and specifications for an ideal benchmarking system; we examine philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization.
 
@@
 Given that benchmarks typically evaluate methods using a range of metrics and datasets, not all of which are relevant to all analysts, flexibility in defining the ranking approach and aggregation of metrics is preferable.
-Such flexibility, which requires access to various artifacts, is typically not part of published benchmarks today \cite{Sonrel2023-te}.
+Such flexibility is typically not part of published benchmarks today \cite{Sonrel2023-te}.
 However, a well-structured benchmarking system could accommodate flexible filtering and aggregation of metrics and datasets, and additionally provide access to the code and software stack that the analyst would need to apply the selected method to their own dataset.
@@
 Benchmarks can also be of high value to \emph{scientific journals} and \emph{funding agencies}.
-Well-executed benchmark studies can be highly utilized, influential and cited, and may point to gaps in the current body of methods, guiding future methodological developments; at the same time, in fast-moving fields like bioinformatics, benchmarks have the tendency to rapidly become stale.
+Well-executed benchmark studies can be highly utilized and influential, and may point to gaps in the current body of methods, guiding future methodological developments; at the same time, in fast-moving fields like bioinformatics, benchmarks have the tendency to rapidly become stale.
 Journals and funding bodies also have a vested interest in ensuring that published or funded method developments are performed to a high standard, that unnecessary redundancy is reduced, and that results are FAIR (findable, accessible, interoperable and reusable) for maximal benefit to the community \cite{Wilkinson2016-bh}.
@@
 Benchmarks are collections of data and source code, ideally executed as workflows within a computing environment.
-Indeed, recent studies highlight that code is frequently open, reusable and versioned, as are input datasets \cite{Sonrel2023-te, cao2023-jz}.
+Indeed, recent studies highlight that code is frequently open, reusable and versioned, as are input datasets \cite{cao2023-jz, Sonrel2023-te}.
 However, \textit{extensibility} of benchmarking studies is generally very low, as is the proportion of benchmarks using a formal workflow system.
@@
 Over 350 workflow languages, platforms or systems  \cite{Wratten2021-el, Amstutz2024-qk} and one standard (Common Workflow Language, CWL) exist \cite{Amstutz2016-vo}.
-Some workflow languages can export workflow definitions to CWL, hence helping computational FAIR guidelines: that is, bundling and processing data together with their metadata, tracking new metadata generation, recording provenance, and being accessible \cite{Goble2020-ps}.
+Some workflow languages can export workflow definitions to CWL, hence helping computational FAIR guidelines, i.e., bundling and processing data together with their metadata, tracking new metadata generation, recording provenance, and being accessible \cite{Goble2020-ps}.
 For benchmarking, we believe both the workflow layout and provenance can be formally defined (Figure~\ref{fig:formalization}).
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and efficient software stack; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results.
 
@@
 Parameter choices influence results interpretation, as they can be included in a benchmark to test performance across a wider range of settings, or to find the optimal parameter settings. 
-Dataset-specific parameters, e.g., the true number of clusters when evaluating clustering algorithms, are not method-specific.
-They can also be seen as (dataset-specific) metadata.
+Dataset-specific parameters, e.g., the true number of clusters when evaluating clustering algorithms, are not method-specific; they can be seen as (dataset-specific) metadata.
 To differentiate between the effect of the parameter and the methods performances themselves, data-specific parameters should be shared across methods during evaluation.
 Similarly, optimizing method-specific parameters for some but not other methods can skew results and should be avoided. 
-Aside of optimal and method-specific parameter searches, comparing a method using different settings against other methods can be viewed as a comparing fully distinct methods.
-This should be clearly communicated in the results.
+Aside of optimal and method-specific parameter searches, comparing a method under different settings against other methods can be viewed as comparing fully distinct methods, which should be clearly communicated in the results.
 For continuous and collaborative benchmarking, parameter optimization and usage should be governed and coordinated between all contributors, ideally ensuring comparability and interpretability of the results.
@@
 Users need a system that supports building benchmarks both ways, transitioning to centralization as the design matures.
-However, becoming centralized is not cost-free: it implies gatekeeping (contributors authenticate to the system and get access to computational resources) and adding measures to build trust \textit{via} transparency.
+However, becoming centralized is not cost-free: it implies gatekeeping (contributors authenticate to the system and get access to computational resources) and requires additional measures that build trust (e.g., transparency).
 
@@
 
-Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods)\cite{weber2019-el,jelizarow2010-kr}.
+Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods) \cite{Weber2019-el, Jelizarow2010-kr}.
 One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a new method.
@@
 Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then harder to establish.
-Here,  recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance at a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial benchmark formalization could even provide the basis for the formal preregistration.
+Here,  recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance at a registry \cite{Sullivan2019-wh, Olevska2021-if}.
 In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, but the challenge to assess neutrality directly remains.
@@
 Finally, there is a tension between preregistration and the concept of `continuous' benchmarking, where new datasets, methods and new metrics are added over time.
-The logic of continuous benchmarking is that as the problem becomes more understood in the field, it becomes increasingly clear what are good ways to evaluate the performance of methods.
+The logic of continuous benchmarking is that as the computational task becomes more understood, it becomes increasingly clear how to best evaluate the performance of methods.
 
@@
 Software plays an important role in the current scientific landscape \cite{Howison2015-rc}, since it is widely used throughout the scientific process, including data collection, simulation, analysis and reporting.
-Researchers are encouraged to follow FAIR guidelines and reproducibility in their work.
-Ideally, this would increase benchmark (software) reuse by the wider scientific community.
-In practice, however, it is sometimes easier to develop new code than to reuse existing software \cite{Trisovic2022-ol}.
+Even if open sourced, however, it is sometimes easier to develop new code than to reuse existing software \cite{Trisovic2022-ol}.
 This leads to the phenomenon of `academic abandonware', where projects are forgotten in code repositories and not maintained after delivering a research output.
 
-Operational costs are given by the expected compute and storage requirements.
+Operational costs of benchmarks are given by the expected compute and storage requirements.
 In the context of benchmarking, methods under evaluation typically have variable computational demands.
@@
 Furthermore, intuitive APIs and sensible defaults \cite{Proctor2018-do} are factors that contribute to highly usable systems.
-Structured approaches to documentation, including adopting frameworks like diataxis \cite{ProcidaUnknown-vf}, can facilitate adoption and new contributions to a benchmark.
+Structured approaches to documentation, including frameworks such as diataxis \cite{ProcidaUnknown-vf}, can facilitate adoption and new contributions to a benchmark.
 
@@
 Computational reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}.
-During benchmarking, a workflow runs code transforming inputs into outputs: this execution is affected by the base system (OS, compiler toolchains, libraries) and its configuration.
+During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system (OS, compiler toolchains, libraries) and its configuration.
 The benchmark definition should control the execution environment.
@@
 Hence, data, code and operating system dependencies are all accessible during the benchmarking process.
-For reproducibility, transparency and to build trust \cite{Laine2007-py}, this accessibility and freedom should be extended for others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how someone is implementing a method), change (bug fix, contribute) and improve (extend) benchmarks.
+For reproducibility, transparency and to build trust \cite{Laine2007-py}, this accessibility and freedom should be extended to others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how someone is implementing a method), change (bug fix, contribute) and improve (extend) benchmarks.
 Such freedoms are granted (or restricted) by software and content licenses, as are copyright and authorship and attribution \cite{Kreutzer2014-ua}.
@@
     \item Flexibility vs.
-constraints:  An unconstrained benchmark grants method contributors greater flexibility but requires more effort from the benchmark maintainer (maintenance, version support).
+constraints:  An unconstrained benchmark grants method contributors greater flexibility but requires more effort from the benchmarker (maintenance, version support).
 Introducing constraints (such as software dependencies or common data formats) adds initial setup work but can boost maintainability through built-in validation and quality checks.
-    \item  Achieving software reproducibility needs discipline and commitment \cite{Lamb2021-tb}.
-A benchmarker can state the desired degree of replicability and constraints; different tasks require different levels of compliance.
+    \item  Achieving software reproducibility needs discipline and commitment \cite{Lamb2021-tb}. A benchmarker can state the desired degree of replicability and constraints. 
     \item  Security concerns: decentralized runs lower the participation barrier, but also increase the attack  surface.
@@
 proprietary formats.
-File formats specific to a programming language should be avoided since those formats may change over time, security risks can arise \cite{Huynh2023-pq, Bleih2024-lv}, and in general they are not interoperable.
+File formats specific to a programming language should be avoided since those formats may change over time, security risks can arise \cite{Huynh2023-pq, Bleih2024-lv}, and they are not always interoperable.
 Likewise, proprietary formats should be avoided, since they may change and could require specific (proprietary) software to read.
@@
 Benchmarking is a cornerstone practice in bioinformatics, impacting how methods are conceived, developed, published and chosen.
-Benchmarking is driven by the effort of people with the expertise to address (biological and technical) problems.
-This expertise can be leveraged by current practices of community benchmarking, so scientific advance is eased by joint efforts to catalog open problems, generate (or simulate) appropriate input data, develop methods to address them, and evaluate their performances.
-Community benchmarking has been proved successful and very productive \cite{Moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.
+Benchmarking is driven by the effort of people with the expertise to address biological and technical problems.
+This expertise can be leveraged by current practices of community benchmarking, so scientific advances are eased by joint efforts to catalog open problems, generate (or simulate) appropriate input data, develop methods to address them, and evaluate their performances.
+Community benchmarking has been proven successful and  productive \cite{Moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.
 
-Here we discussed the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting technologies to potentially ease the technical, scientific and contribution challenges.
+Here, we discussed the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting strategies and technologies to potentially ease the technical, scientific and contribution challenges.
 

### 2024-11-18 manifesto-main.tex
mark's edits, before accepting

@@
 \author[1,2,*]{Izaskun Mallona}
-\author[1,2,3]{Almut Lütge}
-\author[2,4]{Charlotte Soneson}
+\author[2,3]{Charlotte Soneson}
 \author[1]{Ben Carrillo}
+\author[1,2,4]{Almut Lütge}
+\author[1]{Daniel Incicau}
 \author[1,2]{Reto Gerber}
-\author[1]{Daniel Incicau}
 \author[1,2]{Anthony Sonrel}
@@
 \affil[2]{SIB Swiss Institute of Bioinformatics, University of Zurich, 8057 Zurich, Switzerland}
-\affil[3]{Swiss Data Science Centre, 8092 Zurich, Switzerland}
-\affil[4]{Friedrich Miescher Institute for Biomedical Research, 4056 Basel, Switzerland}
+\affil[3]{Friedrich Miescher Institute for Biomedical Research, 4056 Basel, Switzerland}
+\affil[4]{Swiss Data Science Centre, 8092 Zurich, Switzerland}
 
@@
 
-Benchmarking (further defined below) is a critical step in method development.
+Benchmarking (defined below) is a critical step in method development.
 Such studies are typically published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{cao2023-jz}.
@@
 
-For \emph{method developers}, benchmarking allows their method to be compared to the state of the art, ideally using neutral datasets and metrics to avoid intrinsic bias \cite{Boulesteix2013-vy, Weber2019-el}.
+For \emph{method developers}, benchmarking allows their method to be compared to the state of the art, ideally using neutral datasets and metrics to avoid intrinsic bias \cite{Boulesteix2013-vy, Weber2019-el}; a recent call for reporting checklists was also made \cite{Brooks2024-rq}.
 A diverse, accessible set of benchmarking datasets and metrics can address this bias and lower the entry barrier for method developers.
@@
 Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then harder to establish.
-Here,  recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance at a registry \cite{Sullivan2019-wh, Olevska2021-if}.
+Here, recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance at a registry \cite{Sullivan2019-wh, Olevska2021-if, Teo2024-wk}.
 In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, but the challenge to assess neutrality directly remains.
@@
 
-Here, we discussed the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting strategies and technologies to potentially ease the technical, scientific and contribution challenges.
+Here, we discussed the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting strategies and technologies to potentially ease the technical and scientific challenges as well as how to organize contributions from communities.
 

### 2025-01-31 review_cycle_1/letter_to_editor.tex
Add figure draft

\documentclass[a4paper]{article}

\usepackage[margin=1in]{geometry}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\hypersetup{
	unicode,
}

\newcommand{\submission}{Building a continuous benchmarking ecosystem in bioinformatics}
\newcommand{\journal}{PLoS Computational Biology}
\newcommand{\submitter}{Izaskun Mallona and Mark D. Robinson}
\newcommand{\editor}{Dr BF Francis Ouellette, Dr Patricia M Palagi}

\begin{document}

\noindent Dear \editor,\\

Thank you for processing our manuscript \texttt{PCOMPBIOL-D-24-02003}.

\subsection*{Reviewer 1}

\end{document}

### 2025-02-07 review_cycle_1/highlighted_manifesto_main.tex
reviewer #1, reviewer #2

--- 2024-11-18 manifesto-main.tex
@@
 \usepackage{hyperref}
+\usepackage{xcolor}
+\usepackage[normalem]{ulem}
 
@@
 
+\begin{figure}
+    \centering
+    %\includegraphics[width=0.8\linewidth]{figures/fig_benchmarking.pdf}
+    {\color{red} \caption[Definitions.]{Benchmarks, workflows and benchmarking systems capabilities.}\label{fig:definitions}}
+\end{figure}
+
 We define a benchmark as a conceptual framework to evaluate the performance of computational methods for a given task, though it can extend to assessing other aspects, like data-generating technologies.
-A benchmark thus requires a well-defined task and a definition of correctness, or ground-truth, to be precisely defined in advance.
+A benchmark thus requires a well-defined task and a definition of correctness, or ground-truth, to be precisely defined in advance {\color{red} (Figure~\ref{fig:definitions})}.
 A benchmark then consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, and metrics.
@@
 Can benchmarking be systematized?
-Benchmarks should be orchestrated and run by \emph{benchmarking systems} to create `artifacts' (code snapshots, file outputs to be shared, performance outputs).
-The various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
-Such a system should facilitate components to be public, open and allow contributions (e.g., corrections or new components), such that the benchmarks of the future can run more continuously.
+{\color{red}Benchmarking aims to calculate, collect and report performance metrics of methods aiming to solve a task.
+Such process normally requires collecting the data, running the methods in an automated manner.
+Some of these tasks can be automated by workflow systems.
+Others, such as sharing results in an interactive manner and, particularly, result interpretation, go beyond what workflows offer.
+Similarly, collaborative benchmarking and challenges add extra layers of complexity, including collaboration tracking, hidding ground truths, or the ability to release multiple versions of the same benchmark as it grows (Figure~\ref{fig:definitions}).
+Moreover, scientific benchmarking borrows procedures from the scientific method and from scientific publishing, hence having expectations of fairness, reproducibility, transparency, and trust.
+\sout{Benchmarks should be orchestrated and run by}} \emph{Benchmarking systems} {\color{red} aim to orchestrate the workflow management, benchmarking and community engagement aspects of the process  \sout{to create} to generate benchmark} `artifacts' (code snapshots, file outputs to be shared, performance outputs) {\color{red} in a systematic way and following good practices and standards.
+Namely,}  the various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
+Such {\color{red}\sout{a}benchmarking} systems should facilitate components to be public, open and allow contributions (e.g., corrections or new components), such that the benchmarks of the future can run more continuously.
 At the outset of a benchmark, a benchmark `definition' could be envisioned, which could give a formal specification of the entire set of components (and the pattern of artifacts to be generated).
 The benchmark definition could even be expressed as a single configuration file that specifies the scope and topology of components to include, details of the repositories (code implementations with versions), the instructions to create software environments (accessible across compute architectures), the parameters used, and what components to snapshot for a release.
-A benchmark definition in terms of layers and their corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.
+A benchmark definition in terms of layers and their corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.}
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figures/bens_layers.pdf}
+    \includegraphics[width=0.8\linewidth]{../original_submission/figures/bens_layers.pdf}
     \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges.
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{figures/bens_formalization.pdf}
+    \includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
     \caption[Benchmark formalization.]{Benchmark formalization.
@@
 Working together would mean sharing not only datasets and code snippets, but also execution environments (software but perhaps also hardware), alignment on parameters to evaluate and eventually sharing, consolidating, and interpreting results.
-However, because there are no standards of how benchmarks are implemented \cite{Sonrel2023-te}, another disadvantage of decentralized benchmarks is that, while these components may eventually be openly shared as a product of a research study, the contents of a benchmark are non-interoperable, even across benchmarks of the same task.
+However, because there are not many standards of how benchmarks are implemented \cite{Sonrel2023-te}, another disadvantage of decentralized benchmarks is that, while these components may eventually be openly shared as a product of a research study, the contents of a benchmark are non-interoperable, even across benchmarks of the same task.
 Furthermore, decentralized benchmarks have been criticized on the basis of `inflating model performance' due to decisions made in preprocessing or parameter tuning \cite{Luecken2024-fk}; such issues could be mitigated by gathering the `wisdom of crowds' that is implicit in running centralized benchmarks.
@@
 
-\hfill
+%\hfill
 
@@
 
-\section*{Funding}
-
-MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36\_14).
+% the two sections below have been added to the online system now
+%\section*{Funding}
+%MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36\_14).
 CS is supported by the Novartis Research Foundation.
 
-\section*{Disclosure statement}
-
-No potential conflict of interest is reported by the authors
+%\section*{Competing interests}
+%No conflicts of interest are declared by the authors.
 

### 2025-02-07 review_cycle_1/highlighted_manifesto_main.tex
add placeholders

@@
 Journals and funding bodies also have a vested interest in ensuring that published or funded method developments are performed to a high standard, that unnecessary redundancy is reduced, and that results are FAIR (findable, accessible, interoperable and reusable) for maximal benefit to the community \cite{Wilkinson2016-bh}.
-Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
+{\color{red} Recently published benchmarks and challenges have been conceived as or ported to benchmarking systems to ensure these properties.} Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
+
+{\color{red} Discuss DREAM, CASP etc and scientific communities here (reviewer 1 and 2).}
 
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
 
@@
 
+{\color{red} Here all content on whether software is part of the backend or not, conda, pixi, web assembly etc - reviewer 3}
+
 \section*{Software licenses and attribution}

### 2025-02-10 review_cycle_1/highlighted_manifesto_main.tex
reviewer 2 and 3 now

@@
 
+{\color{red}
+Aside of task completion-tailored metrics, computing performance (i.e.
+wallclock, on-disk read and write operations), requirements (e.g.
+GPUs, minimum required memory) and their scalability depending on the input size (e.g.
+whether runtimes increase linearly, quadratically, etc) are key, and routine, benchmarking metrics.
+To measure these accurately, benchmarking systems ideally would provide means to: i, characterize the hardware running the benchmark, so methods runtimes are comparable if run on homogeneous systems;  ii, track performances (CPU or GPU usage, memory) when running each benchmark component; iii, introduce little overhead during the profiling (so profiling does not considerably slow down runtimes); iv, collect variability in performance by running each method several times; v, go beyond reporting average resources usage and provide them over execution time (e.g.
+so embarrasingly parallel algorithms can be told apart from largely single-cored runs with few parallelism bursts).
+}
+
 \section*{Collaboration, gatekeeping and benchmarking infrastructure go hand in hand}
@@
 
+{\color{red}
+Mature communities are key in sustaining and reflecting the existing challenge in a given field.
+For instance, CASP, a continuous benchmarking by nature, has a established community defining the critical assesments scope, input data, target definitions, and composite and normalized metrics to equally weight multiple performance scores \cite{simpkin2023tertiary}.
+At the continuous benchmark level, CASP keeps track of past performances and introduces metrics to evaluate possible artifacts, i.e.
+whether round progress could be arising from different target difficulty \cite{kryshtafovych2019critical}.
+To facilitate community contribution, we believe benchmarking systems can lower the barrier to contributing, hence making the contribution efforts less likely to be biased.
+}
+
 Finally, there is a tension between preregistration and the concept of `continuous' benchmarking, where new datasets, methods and new metrics are added over time.
@@
 Computational reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}.
-During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system (OS, compiler toolchains, libraries) and its configuration.
+During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system ({\color{red}\sout{OS} operating system}, compiler toolchains, libraries) and its configuration.
 The benchmark definition should control the execution environment.
@@
 Making the long-term replicability of a benchmark tractable implies adopting a sane software management system.
-Several solutions exist, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.
+Several solutions exist, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.
 
@@
 
-{\color{red} Here all content on whether software is part of the backend or not, conda, pixi, web assembly etc - reviewer 3}
+{\color{red} 
+The responsibility of handling the software environment poses an interesting challenge to benchmarking systems.
+On one hand, some workflow managers provide means to install and activate software and re-run workflows when software specification changes, even if the workflow topology remains the same.
+For these, software is the workflow manager's responsability.
+On the other hand, some benchmarks might be addressed to comparing the same workflow topology run with different software backends (i.e., with conda vs compilled) or with different versions of some software dependencies components (i.e.
+several samtools/htslib releases).
+Benchmarking systems should provide enough flexibility to take control and make the latter possible.
+
+Finally, scientific software management handling is known for its nuances, including abundant dependencies with known incompatibilities and complex installation recipes \cite{droge2023-ax}.
+Scientific software management is handled via traditional (linux) package managers (Ri.e.
+PMs, DEBs), central software stacks and enviornment modules (i.e.
+Lmod), conda, containers, or read-only network filesystems providing ready-to-use installations (CernVM-FS).
+These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; conda is speeding up their deployment procedures (via py-rattler/pixi) and allowing compilation to Web assembly (to execute code within Web browser)s; or apptainer (former Singularity) provides rootless containarization suitable to run in scientific clusters.
+}
 

### 2025-02-10 review_cycle_1/highlighted_manifesto_main.tex
Add new figure

@@
     \centering
-    %\includegraphics[width=0.8\linewidth]{figures/fig_benchmarking.pdf}
-    {\color{red} \caption[Definitions.]{Benchmarks, workflows and benchmarking systems capabilities.}\label{fig:definitions}}
+    \includegraphics[width=\linewidth]{figures/fig_benchmarking.pdf}
+    {\color{red} \caption[Run types, properties, capabilities and features.]{Solo- and community-benchmarks, workflows, and benchmarking system-enabled (\textit{bs}) benchmarks differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
+These run intrinsic properties have consequences in terms of features (i.e.
+transparency, forkability, availability for metaanalysis) and technical capabilities (i.e.
+software versioning or ability to run anonymously and locally (stealth mode).
+\label{fig:definitions}}
 \end{figure}

### 2025-02-10 review_cycle_1/highlighted_manifesto_main.tex
syntax fixes

@@
 Such studies are typically published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{cao2023-jz}.
-A benchmarking study can be run by a single person or a small group, or it can be organized as a `challenge' or within a hackathon.
+A benchmarking study can be run by a single person {\color{red} (solo benchmarking)} or a small group, or it can be organized as a `challenge' or within a hackathon.
 In this work, we present a perspective on building a robust benchmarking ecosystem in bioinformatics, coming primarily from the methodologist's viewpoint (researchers developing computational tools).
@@
 
-In our discussion, we outline concepts and specifications for an ideal benchmarking system; we examine philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization.
+In our discussion, we outline concepts {\color{red} related to workflow automation, benchmark formalization and benchmarking-specific tasks beyond the former, including \sout{and specifications for an ideal benchmarking system; we examine}} philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization.
 
@@
 \label{fig:definitions}}
+    }
 \end{figure}
@@
 {\color{red}Benchmarking aims to calculate, collect and report performance metrics of methods aiming to solve a task.
-Such process normally requires collecting the data, running the methods in an automated manner.
+Such process normally requires collecting appropriate data and running methods in an automated manner.
 Some of these tasks can be automated by workflow systems.
-Others, such as sharing results in an interactive manner and, particularly, result interpretation, go beyond what workflows offer.
-Similarly, collaborative benchmarking and challenges add extra layers of complexity, including collaboration tracking, hidding ground truths, or the ability to release multiple versions of the same benchmark as it grows (Figure~\ref{fig:definitions}).
+Others, such as sharing results in an interactive manner and, particularly, result interpretation and making results reusable by others, go beyond what workflows offer.
+Similarly, collaborative benchmarking and challenges add extra layers of complexity, including collaboration tracking, hiding ground truths, or the ability to release multiple versions of the same benchmark as it develops (Figure~\ref{fig:definitions}).
 Moreover, scientific benchmarking borrows procedures from the scientific method and from scientific publishing, hence having expectations of fairness, reproducibility, transparency, and trust.
@@
 The benchmark definition could even be expressed as a single configuration file that specifies the scope and topology of components to include, details of the repositories (code implementations with versions), the instructions to create software environments (accessible across compute architectures), the parameters used, and what components to snapshot for a release.
-A benchmark definition in terms of layers and their corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.}
+A benchmark definition in terms of layers and their corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.
 
@@
 
-{\color{red} Discuss DREAM, CASP etc and scientific communities here (reviewer 1 and 2).}
+{\color{red} Collaborative benchmarking efforts enunciate field-specific challenges and provide a venue for \emph{communities} to address them. For instance, DREAM
+
+GIAB
+
+CASP
+
+have successfully shown how collaborative benchmarks can help in moving fields forward.}
 
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environments handling from workflow execution is further discussed in section~\ref{softwareenvs})}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
 
@@
 
-\subsection*{Reproducible software environments}
+\subsection*{Reproducible software environments \label{softwareenvs}}
 
@@
 Making the long-term replicability of a benchmark tractable implies adopting a sane software management system.
-Several solutions exist, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.
+Several solutions exist, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{droge2023-ax}.
 
@@
 This conundrum extends to the operating system layer and to system dependencies (e.g., compilers).
-In terms of benchmark re-runnability, fully free software system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{Droge2023-ax}.
+In terms of benchmark re-runnability, fully free software system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{droge2023-ax}.
 

### 2025-02-10 review_cycle_1/highlighted_manifesto_main.tex
More

@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environments handling from workflow execution is further discussed in section~\ref{softwareenvs})}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environments handling from workflow execution is further discussed in section~`Reproducible software environments')}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
 
@@
 Benchmark contributions should be licensed and associated to scholarly names including email and ORCID (as the canonical identifier), hence allowing seamless reuse, transparency and attribution.
+{\color{red} As a consequence, properly licensed benchmarking pieces (e.g.
+a procedure to compute a clustering metric; or a clustering method) could be reused across benchmarks; we envision assisted benchmark design tools to leverage and reuse benchmark components, perhaps guided by AI.}
 
@@
 Additionally, static and dynamic code analysis in the continuous integration process can aid in approving potentially risky submissions.
+    \item Versioning granularity: a single benchmark could address performance of a given method across its versions when run on the same datasets; independently, this benchmark could be updated so several benchmark versions are released.
+Benchmarking systems should provide versioning at both component and benchmark levels.
 \end{itemize}
@@
 
+{\color{red}
+\subsection*{Funding}
+@Mark
+}
 \section*{Author contributions}
 
-Conceptualization: all. Visualization: BC. Original draft preparation: all. Writing, reviewing and editing: all. Supervision: IM, MDR. Funding acquisition: MDR.
+Conceptualization: all. {\color{red} \sout{Visualization} Figures: IM,} BC. Original draft preparation: all. Writing, reviewing and editing: all. Supervision: IM, MDR. Funding acquisition: MDR.
 

### 2025-02-10 review_cycle_1/highlighted_manifesto_main.tex
add giab, casp, dream

@@
 
-{\color{red} Collaborative benchmarking efforts enunciate field-specific challenges and provide a venue for \emph{communities} to address them. For instance, DREAM
-
-GIAB
-
-CASP
-
-have successfully shown how collaborative benchmarks can help in moving fields forward.}
+{\color{red} Collaborative benchmarking efforts enunciate field-specific challenges and provide a venue for \emph{communities} to address them.
+For instance, DREAM \cite{meyer2021advances}, GIAB \cite{zook2019open} and CASP \cite{moult2005-ne} challenges and benchmarks have demonstrated how collaborative benchmarks can help in moving fields forward.}
 
@@
 This expertise can be leveraged by current practices of community benchmarking, so scientific advances are eased by joint efforts to catalog open problems, generate (or simulate) appropriate input data, develop methods to address them, and evaluate their performances.
-Community benchmarking has been proven successful and  productive \cite{Moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.
+Community benchmarking has been proven successful and  productive \cite{moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.
 

### 2025-02-12 original_submission/manifesto-main.tex
Add figure draft
Mark changes

--- 2024-11-18 manifesto-main.tex
@@
 
-\affil[*]{\{mark.robinson,izaskun.mallona\}@mls.uzh.ch}
+\affil[*]{\{izaskun.mallona,mark.robinson\}@mls.uzh.ch}
 
@@
 
+% the two sections below have been added to the online system now
 \section*{Funding}
-
 MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36\_14).
@@
 
-\section*{Disclosure statement}
-
-No potential conflict of interest is reported by the authors
+\section*{Competing interests}
+No conflicts of interest are declared by the authors.
 

### 2025-02-12 review_cycle_1/highlighted_manifesto_main.tex
Prior to Mark's edits
Mark changes

@@
     \includegraphics[width=\linewidth]{figures/fig_benchmarking.pdf}
-    {\color{red} \caption[Run types, properties, capabilities and features.]{Solo- and community-benchmarks, workflows, and benchmarking system-enabled (\textit{bs}) benchmarks differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
-These run intrinsic properties have consequences in terms of features (i.e.
-transparency, forkability, availability for metaanalysis) and technical capabilities (i.e.
-software versioning or ability to run anonymously and locally (stealth mode).
+    {\color{red} \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
+These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e.
+software versioning or ability to run anonymously and locally in stealth mode).
 \label{fig:definitions}}
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environments handling from workflow execution is further discussed in section~`Reproducible software environments')}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environment handling from workflow execution is further discussed in section~`Reproducible software environments')}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
 
@@
 {\color{red}
-Aside of task completion-tailored metrics, computing performance (i.e.
-wallclock, on-disk read and write operations), requirements (e.g.
-GPUs, minimum required memory) and their scalability depending on the input size (e.g.
-whether runtimes increase linearly, quadratically, etc) are key, and routine, benchmarking metrics.
-To measure these accurately, benchmarking systems ideally would provide means to: i, characterize the hardware running the benchmark, so methods runtimes are comparable if run on homogeneous systems;  ii, track performances (CPU or GPU usage, memory) when running each benchmark component; iii, introduce little overhead during the profiling (so profiling does not considerably slow down runtimes); iv, collect variability in performance by running each method several times; v, go beyond reporting average resources usage and provide them over execution time (e.g.
-so embarrasingly parallel algorithms can be told apart from largely single-cored runs with few parallelism bursts).
+Aside of task completion-tailored metrics, computing performance (i.e., wallclock, on-disk read-write operations), requirements (e.g., GPUs, minimum memory) and their scalability depending on the input size (e.g., whether runtimes increase linearly, quadratically, etc) are key routine benchmarking metrics.
+To measure these accurately, benchmarking systems ideally would provide means to: i) characterize the hardware running the benchmark, so method runtimes are comparable if run on homogeneous systems;  ii) track performance (CPU or GPU usage, memory) when running each benchmark component; iii) introduce minimal overhead during profiling (so profiling does not considerably slow down runtimes); iv) collect variability in performance by running each method several times; and, v) go beyond reporting average resource usage and provide them over execution time (e.g., so embarrassingly-parallel algorithms can be differentiated from largely single-cored runs with few parallelism bursts).
 }
@@
 {\color{red}
-Mature communities are key in sustaining and reflecting the existing challenge in a given field.
-For instance, CASP, a continuous benchmarking by nature, has a established community defining the critical assesments scope, input data, target definitions, and composite and normalized metrics to equally weight multiple performance scores \cite{simpkin2023tertiary}.
-At the continuous benchmark level, CASP keeps track of past performances and introduces metrics to evaluate possible artifacts, i.e.
-whether round progress could be arising from different target difficulty \cite{kryshtafovych2019critical}.
-To facilitate community contribution, we believe benchmarking systems can lower the barrier to contributing, hence making the contribution efforts less likely to be biased.
+Mature communities are key to sustaining and reflecting the existing challenges in a given field.
+For instance, CASP, a continuous benchmark by nature, has a established community that defines the scope of critical assessment, input data, target definitions, and composite and normalized metrics to equally weight multiple performance scores \cite{simpkin2023tertiary}.
+At the continuous benchmark level, CASP keeps track of past performances and introduces metrics to evaluate possible artifacts, i.e., whether round progress could arises from different target difficulty \cite{kryshtafovych2019critical}.
+To facilitate community contribution, we believe that benchmarking systems can lower the barrier to contributing, hence making the contribution efforts less likely to be biased.
 }
@@
 {\color{red} 
-The responsibility of handling the software environment poses an interesting challenge to benchmarking systems.
+The responsibility of handling software environments poses an interesting challenge to benchmarking systems.
 On one hand, some workflow managers provide means to install and activate software and re-run workflows when software specification changes, even if the workflow topology remains the same.
-For these, software is the workflow manager's responsability.
-On the other hand, some benchmarks might be addressed to comparing the same workflow topology run with different software backends (i.e., with conda vs compilled) or with different versions of some software dependencies components (i.e.
-several samtools/htslib releases).
+For these, software is the workflow manager's responsibility.
+On the other hand, some benchmarks might be explicitly designed to compare the same workflow topology run with different software backends (i.e., with conda vs compiled) or with different versions of a dependency (e.g., samtools/htslib releases).
 Benchmarking systems should provide enough flexibility to take control and make the latter possible.
@@
 Finally, scientific software management handling is known for its nuances, including abundant dependencies with known incompatibilities and complex installation recipes \cite{droge2023-ax}.
-Scientific software management is handled via traditional (linux) package managers (Ri.e.
-PMs, DEBs), central software stacks and enviornment modules (i.e.
-Lmod), conda, containers, or read-only network filesystems providing ready-to-use installations (CernVM-FS).
-These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; conda is speeding up their deployment procedures (via py-rattler/pixi) and allowing compilation to Web assembly (to execute code within Web browser)s; or apptainer (former Singularity) provides rootless containarization suitable to run in scientific clusters.
+Scientific software management is handled via traditional (linux) package managers (RPMs, DEBs), central software stacks and environment modules (i.e., Lmod), conda, containers, or read-only network filesystems providing ready-to-use installations (CernVM-FS).
+These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; conda is speeding up their deployment procedures (via py-rattler/pixi) and allowing compilation to Web assembly (to execute code within Web browser); or apptainer (former Singularity) provides rootless containerization suitable to run in scientific clusters.
 }
@@
 Benchmark contributions should be licensed and associated to scholarly names including email and ORCID (as the canonical identifier), hence allowing seamless reuse, transparency and attribution.
-{\color{red} As a consequence, properly licensed benchmarking pieces (e.g.
-a procedure to compute a clustering metric; or a clustering method) could be reused across benchmarks; we envision assisted benchmark design tools to leverage and reuse benchmark components, perhaps guided by AI.}
+{\color{red} As a consequence, properly-licensed benchmark components (e.g., a procedure to compute a clustering metric; or a clustering method) could be reused across benchmarks; we envision assisted benchmark design tools to leverage and reuse benchmark components, perhaps guided by AI.}
 
@@
 {\color{red}
-\subsection*{Funding}
-@Mark
+\section*{Funding}
+MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36\_14).
+CS is supported by the Novartis Research Foundation.
+
+\section*{Conflict of interest}
+
+We report no conflict of interest.
 }

### 2025-02-12 review_cycle_1/highlighted_manifesto_main.tex
Accepting Mark's comments, first round

@@
 Journals and funding bodies also have a vested interest in ensuring that published or funded method developments are performed to a high standard, that unnecessary redundancy is reduced, and that results are FAIR (findable, accessible, interoperable and reusable) for maximal benefit to the community \cite{Wilkinson2016-bh}.
-{\color{red} Recently published benchmarks and challenges have been conceived as or ported to benchmarking systems to ensure these properties.} Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
+{\color{red} Recently published benchmarks and challenges have been conceived as or ported to benchmarking systems to ensure these properties \cite{Nevers_2022,Bryce_Smith_2023, Pardo_Palacios_2023}.} Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
 
@@
 {\color{red}
-Aside of task completion-tailored metrics, computing performance (i.e., wallclock, on-disk read-write operations), requirements (e.g., GPUs, minimum memory) and their scalability depending on the input size (e.g., whether runtimes increase linearly, quadratically, etc) are key routine benchmarking metrics.
-To measure these accurately, benchmarking systems ideally would provide means to: i) characterize the hardware running the benchmark, so method runtimes are comparable if run on homogeneous systems;  ii) track performance (CPU or GPU usage, memory) when running each benchmark component; iii) introduce minimal overhead during profiling (so profiling does not considerably slow down runtimes); iv) collect variability in performance by running each method several times; and, v) go beyond reporting average resource usage and provide them over execution time (e.g., so embarrassingly-parallel algorithms can be differentiated from largely single-cored runs with few parallelism bursts).
+Aside of task-tailored scores (i.e., intrinsic or extrinsic cluster validity measures for clustering tasks), collecting metrics to report computing performance (i.e., wallclock, on-disk read-write operations), requirements (e.g., GPUs, minimum memory) and their scalability depending on the input size (e.g., whether runtimes increase linearly, quadratically, etc) is key during benchmarking.
+To measure these performance metrics accurately, benchmarking systems ideally would provide means to: i) characterize the hardware running the benchmark, so method runtimes are comparable;  ii) track performance (CPU or GPU usage, memory) when running each benchmark component; iii) introduce minimal overhead during profiling; iv) collect variability in performance by running each method several times; and, v) go beyond reporting average resource usage and provide them over execution time (e.g., so embarrassingly-parallel algorithms can be differentiated from largely single-cored runs with few parallelism bursts).
 }
@@
 For instance, CASP, a continuous benchmark by nature, has a established community that defines the scope of critical assessment, input data, target definitions, and composite and normalized metrics to equally weight multiple performance scores \cite{simpkin2023tertiary}.
-At the continuous benchmark level, CASP keeps track of past performances and introduces metrics to evaluate possible artifacts, i.e., whether round progress could arises from different target difficulty \cite{kryshtafovych2019critical}.
+At the continuous benchmark level, CASP keeps track of past performances and introduces metrics to evaluate possible artifacts, i.e., whether CASP round progress could arise from different target difficulty \cite{kryshtafovych2019critical}.
 To facilitate community contribution, we believe that benchmarking systems can lower the barrier to contributing, hence making the contribution efforts less likely to be biased.
@@
 On the other hand, some benchmarks might be explicitly designed to compare the same workflow topology run with different software backends (i.e., with conda vs compiled) or with different versions of a dependency (e.g., samtools/htslib releases).
-Benchmarking systems should provide enough flexibility to take control and make the latter possible.
+Benchmarking systems should provide enough flexibility to make all variants possible (i.e., decouple the benchmark topology from the software backend).
 
@@
 Additionally, static and dynamic code analysis in the continuous integration process can aid in approving potentially risky submissions.
-    \item Versioning granularity: a single benchmark could address performance of a given method across its versions when run on the same datasets; independently, this benchmark could be updated so several benchmark versions are released.
-Benchmarking systems should provide versioning at both component and benchmark levels.
+    \item {\color{red} Versioning granularity: a single benchmark could address performance of a given method across its versions when run on the same datasets; independently, this benchmark could be updated so several benchmark versions are released.
+Benchmarking systems should provide versioning at both component and benchmark levels.}
 \end{itemize}

### 2025-02-12 review_cycle_1/highlighted_manifesto_main.tex
Fix fig1 typo

@@
 These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e.
-software versioning or ability to run anonymously and locally in stealth mode).
+software versioning or ability to run locally without leaving traces - in \textit{incognito} mode).
+Green, blue and red ticks depict available, perhaps available and unavailable features, respectively.
 \label{fig:definitions}}
@@
 For instance, CASP, a continuous benchmark by nature, has a established community that defines the scope of critical assessment, input data, target definitions, and composite and normalized metrics to equally weight multiple performance scores \cite{simpkin2023tertiary}.
-At the continuous benchmark level, CASP keeps track of past performances and introduces metrics to evaluate possible artifacts, i.e., whether CASP round progress could arise from different target difficulty \cite{kryshtafovych2019critical}.
+At the continuous benchmark level, CASP keeps track of past performances and introduces metrics to evaluate possible artifacts, i.e., whether progress during a CASP round could arise from different target difficulty \cite{kryshtafovych2019critical}.
 To facilitate community contribution, we believe that benchmarking systems can lower the barrier to contributing, hence making the contribution efforts less likely to be biased.
@@
 For these, software is the workflow manager's responsibility.
-On the other hand, some benchmarks might be explicitly designed to compare the same workflow topology run with different software backends (i.e., with conda vs compiled) or with different versions of a dependency (e.g., samtools/htslib releases).
+On the other hand, some benchmarks might be designed to compare the same workflow topology run with different software backends (i.e., with conda vs compiled) or with different versions of a dependency (e.g., samtools/htslib releases).
 Benchmarking systems should provide enough flexibility to make all variants possible (i.e., decouple the benchmark topology from the software backend).

### 2025-02-17 review_cycle_1/highlighted_manifesto_main.tex
Add aligned figure and other changes by Charlotte

@@
     \centering
-    \includegraphics[width=\linewidth]{figures/fig_benchmarking.pdf}
+    \includegraphics[width=\linewidth]{review_cycle_1/figures/fig_benchmarking_aligned.pdf}
     {\color{red} \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
@@
 We define a benchmark as a conceptual framework to evaluate the performance of computational methods for a given task, though it can extend to assessing other aspects, like data-generating technologies.
-A benchmark thus requires a well-defined task and a definition of correctness, or ground-truth, to be precisely defined in advance {\color{red} (Figure~\ref{fig:definitions})}.
+A benchmark thus requires a well-defined task and typically a definition of correctness, or ground-truth, to be precisely defined in advance {\color{red} (Figure~\ref{fig:definitions})}.
 A benchmark then consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, and metrics.
@@
 The benchmark definition could even be expressed as a single configuration file that specifies the scope and topology of components to include, details of the repositories (code implementations with versions), the instructions to create software environments (accessible across compute architectures), the parameters used, and what components to snapshot for a release.
-A benchmark definition in terms of layers and their corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.
+A benchmark definition in terms of layers and the corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.
 
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environment handling from workflow execution is further discussed in section~`Reproducible software environments')}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments')}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
 
@@
 Making the long-term replicability of a benchmark tractable implies adopting a sane software management system.
-Several solutions exist, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{droge2023-ax}.
+Several solutions exist, from containerization (e.g., \texttt{apptainer} or \texttt{Singularity}) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{droge2023-ax}.
 
@@
 Finally, scientific software management handling is known for its nuances, including abundant dependencies with known incompatibilities and complex installation recipes \cite{droge2023-ax}.
-Scientific software management is handled via traditional (linux) package managers (RPMs, DEBs), central software stacks and environment modules (i.e., Lmod), conda, containers, or read-only network filesystems providing ready-to-use installations (CernVM-FS).
-These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; conda is speeding up their deployment procedures (via py-rattler/pixi) and allowing compilation to Web assembly (to execute code within Web browser); or apptainer (former Singularity) provides rootless containerization suitable to run in scientific clusters.
+Scientific software management is handled via traditional (Linux) package managers (RPMs, DEBs), central software stacks and environment modules (i.e., Lmod), conda, containers, or read-only network filesystems providing ready-to-use installations (CernVM-FS).
+These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; \texttt{conda} is speeding up their deployment procedures (via \texttt{py-rattler}/\texttt{pixi}) and allowing compilation to Web assembly (to execute code within a web browser); and \texttt{apptainer} (former \texttt{Singularity}) provides rootless containerization suitable for running in scientific clusters.
 }

### 2025-02-17 review_cycle_1/highlighted_manifesto_main.tex
Update figure

@@
 \emph{Hardware}: infrastructure and its costs.
-\emph{Data}: dataset archival, openness, interoperability, archival, and selection.
+\emph{Data}: dataset archival, openness, interoperability, and selection.
 \emph{Software}: method implementations, reproducibility, workflow execution, continuous integration and delivery (CI/CD), versioning, quality assurance (QA).
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
+    \includegraphics[width=0.75\linewidth]{figures/bens_formalization_arrows.pdf}
     \caption[Benchmark formalization.]{Benchmark formalization.

### 2025-02-21 review_cycle_1/highlighted_manifesto_main.tex
Prior to adding the unmarked one

@@
 
-\affil[*]{\{izaskun.mallona,mark.robinson\}@mls.uzh.ch}
+\affil[*]{Correspondence: izaskun.mallona@gmail.com,mark.robinson@mls.uzh.ch}
 
@@
 {\color{red}Benchmarking aims to calculate, collect and report performance metrics of methods aiming to solve a task.
-Such process normally requires collecting appropriate data and running methods in an automated manner.
+Such a process normally requires collecting appropriate data and running methods in an automated manner.
 Some of these tasks can be automated by workflow systems.
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{figures/bens_formalization_arrows.pdf}
+        %\includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
+    \includegraphics[width=0.75\linewidth]{figures/bens_formalization_arrows.pdf} % alternative
     \caption[Benchmark formalization.]{Benchmark formalization.
@@
 Aside of task-tailored scores (i.e., intrinsic or extrinsic cluster validity measures for clustering tasks), collecting metrics to report computing performance (i.e., wallclock, on-disk read-write operations), requirements (e.g., GPUs, minimum memory) and their scalability depending on the input size (e.g., whether runtimes increase linearly, quadratically, etc) is key during benchmarking.
-To measure these performance metrics accurately, benchmarking systems ideally would provide means to: i) characterize the hardware running the benchmark, so method runtimes are comparable;  ii) track performance (CPU or GPU usage, memory) when running each benchmark component; iii) introduce minimal overhead during profiling; iv) collect variability in performance by running each method several times; and, v) go beyond reporting average resource usage and provide them over execution time (e.g., so embarrassingly-parallel algorithms can be differentiated from largely single-cored runs with few parallelism bursts).
+To measure these performance metrics accurately, benchmarking systems ideally would provide means to: i) characterize the hardware running the benchmark, so method runtimes are comparable;  ii) track performance (CPU or GPU usage, memory) when running each benchmark component; iii) introduce minimal overhead during profiling; iv) collect variability in performance by running each method several times; and, v) go beyond reporting average resource usage and provide them over execution time (e.g., so embarrassingly-parallel algorithms can be differentiated from largely single-cored runs with few bursts of parallel execution).
 }
@@
 CS is supported by the Novartis Research Foundation.
+The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
 

### 2025-02-21 review_cycle_1/updated_manifesto_main.tex
add updated, not highlighted, tex

--- 2025-02-21 review_cycle_1/highlighted_manifesto_main.tex
@@
 \usepackage{authblk}
-\usepackage{xcolor} % Load the xcolor package for color options
+%\usepackage{xcolor} % Load the xcolor package for color options
 \usepackage{microtype}
@@
 \usepackage{hyperref}
-\usepackage{xcolor}
+
 \usepackage[normalem]{ulem}
@@
 Such studies are typically published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{cao2023-jz}.
-A benchmarking study can be run by a single person {\color{red} (solo benchmarking)} or a small group, or it can be organized as a `challenge' or within a hackathon.
+A benchmarking study can be run by a single person  (solo benchmarking) or a small group, or it can be organized as a `challenge' or within a hackathon.
 In this work, we present a perspective on building a robust benchmarking ecosystem in bioinformatics, coming primarily from the methodologist's viewpoint (researchers developing computational tools).
@@
 
-In our discussion, we outline concepts {\color{red} related to workflow automation, benchmark formalization and benchmarking-specific tasks beyond the former, including \sout{and specifications for an ideal benchmarking system; we examine}} philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization.
+In our discussion, we outline concepts related to workflow automation, benchmark formalization and benchmarking-specific tasks beyond the former, including philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization.
 
@@
     \includegraphics[width=\linewidth]{review_cycle_1/figures/fig_benchmarking_aligned.pdf}
-    {\color{red} \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
+     \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
 These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e.
@@
 \label{fig:definitions}}
-    }
+
 \end{figure}
@@
 We define a benchmark as a conceptual framework to evaluate the performance of computational methods for a given task, though it can extend to assessing other aspects, like data-generating technologies.
-A benchmark thus requires a well-defined task and typically a definition of correctness, or ground-truth, to be precisely defined in advance {\color{red} (Figure~\ref{fig:definitions})}.
+A benchmark thus requires a well-defined task and typically a definition of correctness, or ground-truth, to be precisely defined in advance  (Figure~\ref{fig:definitions}).
 A benchmark then consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, and metrics.
@@
 Can benchmarking be systematized?
-{\color{red}Benchmarking aims to calculate, collect and report performance metrics of methods aiming to solve a task.
+Benchmarking aims to calculate, collect and report performance metrics of methods aiming to solve a task.
 Such a process normally requires collecting appropriate data and running methods in an automated manner.
@@
 Moreover, scientific benchmarking borrows procedures from the scientific method and from scientific publishing, hence having expectations of fairness, reproducibility, transparency, and trust.
-\sout{Benchmarks should be orchestrated and run by}} \emph{Benchmarking systems} {\color{red} aim to orchestrate the workflow management, benchmarking and community engagement aspects of the process  \sout{to create} to generate benchmark} `artifacts' (code snapshots, file outputs to be shared, performance outputs) {\color{red} in a systematic way and following good practices and standards.
-Namely,}  the various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
-Such {\color{red}\sout{a}benchmarking} systems should facilitate components to be public, open and allow contributions (e.g., corrections or new components), such that the benchmarks of the future can run more continuously.
+\emph{Benchmarking systems} aim to orchestrate the workflow management, benchmarking and community engagement aspects of the process to generate benchmark `artifacts' (code snapshots, file outputs to be shared, performance outputs) in a systematic way and following good practices and standards.
+Namely,  the various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
+Such systems should facilitate components to be public, open and allow contributions (e.g., corrections or new components), such that the benchmarks of the future can run more continuously.
 At the outset of a benchmark, a benchmark `definition' could be envisioned, which could give a formal specification of the entire set of components (and the pattern of artifacts to be generated).
@@
 Journals and funding bodies also have a vested interest in ensuring that published or funded method developments are performed to a high standard, that unnecessary redundancy is reduced, and that results are FAIR (findable, accessible, interoperable and reusable) for maximal benefit to the community \cite{Wilkinson2016-bh}.
-{\color{red} Recently published benchmarks and challenges have been conceived as or ported to benchmarking systems to ensure these properties \cite{Nevers_2022,Bryce_Smith_2023, Pardo_Palacios_2023}.} Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
+Recently published benchmarks and challenges have been conceived as or ported to benchmarking systems to ensure these properties \cite{Nevers_2022,Bryce_Smith_2023, Pardo_Palacios_2023}.
+Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
 
-{\color{red} Collaborative benchmarking efforts enunciate field-specific challenges and provide a venue for \emph{communities} to address them.
-For instance, DREAM \cite{meyer2021advances}, GIAB \cite{zook2019open} and CASP \cite{moult2005-ne} challenges and benchmarks have demonstrated how collaborative benchmarks can help in moving fields forward.}
+Collaborative benchmarking efforts enunciate field-specific challenges and provide a venue for \emph{communities} to address them.
+For instance, DREAM \cite{meyer2021advances}, GIAB \cite{zook2019open} and CASP \cite{moult2005-ne} challenges and benchmarks have demonstrated how collaborative benchmarks can help in moving fields forward.
 
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments')}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack (the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments'); managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results (Figure~\ref{fig:definitions})}
 
@@
 
-{\color{red}
 Aside of task-tailored scores (i.e., intrinsic or extrinsic cluster validity measures for clustering tasks), collecting metrics to report computing performance (i.e., wallclock, on-disk read-write operations), requirements (e.g., GPUs, minimum memory) and their scalability depending on the input size (e.g., whether runtimes increase linearly, quadratically, etc) is key during benchmarking.
 To measure these performance metrics accurately, benchmarking systems ideally would provide means to: i) characterize the hardware running the benchmark, so method runtimes are comparable;  ii) track performance (CPU or GPU usage, memory) when running each benchmark component; iii) introduce minimal overhead during profiling; iv) collect variability in performance by running each method several times; and, v) go beyond reporting average resource usage and provide them over execution time (e.g., so embarrassingly-parallel algorithms can be differentiated from largely single-cored runs with few bursts of parallel execution).
-}
 
@@
 
-{\color{red}
 Mature communities are key to sustaining and reflecting the existing challenges in a given field.
@@
 To facilitate community contribution, we believe that benchmarking systems can lower the barrier to contributing, hence making the contribution efforts less likely to be biased.
-}
 
@@
 Computational reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}.
-During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system ({\color{red}\sout{OS} operating system}, compiler toolchains, libraries) and its configuration.
+During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system (operating system, compiler toolchains, libraries) and its configuration.
 The benchmark definition should control the execution environment.
@@
 
-{\color{red} 
 The responsibility of handling software environments poses an interesting challenge to benchmarking systems.
@@
 These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; \texttt{conda} is speeding up their deployment procedures (via \texttt{py-rattler}/\texttt{pixi}) and allowing compilation to Web assembly (to execute code within a web browser); and \texttt{apptainer} (former \texttt{Singularity}) provides rootless containerization suitable for running in scientific clusters.
-}
 
@@
 Benchmark contributions should be licensed and associated to scholarly names including email and ORCID (as the canonical identifier), hence allowing seamless reuse, transparency and attribution.
-{\color{red} As a consequence, properly-licensed benchmark components (e.g., a procedure to compute a clustering metric; or a clustering method) could be reused across benchmarks; we envision assisted benchmark design tools to leverage and reuse benchmark components, perhaps guided by AI.}
+As a consequence, properly-licensed benchmark components (e.g., a procedure to compute a clustering metric; or a clustering method) could be reused across benchmarks; we envision assisted benchmark design tools to leverage and reuse benchmark components, perhaps guided by AI.
 
@@
 Additionally, static and dynamic code analysis in the continuous integration process can aid in approving potentially risky submissions.
-    \item {\color{red} Versioning granularity: a single benchmark could address performance of a given method across its versions when run on the same datasets; independently, this benchmark could be updated so several benchmark versions are released.
-Benchmarking systems should provide versioning at both component and benchmark levels.}
+    \item Versioning granularity: a single benchmark could address performance of a given method across its versions when run on the same datasets; independently, this benchmark could be updated so several benchmark versions are released.
+Benchmarking systems should provide versioning at both component and benchmark levels.
 \end{itemize}
@@
 
-{\color{red}
 \section*{Funding}
@@
 We report no conflict of interest.
-}
+
 \section*{Author contributions}
 
-Conceptualization: all. {\color{red} \sout{Visualization} Figures: IM,} BC. Original draft preparation: all. Writing, reviewing and editing: all. Supervision: IM, MDR. Funding acquisition: MDR.
+Conceptualization: all. Figures: IM, BC. Original draft preparation: all. Writing, reviewing and editing: all. Supervision: IM, MDR. Funding acquisition: MDR.
 

### 2025-04-23 review_cycle_1/highlighted_manifesto_main.tex
before updating

@@
 
-\affil[*]{Correspondence: izaskun.mallona@gmail.com,mark.robinson@mls.uzh.ch}
+\affil[*]{Correspondence: izaskun.mallona{\color{red}@gmail.com},mark.robinson@mls.uzh.ch}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{review_cycle_1/figures/fig_benchmarking_aligned.pdf}
+    \includegraphics[width=\linewidth]{review_cycle_1/figures/fig1_fig_benchmarking_aligned.pdf}
     {\color{red} \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{../original_submission/figures/bens_layers.pdf}
+    \includegraphics[width=0.8\linewidth]{../original_submission/figures/fig2_bens_layers.pdf}
     \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges.
@@
         %\includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
-    \includegraphics[width=0.75\linewidth]{figures/bens_formalization_arrows.pdf} % alternative
+    \includegraphics[width=0.75\linewidth]{figures/fig3_bens_formalization_arrows.pdf} % alternative
     \caption[Benchmark formalization.]{Benchmark formalization.

### 2025-04-23 0_original_submission/letter_to_editor.tex
before generating the bbl

--- 2024-10-31 letter_to_editor.tex
@@
 
-\newcommand{\submission}{X}
-\newcommand{\journal}{Y}
+\newcommand{\submission}{Building a continuous benchmarking ecosystem in bioinformatics}
+\newcommand{\journal}{NAR Genomics and Bioinformatics}
+\newcommand{\submitter}{Izaskun Mallona and Mark D. Robinson}
+\newcommand{\editor}{Dr Ana Rojas}
 
@@
 
-\noindent Dear Editor,\\
+Dear \editor,\\
 
@@
 
-In this manuscript we describe the desired properties of benchmarking systems, e.g., software solutions to facilitate, standardize, coordinate and help distributing benchmarks.\\
+In this manuscript we argue benchmarking can benefit from benchmarking systems, e.g., software solutions to facilitate, standardize, coordinate and help distributing benchmarks, and discuss the desired properties of such benchmarking systems.\\
 
-We think our paper matches \journal's call for .\\
+We think our paper matches \journal's scope and interest in robust and reproducible benchmarking of genomics and bioinformatics methods.\\
+
+Our manuscript is available as a preprint at \url{https://arxiv.org/abs/2409.15472}.\\
 
@@
 
-Izaskun Mallona\\
+\submitter\\
 \indent University of Zurich\\
@@
     \item Prof Dr Johannes Koester, \texttt{johannes.koester@uni-due.de}, professor for Bioinformatics and Computational Oncology, Institute for AI in Medicine (IKIM), University Medicine Essen.
-    \item Prof Dr Anne-Laure Boulesteix, \texttt{boulesteix@ibe.med.uni-muenchen.de}, leader of the Biometry in Molecular Medicine group, LMU Munich
+    \item Prof Dr Anne-Laure Boulesteix, \texttt{boulesteix@ibe.med.uni-muenchen.de}, leader of the Biometry in Molecular Medicine group, LMU Munich.
+    \item Prof Dr. Serguei Mangul, \texttt{serghei.mangul@gmail.com}, assistant professor of clinical pharmacy, University of Southern California.
+    \item Prof Dr. Jean Yang, \texttt{jean.yang@sydney.edu.au}, professor, School of Mathematics and Statistics of the University of Sydney.
+    \item Prof. Dr. Gregory Grann, \texttt{ggrant@pennmedicine.upenn.edu}, associate professor of genetics, Perelman School of Medicine
 \end{itemize}

### 2025-04-23 1_first_review/manuscript.tex
before generating the bbl

--- 2025-02-21 review_cycle_1/updated_manifesto_main.tex
@@
-% Emerald Publishing - Construction Innovation Submission Template
+1_% Emerald Publishing - Construction Innovation Submission Template
 % by Oleksandr Melnyk
@@
     \centering
-    \includegraphics[width=\linewidth]{review_cycle_1/figures/fig_benchmarking_aligned.pdf}
+    \includegraphics[width=\linewidth]{figures/fig1_fig_benchmarking_aligned.pdf}
      \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{../original_submission/figures/bens_layers.pdf}
+    \includegraphics[width=0.8\linewidth]{figures/fig2_bens_layers.pdf}
     \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges.
@@
         %\includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
-    \includegraphics[width=0.75\linewidth]{figures/bens_formalization_arrows.pdf} % alternative
+    \includegraphics[width=0.75\linewidth]{./figures/fig3_bens_formalization_arrows.pdf} % alternative
     \caption[Benchmark formalization.]{Benchmark formalization.

### 2025-04-23 1_first_review/tracked_changes.tex
before generating the bbl

--- 2025-04-23 review_cycle_1/highlighted_manifesto_main.tex
@@
     \centering
-    \includegraphics[width=\linewidth]{review_cycle_1/figures/fig1_fig_benchmarking_aligned.pdf}
+    \includegraphics[width=\linewidth]{./figures/fig1_fig_benchmarking_aligned.pdf}
     {\color{red} \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{../original_submission/figures/fig2_bens_layers.pdf}
+    \includegraphics[width=0.8\linewidth]{./figures/fig2_bens_layers.pdf}
     \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges.

### 2025-04-23 1_first_review/manuscript.tex

@@
-1_% Emerald Publishing - Construction Innovation Submission Template
+% Emerald Publishing - Construction Innovation Submission Template
 % by Oleksandr Melnyk
@@
 
-%\usepackage{makecell} % multiline table cells
+%\usepackage[backend=biber]{biblatex}
 
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack (the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments'); managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results (Figure~\ref{fig:definitions})}
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack (the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments'); managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results (Figure~\ref{fig:definitions}).
 
@@
 
-\bibliographystyle{vancouver} % elsarticle-num} % vancouver}  
-\bibliography{clean} % your .bib file
+%\bibliographystyle{vancouver} % elsarticle-num} % vancouver}  
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% \bibliography{clean} % your .bib file
+
+\begin{thebibliography}{10}
+
+\bibitem{cao2023-jz}
+Cao Y, Yu L, Torkel M, Kim S, Lin Y, Yang P, et~al.
+\newblock The current landscape and emerging challenges of benchmarking
+  single-cell methods.
+\newblock bioRxiv. 2023; p. 2023.12.19.572303.
+
+\bibitem{Sonrel2023-te}
+Sonrel A, Luetge A, Soneson C, Mallona I, Germain PL, Knyazev S, et~al.
+\newblock Meta-analysis of (single-cell method) benchmarks reveals the need for
+  extensibility and interoperability.
+\newblock Genome Biol. 2023;24:119.
+
+\bibitem{Boulesteix2013-vy}
+Boulesteix AL, Lauer S, Eugster MJA.
+\newblock A plea for neutral comparison studies in computational sciences.
+\newblock PLoS One. 2013;8:e61562.
+
+\bibitem{Weber2019-el}
+Weber LM, Saelens W, Cannoodt R, Soneson C, Hapfelmeier A, Gardner PP, et~al.
+\newblock Essential guidelines for computational method benchmarking.
+\newblock Genome Biol. 2019;20:125.
+
+\bibitem{Brooks2024-rq}
+Brooks TG, Lahens NF, Mrčela A, Grant GR.
+\newblock Challenges and best practices in omics benchmarking.
+\newblock Nature Review Genetics. 2024;25(5):326--339.
+
+\bibitem{Rich2024-jh}
+Rich JM, Moses L, Einarsson PH, Jackson K, Luebbert L, Booeshaghi AS, et~al.
+\newblock The impact of package selection and versioning on single-cell
+  {RNA}-seq analysis.
+\newblock bioRxivorg. 2024; p. 2024.04.04.588111.
+
+\bibitem{Wilkinson2016-bh}
+Wilkinson MD, Dumontier M, Aalbersberg IJJ, Appleton G, Axton M, Baak A, et~al.
+\newblock The {FAIR} Guiding Principles for scientific data management and
+  stewardship.
+\newblock Sci Data. 2016;3:160018.
+
+\bibitem{Nevers_2022}
+Nevers Y, Jones TEM, Jyothi D, Yates B, Ferret M, Portell-Silva L, et~al.
+\newblock The Quest for Orthologs orthology benchmark service in 2022.
+\newblock Nucleic Acids Research. 2022;50(W1):W623–W632.
+\newblock doi:{10.1093/nar/gkac330}.
+
+\bibitem{Bryce_Smith_2023}
+Bryce-Smith S, Burri D, Gazzara MR, Herrmann CJ, Danecka W, Fitzsimmons CM,
+  et~al.
+\newblock Extensible benchmarking of methods that identify and quantify
+  polyadenylation sites from RNA-seq data.
+  2023;doi:{10.1101/2023.06.23.546284}.
+
+\bibitem{Pardo_Palacios_2023}
+Pardo-Palacios FJ, Wang D, Reese F, Diekhans M, Carbonell-Sala S, Williams B,
+  et~al.
+\newblock Systematic assessment of long-read RNA-seq methods for transcript
+  identification and quantification. 2023;doi:{10.1101/2023.07.25.550582}.
+
+\bibitem{meyer2021advances}
+Meyer P, Saez-Rodriguez J.
+\newblock Advances in systems biology modeling: 10 years of crowdsourcing DREAM
+  challenges.
+\newblock Cell Systems. 2021;12(6):636--653.
+
+\bibitem{zook2019open}
+Zook JM, McDaniel J, Olson ND, Wagner J, Parikh H, Heaton H, et~al.
+\newblock An open resource for accurately benchmarking small variant and
+  reference calls.
+\newblock Nature Biotechnology. 2019;37(5):561--566.
+
+\bibitem{moult2005-ne}
+Moult J.
+\newblock A decade of {CASP}: progress, bottlenecks and prognosis in protein
+  structure prediction.
+\newblock Curr Opin Struct Biol. 2005;15:285--289.
+
+\bibitem{Wratten2021-el}
+Wratten L, Wilm A, Göke J.
+\newblock Reproducible, scalable, and shareable analysis pipelines with
+  bioinformatics workflow managers.
+\newblock Nat Methods. 2021;18:1161--1168.
+
+\bibitem{Amstutz2024-qk}
+Amstutz P, Mikheev M, Crusoe MR, Tijanić N, Lampa S. Existing Workflow
+  systems; 2024.
+\newblock Available from: \url{https://s.apache.org/existing-workflow-systems}.
+
+\bibitem{Amstutz2016-vo}
+Amstutz P, Crusoe MR, Tijanić N, Chapman B, Chilton J, Heuer M, et~al.. Common
+  Workflow Language, {v1}.0; 2016.
+\newblock Available from:
+  \url{https://research.manchester.ac.uk/files/57032693/cwl_1.0_workflow.pdf}.
+
+\bibitem{Goble2020-ps}
+Goble C, Cohen-Boulakia S, Soiland-Reyes S, Garijo D, Gil Y, Crusoe MR, et~al.
+\newblock FAIR computational workflows.
+\newblock Data Intell. 2020;2:108--121.
+
+\bibitem{Hanssen2023-uf}
+Hanssen F, Gabernet G, Smith NH, Mertes C, Neogi AG, Brandhoff L, et~al.
+\newblock NCBench: providing an open, reproducible, transparent, adaptable, and
+  continuous benchmark approach for {DNA}-sequencing-based variant calling.
+\newblock F1000Research. 2023;12:1125.
+
+\bibitem{Capella-Gutierrez2017-dh}
+Capella-Gutierrez S, de~la Iglesia D, Haas J, Lourenco A, Fernández JM,
+  Repchevsky D, et~al.
+\newblock Lessons Learned: Recommendations for Establishing Critical Periodic
+  Scientific Benchmarking.
+\newblock bioRxiv. 2017; p. 181677.
+
+\bibitem{Luecken2024-fk}
+Luecken MD, Gigante S, Burkhardt DB, Cannoodt R, Strobl DC, Markov NS, et~al.
+\newblock Defining and benchmarking open problems in single-cell analysis.
+\newblock Res Sq. 2024;.
+
+\bibitem{Khan2018-ot}
+Khan RU, Zhang X, Kumar R, Aboagye EO.
+\newblock Evaluating the performance of {ResNet} model based on image
+  recognition.
+\newblock In: Proceedings of the 2018 International Conference on Computing and
+  Artificial Intelligence. ACM; 2018.Available from:
+  \url{https://dl.acm.org/doi/10.1145/3194452.3194461}.
+
+\bibitem{Jumper2021-oq}
+Jumper J, Evans R, Pritzel A, Green T, Figurnov M, Ronneberger O, et~al.
+\newblock Highly accurate protein structure prediction with AlphaFold.
+\newblock Nature. 2021;596:583--589.
+
+\bibitem{Lutge2021-mt}
+Lütge A, Zyprych-Walczak J, Brykczynska~Kunzmann U, Crowell HL, Calini D,
+  Malhotra D, et~al.
+\newblock CellMixS: quantifying and visualizing batch effects in single-cell
+  RNA-seq data.
+\newblock Life Sci Alliance. 2021;4.
+
+\bibitem{Reinke2024-dw}
+Reinke A, Tizabi MD, Baumgartner M, Eisenmann M, Heckmann-Nötzel D, Kavur AE,
+  et~al.
+\newblock Understanding metric-related pitfalls in image analysis validation.
+\newblock Nat Methods. 2024;21:182--194.
+
+\bibitem{funkyheatmap}
+Cannoodt R, Saelens W. type [;]Available from:
+  \url{https://cran.r-project.org/web/packages/funkyheatmap/index.html}.
+
+\bibitem{Saelens2019-jy}
+Saelens W, Cannoodt R, Todorov H, Saeys Y.
+\newblock A comparison of single-cell trajectory inference methods.
+\newblock Nat Biotechnol. 2019;37:547--554.
+
+\bibitem{Jelizarow2010-kr}
+Jelizarow M, Guillemot V, Tenenhaus A, Strimmer K, Boulesteix AL.
+\newblock Over-optimism in bioinformatics: an illustration.
+\newblock Bioinformatics. 2010;26:1990--1998.
+
+\bibitem{Boulesteix2010-zj}
+Boulesteix AL.
+\newblock Over-optimism in bioinformatics research.
+\newblock Bioinformatics. 2010;26:437--439.
+
+\bibitem{Strobl2024-lp}
+Strobl C, Leisch F.
+\newblock Against the "one method fits all data sets" philosophy for comparison
+  studies in methodological research.
+\newblock Biom J. 2024;66:e2200104.
+
+\bibitem{bettr}
+Marini F, Soneson C. type [; 2024]Available from:
+  \url{https://www.bioconductor.org/packages/release/bioc/html/bettr.html}.
+
+\bibitem{Taherdoost2023-wd}
+Taherdoost H, Madanchian M.
+\newblock Multi-criteria decision making ({MCDM}) methods and concepts.
+\newblock Encyclopedia (Basel, 2021). 2023;3:77--87.
+
+\bibitem{Niessl2022-fk}
+Nießl C, Herrmann M, Wiedemann C, Casalicchio G, Boulesteix AL.
+\newblock Over‐optimism in benchmark studies and the multiplicity of design
+  and analysis options when interpreting their results.
+\newblock Wiley Interdiscip Rev Data Min Knowl Discov. 2022;12:e1441.
+
+\bibitem{omnibenchmark}
+Mallona I, Luetge A, Carrillo B, Incicau D, Gerber R, Sonrel A, et~al.
+\newblock Omnibenchmark (alpha) for continuous and open benchmarking in
+  bioinformatics.
+\newblock arXiv preprint arXiv:240917038. 2024;.
+
+\bibitem{Greenstein2016-os}
+Greenstein S, Zhu F.
+\newblock Open content, Linus’ law, and neutral Point of View.
+\newblock Inf Syst Res. 2016;27:618--635.
+
+\bibitem{Sullivan2019-wh}
+Sullivan I, DeHaven A, Mellor D.
+\newblock Open and reproducible research on open science framework.
+\newblock Curr Protoc Essent Lab Tech. 2019;18:e32.
+
+\bibitem{Olevska2021-if}
+Olevska A, Bert B, Ebrahimi L, Schonfelder G, Heinl C.
+\newblock Ensuring reproducible research requires a support infrastructure: The
+  value of public registries to publishers.
+\newblock Sci Editor. 2021; p. 4--7.
+
+\bibitem{Teo2024-wk}
+Teo AYY, Squair JW, Courtine G, Skinnider MA.
+\newblock Best practices for differential accessibility analysis in single-cell
+  epigenomics.
+\newblock Nature Communications. 2024;15(1):8805.
+
+\bibitem{Siepe2023-tz}
+Siepe BS, Bartoš F, Morris TP, Boulesteix AL, Heck DW, Pawel S.
+\newblock Simulation studies for methodological research in psychology: A
+  standardized template for planning, preregistration, and reporting.
+\newblock PsyArXiv. 2023;.
+
+\bibitem{simpkin2023tertiary}
+Simpkin AJ, Mesdaghi S, S{\'a}nchez~Rodr{\'\i}guez F, Elliott L, Murphy DL,
+  Kryshtafovych A, et~al.
+\newblock Tertiary structure assessment at CASP15.
+\newblock Proteins: Structure, Function, and Bioinformatics.
+  2023;91(12):1616--1635.
+
+\bibitem{kryshtafovych2019critical}
+Kryshtafovych A, Schwede T, Topf M, Fidelis K, Moult J.
+\newblock Critical assessment of methods of protein structure prediction
+  (CASP)—Round XIII.
+\newblock Proteins: Structure, Function, and Bioinformatics.
+  2019;87(12):1011--1020.
+
+\bibitem{Howison2015-rc}
+Howison J, Deelman E, McLennan MJ, Ferreira~da Silva R, Herbsleb JD.
+\newblock Understanding the scientific software ecosystem and its impact:
+  Current and future measures.
+\newblock Res Eval. 2015;24:454--470.
+
+\bibitem{Trisovic2022-ol}
+Trisovic A, Lau MK, Pasquier T, Crosas M.
+\newblock A large-scale study on research code quality and execution.
+\newblock Sci Data. 2022;9:60.
+
+\bibitem{Pike1984-we}
+Pike R, Kernighan B.
+\newblock The {UNIX} system: Program design in the {UNIX} environment.
+\newblock AT T Bell Lab Tech J. 1984;63:1595--1605.
+
+\bibitem{Proctor2018-do}
+Proctor RW, Schneider DW.
+\newblock Hick’s law for choice reaction time: A review.
+\newblock Quarterly Journal of Experimental Psychology. 2018;71:1281--1299.
+
+\bibitem{ProcidaUnknown-vf}
+Procida D. Diátaxis;.
+\newblock Available from: \url{https://diataxis.fr/}.
+
+\bibitem{Hill2024-gf}
+Hill D, Antunes BA, Bertrand A, Nguifo EM, Yon L, Nautré-Domanski J, et~al..
+  Machine learning and reproducibility impact of random numbers; 2024.
+\newblock Available from: \url{https://hal.science/hal-04642175v1}.
+
+\bibitem{Kurtzer2017-mn}
+Kurtzer GM, Sochat V, Bauer MW.
+\newblock Singularity: Scientific containers for mobility of compute.
+\newblock PLoS One. 2017;12:e0177459.
+
+\bibitem{Hoste2012-gg}
+Hoste K, Timmerman J, Georges A, De~Weirdt S.
+\newblock EasyBuild: Building Software with Ease.
+\newblock In: 2012 SC Companion: High Performance Computing, Networking Storage
+  and Analysis. IEEE; 2012. p. 572--582.
+\newblock Available from:
+  \url{https://ieeexplore.ieee.org/abstract/document/6495863}.
+
+\bibitem{Gamblin2015-ll}
+Gamblin T, LeGendre M, Collette MR, Lee GL, Moody A, de~Supinski BR, et~al.
+\newblock The Spack package manager: bringing order to {HPC} software chaos.
+\newblock In: Proceedings of the International Conference for High Performance
+  Computing, Networking, Storage and Analysis. ACM; 2015.Available from:
+  \url{https://dl.acm.org/doi/10.1145/2807591.2807623}.
+
+\bibitem{Blomer2013-jk}
+Blomer J, Buncic P, Meusel R.
+\newblock The {CernVM} file system.
+\newblock CERN, Geneva, Switzerland, Tech Rep. 2013; p. 2--1.
+
+\bibitem{Boissonneault2019-wm}
+Boissonneault M, Oldeman BE, Taylor RP.
+\newblock Providing a unified software environment for Canada's national
+  advanced computing centers.
+\newblock In: Proceedings of the Practice and Experience in Advanced Research
+  Computing on Rise of the Machines (learning). ACM; 2019.Available from:
+  \url{https://dl.acm.org/doi/10.1145/3332186.3332210}.
+
+\bibitem{droge2023-ax}
+Dröge B, Holanda~Rusu V, Hoste K, van Leeuwen C, O'Cais A, Röblitz T.
+\newblock EESSI: A cross‐platform ready‐to‐use optimised scientific
+  software stack.
+\newblock Softw Pract Exp. 2023;53:176--210.
+
+\bibitem{Laine2007-py}
+Laine C, Goodman SN, Griswold ME, Sox HC.
+\newblock Reproducible Research: Moving toward Research the Public Can Really
+  Trust.
+\newblock Ann Intern Med. 2007;146:450.
+
+\bibitem{Kreutzer2014-ua}
+Kreutzer T.
+\newblock Open content: A practical guide to using Creative Commons licences.
+  2014;.
+
+\bibitem{Potter2015-fs}
+Potter ME, Smith T.
+\newblock Making code citable with Zenodo and GitHub.
+\newblock Software Sustainibility Institute. 2015;.
+
+\bibitem{Sicilia2017-ko}
+Sicilia MA, García-Barriocanal E, Sánchez-Alonso S.
+\newblock Community curation in open dataset repositories: Insights from
+  zenodo.
+\newblock Procedia Comput Sci. 2017;106:54--60.
+
+\bibitem{van-de-Sandt2019-wd}
+van~de Sandt S, Nielsen LH, Ioannidis A, Muench A, Henneken E, Accomazzi A,
+  et~al.
+\newblock Practice meets principle: Tracking software and data citations to
+  Zenodo DOIs.
+\newblock arXiv [csDL]. 2019;.
+
+\bibitem{Lamb2021-tb}
+Lamb C, Zacchiroli S.
+\newblock Reproducible Builds: Increasing the integrity of software supply
+  chains.
+\newblock arXiv [csSE]. 2021;.
+
+\bibitem{Huynh2023-pq}
+Huynh D. How Python Data Science Libraries Can Be Hijacked (and What You Can Do
+  About It); 2023.
+\newblock Available from:
+  \url{https://blog.mithrilsecurity.io/how-python-data-science-libraries-can-be-hijacked-and-what-you-can-do-about-it-2/}.
+
+\bibitem{Bleih2024-lv}
+Bleih A. New Vulnerability in {R}'s Deserialization Discovered; 2024-05-02.
+\newblock Available from:
+  \url{https://cyberint.com/blog/research/new-vulnerability-in-rs-deserialization-discovered/}.
+
+\bibitem{LeboUnknown-om}
+Lebo T, Sahoo S, McGuinness D, Belhajjame K, Cheney J, Corsar D, et~al..
+  PROV-O: The PROV ontology;.
+\newblock Available from: \url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
+
+\end{thebibliography}
 

### 2025-04-23 1_first_review/manuscript_with_tracked_changes.tex
Add response to reviewers, further rename files

--- 2025-04-23 1_first_review/manuscript.tex
@@
 \usepackage{authblk}
-%\usepackage{xcolor} % Load the xcolor package for color options
+\usepackage{xcolor} % Load the xcolor package for color options
 \usepackage{microtype}
@@
 \usepackage{hyperref}
-
+\usepackage{xcolor}
 \usepackage[normalem]{ulem}
 
-%\usepackage[backend=biber]{biblatex}
+%\usepackage{makecell} % multiline table cells
 
@@
 Such studies are typically published as either `methods-development papers' (MDPs) where new methods are compared against existing ones; or `benchmark-only papers' (BOPs), where a set of existing methods are compared in a more neutral way, as defined in \cite{cao2023-jz}.
-A benchmarking study can be run by a single person  (solo benchmarking) or a small group, or it can be organized as a `challenge' or within a hackathon.
+A benchmarking study can be run by a single person {\color{red} (solo benchmarking)} or a small group, or it can be organized as a `challenge' or within a hackathon.
 In this work, we present a perspective on building a robust benchmarking ecosystem in bioinformatics, coming primarily from the methodologist's viewpoint (researchers developing computational tools).
@@
 
-In our discussion, we outline concepts related to workflow automation, benchmark formalization and benchmarking-specific tasks beyond the former, including philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization.
+In our discussion, we outline concepts {\color{red} related to workflow automation, benchmark formalization and benchmarking-specific tasks beyond the former, including \sout{and specifications for an ideal benchmarking system; we examine}} philosophical issues like gatekeeping and trust; and provide some technical recommendations, such as standardization.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/fig1_fig_benchmarking_aligned.pdf}
-     \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
+    \includegraphics[width=\linewidth]{./figures/fig1_fig_benchmarking_aligned.pdf}
+    {\color{red} \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
 These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e.
@@
 \label{fig:definitions}}
-
+    }
 \end{figure}
@@
 We define a benchmark as a conceptual framework to evaluate the performance of computational methods for a given task, though it can extend to assessing other aspects, like data-generating technologies.
-A benchmark thus requires a well-defined task and typically a definition of correctness, or ground-truth, to be precisely defined in advance  (Figure~\ref{fig:definitions}).
+A benchmark thus requires a well-defined task and typically a definition of correctness, or ground-truth, to be precisely defined in advance {\color{red} (Figure~\ref{fig:definitions})}.
 A benchmark then consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, and metrics.
@@
 Can benchmarking be systematized?
-Benchmarking aims to calculate, collect and report performance metrics of methods aiming to solve a task.
+{\color{red}Benchmarking aims to calculate, collect and report performance metrics of methods aiming to solve a task.
 Such a process normally requires collecting appropriate data and running methods in an automated manner.
@@
 Moreover, scientific benchmarking borrows procedures from the scientific method and from scientific publishing, hence having expectations of fairness, reproducibility, transparency, and trust.
-\emph{Benchmarking systems} aim to orchestrate the workflow management, benchmarking and community engagement aspects of the process to generate benchmark `artifacts' (code snapshots, file outputs to be shared, performance outputs) in a systematic way and following good practices and standards.
-Namely,  the various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
-Such systems should facilitate components to be public, open and allow contributions (e.g., corrections or new components), such that the benchmarks of the future can run more continuously.
+\sout{Benchmarks should be orchestrated and run by}} \emph{Benchmarking systems} {\color{red} aim to orchestrate the workflow management, benchmarking and community engagement aspects of the process  \sout{to create} to generate benchmark} `artifacts' (code snapshots, file outputs to be shared, performance outputs) {\color{red} in a systematic way and following good practices and standards.
+Namely,}  the various elements of code, especially the details on how datasets are preprocessed and how methods are run, should be explicitly available for scrutiny.
+Such {\color{red}\sout{a}benchmarking} systems should facilitate components to be public, open and allow contributions (e.g., corrections or new components), such that the benchmarks of the future can run more continuously.
 At the outset of a benchmark, a benchmark `definition' could be envisioned, which could give a formal specification of the entire set of components (and the pattern of artifacts to be generated).
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figures/fig2_bens_layers.pdf}
+    \includegraphics[width=0.8\linewidth]{./figures/fig2_bens_layers.pdf}
     \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges.
@@
 Journals and funding bodies also have a vested interest in ensuring that published or funded method developments are performed to a high standard, that unnecessary redundancy is reduced, and that results are FAIR (findable, accessible, interoperable and reusable) for maximal benefit to the community \cite{Wilkinson2016-bh}.
-Recently published benchmarks and challenges have been conceived as or ported to benchmarking systems to ensure these properties \cite{Nevers_2022,Bryce_Smith_2023, Pardo_Palacios_2023}.
-Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
+{\color{red} Recently published benchmarks and challenges have been conceived as or ported to benchmarking systems to ensure these properties \cite{Nevers_2022,Bryce_Smith_2023, Pardo_Palacios_2023}.} Thus, quality assurance, neutrality and transparency of benchmarking systems become highly relevant.
 
-Collaborative benchmarking efforts enunciate field-specific challenges and provide a venue for \emph{communities} to address them.
-For instance, DREAM \cite{meyer2021advances}, GIAB \cite{zook2019open} and CASP \cite{moult2005-ne} challenges and benchmarks have demonstrated how collaborative benchmarks can help in moving fields forward.
+{\color{red} Collaborative benchmarking efforts enunciate field-specific challenges and provide a venue for \emph{communities} to address them.
+For instance, DREAM \cite{meyer2021advances}, GIAB \cite{zook2019open} and CASP \cite{moult2005-ne} challenges and benchmarks have demonstrated how collaborative benchmarks can help in moving fields forward.}
 
@@
         %\includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
-    \includegraphics[width=0.75\linewidth]{./figures/fig3_bens_formalization_arrows.pdf} % alternative
+    \includegraphics[width=0.75\linewidth]{figures/fig3_bens_formalization_arrows.pdf} % alternative
     \caption[Benchmark formalization.]{Benchmark formalization.
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack (the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments'); managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results (Figure~\ref{fig:definitions}).
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack {\color{red}(the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments')}; managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results {\color{red}(Figure~\ref{fig:definitions})}.
 
@@
 
+{\color{red}
 Aside of task-tailored scores (i.e., intrinsic or extrinsic cluster validity measures for clustering tasks), collecting metrics to report computing performance (i.e., wallclock, on-disk read-write operations), requirements (e.g., GPUs, minimum memory) and their scalability depending on the input size (e.g., whether runtimes increase linearly, quadratically, etc) is key during benchmarking.
 To measure these performance metrics accurately, benchmarking systems ideally would provide means to: i) characterize the hardware running the benchmark, so method runtimes are comparable;  ii) track performance (CPU or GPU usage, memory) when running each benchmark component; iii) introduce minimal overhead during profiling; iv) collect variability in performance by running each method several times; and, v) go beyond reporting average resource usage and provide them over execution time (e.g., so embarrassingly-parallel algorithms can be differentiated from largely single-cored runs with few bursts of parallel execution).
+}
 
@@
 
+{\color{red}
 Mature communities are key to sustaining and reflecting the existing challenges in a given field.
@@
 To facilitate community contribution, we believe that benchmarking systems can lower the barrier to contributing, hence making the contribution efforts less likely to be biased.
+}
 
@@
 Computational reproducibility comes from controlling the triad of data, code, and environment \cite{Hill2024-gf}.
-During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system (operating system, compiler toolchains, libraries) and its configuration.
+During benchmarking, a workflow runs code transforming inputs into outputs; execution is affected by the base system ({\color{red}\sout{OS} operating system}, compiler toolchains, libraries) and its configuration.
 The benchmark definition should control the execution environment.
@@
 
+{\color{red} 
 The responsibility of handling software environments poses an interesting challenge to benchmarking systems.
@@
 These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; \texttt{conda} is speeding up their deployment procedures (via \texttt{py-rattler}/\texttt{pixi}) and allowing compilation to Web assembly (to execute code within a web browser); and \texttt{apptainer} (former \texttt{Singularity}) provides rootless containerization suitable for running in scientific clusters.
+}
 
@@
 Benchmark contributions should be licensed and associated to scholarly names including email and ORCID (as the canonical identifier), hence allowing seamless reuse, transparency and attribution.
-As a consequence, properly-licensed benchmark components (e.g., a procedure to compute a clustering metric; or a clustering method) could be reused across benchmarks; we envision assisted benchmark design tools to leverage and reuse benchmark components, perhaps guided by AI.
+{\color{red} As a consequence, properly-licensed benchmark components (e.g., a procedure to compute a clustering metric; or a clustering method) could be reused across benchmarks; we envision assisted benchmark design tools to leverage and reuse benchmark components, perhaps guided by AI.}
 
@@
 
+{\color{red}
 \section*{Funding}
@@
 We report no conflict of interest.
-
+}
 \section*{Author contributions}
 
-Conceptualization: all. Figures: IM, BC. Original draft preparation: all. Writing, reviewing and editing: all. Supervision: IM, MDR. Funding acquisition: MDR.
+Conceptualization: all. {\color{red} \sout{Visualization} Figures: IM,} BC. Original draft preparation: all. Writing, reviewing and editing: all. Supervision: IM, MDR. Funding acquisition: MDR.
 

### 2025-07-16 2_second_review/manuscript_with_tracked_changes.tex
more

--- 2025-04-23 1_first_review/manuscript.tex
@@
     \includegraphics[width=\linewidth]{figures/fig1_fig_benchmarking_aligned.pdf}
-     \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
+     \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is {\color{red} \sout{free software}FOSS}, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
 These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e.
@@
 To our knowledge, no benchmarking platform makes use of a fully expressive benchmark definition formalization language covering workflow generation and benchmarking-specific tasks.
-Such a language would go a long way toward facilitating benchmark conceptualization, community gathering and result sharing.
+{\color{red} Orthogonally, some components of the benchmarking process have been formalized: \texttt{Viash} systematizes workflow components; \texttt{PROV-O} represents provenance data; and \texttt{Research Objects} (\texttt{RO} \cite{bechhofer2013linked}, e.g., \texttt{RO-Crate}\cite{soiland2022packaging}) specify a system to store data (files) and their metadata.
+\sout{Such a language would go a long way toward facilitating benchmark conceptualization, community gathering and result sharing} Reusing pre-existing formalizations and extending them to the benchmarking field would facilitate conceptualization, community gathering and result sharing}.
 
@@
 In most systems, code will be tracked in a \texttt{git} repository (e.g., on GitHub, GitLab).
-The workflow execution and the storage of derived files are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery (e.g., GitHub actions for \texttt{ncbench}) or on arbitrary infrastructure (i.e., laptop, server or cloud; e.g., omnibenchmark \cite{omnibenchmark}).
+The workflow execution and the storage of derived files are typically located either in local deployments (e.g., \texttt{OEB-VRE} for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery (e.g., GitHub actions for \texttt{ncbench}) or on arbitrary infrastructure (i.e., laptop, server or cloud; e.g., omnibenchmark \cite{omnibenchmark}).
 
@@
 However, becoming centralized is not cost-free: it implies gatekeeping (contributors authenticate to the system and get access to computational resources) and requires additional measures that build trust (e.g., transparency).
+
+{\color{red}Large benchmarks are often run using cloud computing, further extending the challenge of ensuring hardware homogeneity (so results can be compared) and adding complexity to the benchmarking system capabilities to automate the provisioning and to group executions so costs are minimized.
+Automating cloud provisioning and constraining hardware flavors contribute to ensuring hardware homogeneity.}
 
@@
 Contributions can be made easier by providing good organization, suggested `good first issues' to tackle, contributor guidelines and a code of conduct.
+
+{\color{red} Authorship, e.g.
+acknowledging contributions, can be captured, reported and credited by using third-party platforms such as Apicuron \cite{hatos2021apicuron} or by bundling standardized contribution metadata such as the Citation File Format (\texttt{CFF}).
+So individual contributions to a benchmark can be acknowledged, browsed, and ideally propagated to academic deliverables (e.g.
+datasets, publications).}
 
@@
 
-Similarly, even if all benchmark code were free software, running it could be hindered by (non-free) proprietary interpreters (e.g., MATLAB).
+Similarly, even if all benchmark code were {\color{red} \sout{free} free and open source (FOSS)} software, running it could be hindered by (non-free) proprietary interpreters (e.g., MATLAB).
 Benchmarking systems should be explicit in allowing or disallowing proprietary software, and be aware and track EULAs (end user license agreements).
 This conundrum extends to the operating system layer and to system dependencies (e.g., compilers).
-In terms of benchmark re-runnability, fully free software system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{droge2023-ax}.
+In terms of benchmark re-runnability, fully {\color{red} \sout{free software} FOSS} system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{droge2023-ax}.
 
@@
 A benchmarking system should automate license metadata handling during data import and export, crediting the original authors if applicable; it also represents an opportunity to enforce that academics use appropriate licenses for data and code.
+
+{\color{red} \section{Hardware}
+
+Benchmarking systems are software meant to orchestrate other software (e.g.
+benchmarked methods).
+Software are run on specific hardware, from commodity computers to dedicated hardware, often large servers or academic and commercial clouds, posing a challenge: benchmarked methods should have access to comparable resources (hardware).
+
+compilation flags
+
+multiarch, ARMs
+
+containers, discuss software then
+
+computing quotas, concurrency, IO
+
+how to profile performance and fingerprint resources
+}
 
@@
 
-\bibliographystyle{vancouver} % elsarticle-num} % vancouver}  
-\bibliography{clean} % your .bib file
+%\bibliographystyle{vancouver} % elsarticle-num} % vancouver}  
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+\bibliography{2_second_review/src_do_not_upload/clean} % your .bib file
+
+% \begin{thebibliography}{10}
+
+%\end{thebibliography}
 

### 2025-07-17 2_second_review/manuscript_with_tracked_changes.tex
first round completed

@@
 
+{\color{red} Transparency and clear governance guidelines provide an effective to detect and react to potential misuse, e.g.
+by explicit rules to validate results.
+Other ethical considerations include facilitating safe interactions and fair resources allocation and efficiency.
+Clear code of conduct statements proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) provide explicit means to ease collaboration.
+In terms of efficiency, comparable resources allocation (e.g., computing power, support) to all contribuors, while also being aware of $CO_2$ emissions, could also be spelled out in the governance model.}
+
 Mature communities are key to sustaining and reflecting the existing challenges in a given field.
@@
 
+{\color{red} \section{Hardware}
+
+Software, both from benchmarking systems or from the benchmarking workflow and hence benchmarked methods, can be run in different hardware, from commodity computers to dedicated hardware, often large servers or academic and commercial clouds.
+Performance profiling (e.g., how fast a benchmarked method runs) requires access to fair and comparable resources.
+
+Comparable resources can be provisioned on demand in cloud computing, or can be ensured by deploying benchmarking systems in controlled and dedicated hardware.
+Ideally, performance results (e.g., wallclock time) could be bundled to metadata describing the used hardware (e.g., CPU characteristics).
+
+It is worth noting that CPU microarchitecture influences performance, particularly when using optimized compilers to improve performance of benchmarked methods at the expense of making binaries microarchitecture-specific.
+This feature is not solved by containerization.
+Benchmarking systems can automate software installations with equivalent degrees of optimization across benchmarked methods, and/or provide multiarchitecture builds.
+
+Finally, benchmarking methods could benefit from the high performance computing-style user quotas and timeouts to cap the maximum memory and/or CPU cycles each method can use.}
+
 \section*{Software licenses and attribution}
@@
 A benchmarking system should automate license metadata handling during data import and export, crediting the original authors if applicable; it also represents an opportunity to enforce that academics use appropriate licenses for data and code.
-
-{\color{red} \section{Hardware}
-
-Benchmarking systems are software meant to orchestrate other software (e.g.
-benchmarked methods).
-Software are run on specific hardware, from commodity computers to dedicated hardware, often large servers or academic and commercial clouds, posing a challenge: benchmarked methods should have access to comparable resources (hardware).
-
-compilation flags
-
-multiarch, ARMs
-
-containers, discuss software then
-
-computing quotas, concurrency, IO
-
-how to profile performance and fingerprint resources
-}
 

### 2025-07-17 2_second_review/manuscript_with_tracked_changes.tex
Accept Mark's changes

@@
 \texttt{OpenEBench} runs workflows in any workflow language, frequently \texttt{nextflow}, and makes use of \texttt{openVRE} as GUI and means of visualization \cite{Capella-Gutierrez2017-dh}.
-\texttt{OpenProblems} uses \texttt{Viash} and \texttt{nextflow} to handle software environments and workflows, respectively, and reports updated leaderboards with custom code \cite{Luecken2024-fk}.
+\texttt{OpenProblems} uses \texttt{Viash} \cite{cannoodt2024viash} and \texttt{nextflow} to handle software environments and workflows, respectively, and reports updated leaderboards with custom code \cite{luecken2025open}.
 To our knowledge, no benchmarking platform makes use of a fully expressive benchmark definition formalization language covering workflow generation and benchmarking-specific tasks.
-{\color{red} Orthogonally, some components of the benchmarking process have been formalized: \texttt{Viash} systematizes workflow components; \texttt{PROV-O} represents provenance data; and \texttt{Research Objects} (\texttt{RO} \cite{bechhofer2013linked}, e.g., \texttt{RO-Crate}\cite{soiland2022packaging}) specify a system to store data (files) and their metadata.
+{\color{red} Orthogonally, some components of the benchmarking process have been formalized: \texttt{Viash} \cite{cannoodt2024viash} systematizes workflow components; \texttt{PROV-O} represents provenance data \cite{LeboUnknown-om}; and \texttt{Research Objects} (\texttt{RO} \cite{bechhofer2013linked}, e.g., \texttt{RO-Crate} \cite{soiland2022packaging}) specify a system to store data (files) and their metadata.
 \sout{Such a language would go a long way toward facilitating benchmark conceptualization, community gathering and result sharing} Reusing pre-existing formalizations and extending them to the benchmarking field would facilitate conceptualization, community gathering and result sharing}.
@@
 However, because there are not many standards of how benchmarks are implemented \cite{Sonrel2023-te}, another disadvantage of decentralized benchmarks is that, while these components may eventually be openly shared as a product of a research study, the contents of a benchmark are non-interoperable, even across benchmarks of the same task.
-Furthermore, decentralized benchmarks have been criticized on the basis of `inflating model performance' due to decisions made in preprocessing or parameter tuning \cite{Luecken2024-fk}; such issues could be mitigated by gathering the `wisdom of crowds' that is implicit in running centralized benchmarks.
+Furthermore, decentralized benchmarks have been criticized on the basis of `inflating model performance' due to decisions made in preprocessing or parameter tuning \cite{luecken2025open}; such issues could be mitigated by gathering the `wisdom of crowds' that is implicit in running centralized benchmarks.
 
@@
 
-{\color{red}Large benchmarks are often run using cloud computing, further extending the challenge of ensuring hardware homogeneity (so results can be compared) and adding complexity to the benchmarking system capabilities to automate the provisioning and to group executions so costs are minimized.
-Automating cloud provisioning and constraining hardware flavors contribute to ensuring hardware homogeneity.}
+{\color{red}Compute-intensive benchmarks are often run using cloud or high performance computing, further extending the challenge to ensure hardware homogeneity (for comparable performances) and adding complexity to the benchmarking system capabilities to automate node provisioning (or node selection) and to group executions such that costs are minimized.
+Some workflow managers, such as \texttt{Snakemake}, provide plugins to defer execution to cloud computing or queuing systems, effectively decoupling workflow specification and execution.}
 
@@
 
-{\color{red} Authorship, e.g.
-acknowledging contributions, can be captured, reported and credited by using third-party platforms such as Apicuron \cite{hatos2021apicuron} or by bundling standardized contribution metadata such as the Citation File Format (\texttt{CFF}).
-So individual contributions to a benchmark can be acknowledged, browsed, and ideally propagated to academic deliverables (e.g.
-datasets, publications).}
+{\color{red} Authorship, e.g., acknowledging contributions, can be captured, reported and credited using third-party platforms such as Apicuron \cite{hatos2021apicuron} or by bundling standardized contribution metadata such as the Citation File Format (\texttt{CFF}).
+In this way, individual contributions to a benchmark can be acknowledged, browsed, and ideally propagated to academic deliverables (e.g., datasets, publications).}
 
@@
 
-{\color{red} Transparency and clear governance guidelines provide an effective to detect and react to potential misuse, e.g.
-by explicit rules to validate results.
-Other ethical considerations include facilitating safe interactions and fair resources allocation and efficiency.
-Clear code of conduct statements proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) provide explicit means to ease collaboration.
-In terms of efficiency, comparable resources allocation (e.g., computing power, support) to all contribuors, while also being aware of $CO_2$ emissions, could also be spelled out in the governance model.}
+{\color{red} Transparency and clear governance guidelines provide an effective means to detect and react to potential misuse, e.g., by explicit rules to validate results.
+Other ethical considerations include facilitating safe interactions and fair resource allocation and efficiency.
+Clear code of conduct statements proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) and provide an explicit means to ease collaboration.
+In terms of efficiency, comparable allocation of resources (e.g., computing power, support) to all contributors, while also being aware of $CO_2$ emissions, could also be spelled out in the governance model.}
 
@@
 
-{\color{red} \section{Hardware}
+{\color{red} \section*{Hardware}
 
@@
 This feature is not solved by containerization.
-Benchmarking systems can automate software installations with equivalent degrees of optimization across benchmarked methods, and/or provide multiarchitecture builds.
+Benchmarking systems can automate software installations with equivalent degrees of optimization across benchmarked methods, and/or provide multi-architecture builds.
 
-Finally, benchmarking methods could benefit from the high performance computing-style user quotas and timeouts to cap the maximum memory and/or CPU cycles each method can use.}
+Finally, benchmarking methods could benefit from high performance computing-style user quotas and timeouts to cap the maximum memory and/or CPU cycles each method can use.}
 
@@
 This expertise can be leveraged by current practices of community benchmarking, so scientific advances are eased by joint efforts to catalog open problems, generate (or simulate) appropriate input data, develop methods to address them, and evaluate their performances.
-Community benchmarking has been proven successful and  productive \cite{moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.
+Community benchmarking has been proven successful and  productive \cite{moult2005-ne,Capella-Gutierrez2017-dh,luecken2025open}}.
 
@@
 
-\bibliography{2_second_review/src_do_not_upload/clean} % your .bib file
+% to-do replace by bibitems here
+\bibliography{2_second_review/src_do_not_upload/clean}
 

### 2025-07-23 2_second_review/manuscript_with_tracked_changes.tex
accept all changes, before replacing to bibitems

@@
 \author[2,3]{Charlotte Soneson}
-\author[1]{Ben Carrillo}
+\author[1,2]{Ben Carrillo}
 \author[1,2,4]{Almut Lütge}
@@
 Currently, most benchmarking systems aim to enhance the workflow formalization and execution with strategies to fulfill benchmarking-specific needs.
-To name a few of these systems, \texttt{ncbench} bundles workflow specification and software management with \texttt{snakemake} to benchmark result visualization with \texttt{Datavzrd} \cite{Hanssen2023-uf}.
+To name a few of these systems, \texttt{ncbench} bundles workflow specification and software management with \texttt{Snakemake} to benchmark result visualization with \texttt{Datavzrd} \cite{Hanssen2023-uf}.
 \texttt{OpenEBench} runs workflows in any workflow language, frequently \texttt{nextflow}, and makes use of \texttt{openVRE} as GUI and means of visualization \cite{Capella-Gutierrez2017-dh}.
@@
 {\color{red}Compute-intensive benchmarks are often run using cloud or high performance computing, further extending the challenge to ensure hardware homogeneity (for comparable performances) and adding complexity to the benchmarking system capabilities to automate node provisioning (or node selection) and to group executions such that costs are minimized.
-Some workflow managers, such as \texttt{Snakemake}, provide plugins to defer execution to cloud computing or queuing systems, effectively decoupling workflow specification and execution.}
+Some workflow managers, such as \texttt{Snakemake}, provide plugins to execute on the cloud or on-premises queuing systems, effectively decoupling workflow specification and execution.}
 
@@
 
-{\color{red} Authorship, e.g., acknowledging contributions, can be captured, reported and credited using third-party platforms such as Apicuron \cite{hatos2021apicuron} or by bundling standardized contribution metadata such as the Citation File Format (\texttt{CFF}).
+{\color{red} Authorship, e.g., acknowledging contributions, can be captured, reported and credited using third-party platforms such as Apicuron \cite{hatos2021apicuron} or by bundling standardized contribution metadata such as the Citation File Format (\texttt{CFF}) \cite{druskat_2021_5171937}.
 In this way, individual contributions to a benchmark can be acknowledged, browsed, and ideally propagated to academic deliverables (e.g., datasets, publications).}
@@
 Other ethical considerations include facilitating safe interactions and fair resource allocation and efficiency.
-Clear code of conduct statements proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) and provide an explicit means to ease collaboration.
+Clear code of conduct statements proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) and provide explicit means to ease collaboration.
 In terms of efficiency, comparable allocation of resources (e.g., computing power, support) to all contributors, while also being aware of $CO_2$ emissions, could also be spelled out in the governance model.}
@@
 
-Software, both from benchmarking systems or from the benchmarking workflow and hence benchmarked methods, can be run in different hardware, from commodity computers to dedicated hardware, often large servers or academic and commercial clouds.
+Software, both from benchmarking systems and from the benchmarking workflow itself (the benchmarked methods), can be run on different hardware, from commodity computers to dedicated hardware, often large servers or academic and commercial clouds.
 Performance profiling (e.g., how fast a benchmarked method runs) requires access to fair and comparable resources.
@@
 This expertise can be leveraged by current practices of community benchmarking, so scientific advances are eased by joint efforts to catalog open problems, generate (or simulate) appropriate input data, develop methods to address them, and evaluate their performances.
-Community benchmarking has been proven successful and  productive \cite{moult2005-ne,Capella-Gutierrez2017-dh,luecken2025open}}.
+Community benchmarking has been proven successful and  productive \cite{moult2005-ne,Capella-Gutierrez2017-dh,luecken2025open}.
 

### 2025-07-23 2_second_review/manuscript.tex
Untrack changes, add bibs as bibitems

--- 2025-04-23 1_first_review/manuscript.tex
@@
 
-%\usepackage[backend=biber]{biblatex}
+% \usepackage[backend=biber]{biblatex}
 
@@
     \includegraphics[width=\linewidth]{figures/fig1_fig_benchmarking_aligned.pdf}
-     \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is free software, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
+     \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is FOSS, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
 These run-intrinsic properties have consequences in terms of features (i.e., transparency, forkability, availability for metaanalysis) and technical capabilities (i.e.
@@
 Currently, most benchmarking systems aim to enhance the workflow formalization and execution with strategies to fulfill benchmarking-specific needs.
-To name a few of these systems, \texttt{ncbench} bundles workflow specification and software management with \texttt{snakemake} to benchmark result visualization with \texttt{Datavzrd} \cite{Hanssen2023-uf}.
+To name a few of these systems, \texttt{ncbench} bundles workflow specification and software management with \texttt{Snakemake} to benchmark result visualization with \texttt{Datavzrd} \cite{Hanssen2023-uf}.
 \texttt{OpenEBench} runs workflows in any workflow language, frequently \texttt{nextflow}, and makes use of \texttt{openVRE} as GUI and means of visualization \cite{Capella-Gutierrez2017-dh}.
-\texttt{OpenProblems} uses \texttt{Viash} and \texttt{nextflow} to handle software environments and workflows, respectively, and reports updated leaderboards with custom code \cite{Luecken2024-fk}.
+\texttt{OpenProblems} uses \texttt{Viash} \cite{cannoodt2024viash} and \texttt{nextflow} to handle software environments and workflows, respectively, and reports updated leaderboards with custom code \cite{luecken2025open}.
 To our knowledge, no benchmarking platform makes use of a fully expressive benchmark definition formalization language covering workflow generation and benchmarking-specific tasks.
-Such a language would go a long way toward facilitating benchmark conceptualization, community gathering and result sharing.
+Orthogonally, some components of the benchmarking process have been formalized: \texttt{Viash} \cite{cannoodt2024viash} systematizes workflow components; \texttt{PROV-O} represents provenance data \cite{LeboUnknown-om}; and \texttt{Research Objects} (\texttt{RO} \cite{bechhofer2013linked}, e.g., \texttt{RO-Crate} \cite{soiland2022packaging}) specify a system to store data (files) and their metadata.
+Reusing pre-existing formalizations and extending them to the benchmarking field would facilitate conceptualization, community gathering and result sharing.
 
@@
 However, becoming centralized is not cost-free: it implies gatekeeping (contributors authenticate to the system and get access to computational resources) and requires additional measures that build trust (e.g., transparency).
+
+Compute-intensive benchmarks are often run using cloud or high performance computing, further extending the challenge to ensure hardware homogeneity (for comparable performances) and adding complexity to the benchmarking system capabilities to automate node provisioning (or node selection) and to group executions such that costs are minimized.
+Some workflow managers, such as \texttt{Snakemake}, provide plugins to execute on the cloud or on-premises queuing systems, effectively decoupling workflow specification and execution.
 
@@
 Contributions can be made easier by providing good organization, suggested `good first issues' to tackle, contributor guidelines and a code of conduct.
+
+Authorship, e.g., acknowledging contributions, can be captured, reported and credited using third-party platforms such as Apicuron \cite{hatos2021apicuron} or by bundling standardized contribution metadata such as the Citation File Format (\texttt{CFF}) \cite{druskat_2021_5171937}.
+In this way, individual contributions to a benchmark can be acknowledged, browsed, and ideally propagated to academic deliverables (e.g., datasets, publications).
 
@@
 Templates exist for simulation studies that could be adapted more generally for benchmarking projects \cite{Siepe2023-tz}.
+
+Transparency and clear governance guidelines provide an effective means to detect and react to potential misuse, e.g., by explicit rules to validate results.
+Other ethical considerations include facilitating safe interactions and fair resource allocation and efficiency.
+Clear code of conduct statements proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) and provide explicit means to ease collaboration.
+In terms of efficiency, comparable allocation of resources (e.g., computing power, support) to all contributors, while also being aware of $CO_2$ emissions, could also be spelled out in the governance model.
 
@@
 
+\section*{Hardware}
+
+Software, both from benchmarking systems and from the benchmarking workflow itself (the benchmarked methods), can be run on different hardware, from commodity computers to dedicated hardware, often large servers or academic and commercial clouds.
+Performance profiling (e.g., how fast a benchmarked method runs) requires access to fair and comparable resources.
+
+Comparable resources can be provisioned on demand in cloud computing, or can be ensured by deploying benchmarking systems in controlled and dedicated hardware.
+Ideally, performance results (e.g., wallclock time) could be bundled to metadata describing the used hardware (e.g., CPU characteristics).
+
+It is worth noting that CPU microarchitecture influences performance, particularly when using optimized compilers to improve performance of benchmarked methods at the expense of making binaries microarchitecture-specific.
+This feature is not solved by containerization.
+Benchmarking systems can automate software installations with equivalent degrees of optimization across benchmarked methods, and/or provide multi-architecture builds.
+
+Finally, benchmarking methods could benefit from high performance computing-style user quotas and timeouts to cap the maximum memory and/or CPU cycles each method can use.
+
 \section*{Software licenses and attribution}
@@
 
-Similarly, even if all benchmark code were free software, running it could be hindered by (non-free) proprietary interpreters (e.g., MATLAB).
+Similarly, even if all benchmark code were free and open source (FOSS) software, running it could be hindered by (non-free) proprietary interpreters (e.g., MATLAB).
 Benchmarking systems should be explicit in allowing or disallowing proprietary software, and be aware and track EULAs (end user license agreements).
 This conundrum extends to the operating system layer and to system dependencies (e.g., compilers).
-In terms of benchmark re-runnability, fully free software system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{droge2023-ax}.
+In terms of benchmark re-runnability, fully FOSS system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{droge2023-ax}.
 
@@
 This expertise can be leveraged by current practices of community benchmarking, so scientific advances are eased by joint efforts to catalog open problems, generate (or simulate) appropriate input data, develop methods to address them, and evaluate their performances.
-Community benchmarking has been proven successful and  productive \cite{moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.
+Community benchmarking has been proven successful and  productive \cite{moult2005-ne,Capella-Gutierrez2017-dh,luecken2025open}.
 
@@
 
-% \bibliography{clean} % your .bib file
+%\bibliography{src_do_not_upload/clean}
 
@@
 
-\bibitem{Luecken2024-fk}
+\bibitem{cannoodt2024viash}
+Cannoodt R, Cannoodt H, Schaumont D, Waldrant K, Van~de Kerckhove E, Boschmans
+  A, et~al.
+\newblock Viash: A meta-framework for building reusable workflow modules.
+\newblock Journal of Open Source Software. 2024;9(93):6089.
+
+\bibitem{luecken2025open}
 Luecken MD, Gigante S, Burkhardt DB, Cannoodt R, Strobl DC, Markov NS, et~al.
 \newblock Defining and benchmarking open problems in single-cell analysis.
-\newblock Res Sq. 2024;.
+\newblock Nature Biotechnology. 2025;43(7):1035--1040.
+\newblock doi:{10.1038/s41587-025-02694-w}.
+
+\bibitem{LeboUnknown-om}
+Lebo T, Sahoo S, McGuinness D, Belhajjame K, Cheney J, Corsar D, et~al..
+  PROV-O: The PROV ontology;.
+\newblock Available from: \url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
+
+\bibitem{bechhofer2013linked}
+Bechhofer S, Buchan I, De~Roure D, Missier P, Ainsworth J, Bhagat J, et~al.
+\newblock Why linked data is not enough for scientists.
+\newblock Future Generation Computer Systems. 2013;29(2):599--611.
+
+\bibitem{soiland2022packaging}
+Soiland-Reyes S, Sefton P, Crosas M, Castro LJ, Coppens F, Fern{\'a}ndez JM,
+  et~al.
+\newblock Packaging research artefacts with RO-Crate.
+\newblock Data Science. 2022;5(2):97--138.
 
@@
 \newblock arXiv preprint arXiv:240917038. 2024;.
+
+\bibitem{hatos2021apicuron}
+Hatos A, Quaglia F, Piovesan D, Tosatto SC.
+\newblock APICURON: a database to credit and acknowledge the work of
+  biocurators.
+\newblock Database. 2021;2021:baab019.
+
+\bibitem{druskat_2021_5171937}
+Druskat S, Spaaks JH, Chue~Hong N, Haines R, Baker J, Bliven S, et~al..
+  Citation File Format; 2021.
+\newblock Available from: \url{https://doi.org/10.5281/zenodo.5171937}.
 
@@
 
-\bibitem{LeboUnknown-om}
-Lebo T, Sahoo S, McGuinness D, Belhajjame K, Cheney J, Corsar D, et~al..
-  PROV-O: The PROV ontology;.
-\newblock Available from: \url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
-
 \end{thebibliography}

### 2025-07-23 2_second_review/manuscript_with_tracked_changes.tex
Add bbls to manuscript with tracked changes too, so no bibfile is needed

@@
 
-% to-do replace by bibitems here
-\bibliography{2_second_review/src_do_not_upload/clean}
+%\bibliography{2_second_review/src_do_not_upload/clean}
 
@@
 
-%\end{thebibliography}
+\begin{thebibliography}{10}
+
+\bibitem{cao2023-jz}
+Cao Y, Yu L, Torkel M, Kim S, Lin Y, Yang P, et~al.
+\newblock The current landscape and emerging challenges of benchmarking
+  single-cell methods.
+\newblock bioRxiv. 2023; p. 2023.12.19.572303.
+
+\bibitem{Sonrel2023-te}
+Sonrel A, Luetge A, Soneson C, Mallona I, Germain PL, Knyazev S, et~al.
+\newblock Meta-analysis of (single-cell method) benchmarks reveals the need for
+  extensibility and interoperability.
+\newblock Genome Biol. 2023;24:119.
+
+\bibitem{Boulesteix2013-vy}
+Boulesteix AL, Lauer S, Eugster MJA.
+\newblock A plea for neutral comparison studies in computational sciences.
+\newblock PLoS One. 2013;8:e61562.
+
+\bibitem{Weber2019-el}
+Weber LM, Saelens W, Cannoodt R, Soneson C, Hapfelmeier A, Gardner PP, et~al.
+\newblock Essential guidelines for computational method benchmarking.
+\newblock Genome Biol. 2019;20:125.
+
+\bibitem{Brooks2024-rq}
+Brooks TG, Lahens NF, Mrčela A, Grant GR.
+\newblock Challenges and best practices in omics benchmarking.
+\newblock Nature Review Genetics. 2024;25(5):326--339.
+
+\bibitem{Rich2024-jh}
+Rich JM, Moses L, Einarsson PH, Jackson K, Luebbert L, Booeshaghi AS, et~al.
+\newblock The impact of package selection and versioning on single-cell
+  {RNA}-seq analysis.
+\newblock bioRxivorg. 2024; p. 2024.04.04.588111.
+
+\bibitem{Wilkinson2016-bh}
+Wilkinson MD, Dumontier M, Aalbersberg IJJ, Appleton G, Axton M, Baak A, et~al.
+\newblock The {FAIR} Guiding Principles for scientific data management and
+  stewardship.
+\newblock Sci Data. 2016;3:160018.
+
+\bibitem{Nevers_2022}
+Nevers Y, Jones TEM, Jyothi D, Yates B, Ferret M, Portell-Silva L, et~al.
+\newblock The Quest for Orthologs orthology benchmark service in 2022.
+\newblock Nucleic Acids Research. 2022;50(W1):W623–W632.
+\newblock doi:{10.1093/nar/gkac330}.
+
+\bibitem{Bryce_Smith_2023}
+Bryce-Smith S, Burri D, Gazzara MR, Herrmann CJ, Danecka W, Fitzsimmons CM,
+  et~al.
+\newblock Extensible benchmarking of methods that identify and quantify
+  polyadenylation sites from RNA-seq data.
+  2023;doi:{10.1101/2023.06.23.546284}.
+
+\bibitem{Pardo_Palacios_2023}
+Pardo-Palacios FJ, Wang D, Reese F, Diekhans M, Carbonell-Sala S, Williams B,
+  et~al.
+\newblock Systematic assessment of long-read RNA-seq methods for transcript
+  identification and quantification. 2023;doi:{10.1101/2023.07.25.550582}.
+
+\bibitem{meyer2021advances}
+Meyer P, Saez-Rodriguez J.
+\newblock Advances in systems biology modeling: 10 years of crowdsourcing DREAM
+  challenges.
+\newblock Cell Systems. 2021;12(6):636--653.
+
+\bibitem{zook2019open}
+Zook JM, McDaniel J, Olson ND, Wagner J, Parikh H, Heaton H, et~al.
+\newblock An open resource for accurately benchmarking small variant and
+  reference calls.
+\newblock Nature Biotechnology. 2019;37(5):561--566.
+
+\bibitem{moult2005-ne}
+Moult J.
+\newblock A decade of {CASP}: progress, bottlenecks and prognosis in protein
+  structure prediction.
+\newblock Curr Opin Struct Biol. 2005;15:285--289.
+
+\bibitem{Wratten2021-el}
+Wratten L, Wilm A, Göke J.
+\newblock Reproducible, scalable, and shareable analysis pipelines with
+  bioinformatics workflow managers.
+\newblock Nat Methods. 2021;18:1161--1168.
+
+\bibitem{Amstutz2024-qk}
+Amstutz P, Mikheev M, Crusoe MR, Tijanić N, Lampa S. Existing Workflow
+  systems; 2024.
+\newblock Available from: \url{https://s.apache.org/existing-workflow-systems}.
+
+\bibitem{Amstutz2016-vo}
+Amstutz P, Crusoe MR, Tijanić N, Chapman B, Chilton J, Heuer M, et~al.. Common
+  Workflow Language, {v1}.0; 2016.
+\newblock Available from:
+  \url{https://research.manchester.ac.uk/files/57032693/cwl_1.0_workflow.pdf}.
+
+\bibitem{Goble2020-ps}
+Goble C, Cohen-Boulakia S, Soiland-Reyes S, Garijo D, Gil Y, Crusoe MR, et~al.
+\newblock FAIR computational workflows.
+\newblock Data Intell. 2020;2:108--121.
+
+\bibitem{Hanssen2023-uf}
+Hanssen F, Gabernet G, Smith NH, Mertes C, Neogi AG, Brandhoff L, et~al.
+\newblock NCBench: providing an open, reproducible, transparent, adaptable, and
+  continuous benchmark approach for {DNA}-sequencing-based variant calling.
+\newblock F1000Research. 2023;12:1125.
+
+\bibitem{Capella-Gutierrez2017-dh}
+Capella-Gutierrez S, de~la Iglesia D, Haas J, Lourenco A, Fernández JM,
+  Repchevsky D, et~al.
+\newblock Lessons Learned: Recommendations for Establishing Critical Periodic
+  Scientific Benchmarking.
+\newblock bioRxiv. 2017; p. 181677.
+
+\bibitem{cannoodt2024viash}
+Cannoodt R, Cannoodt H, Schaumont D, Waldrant K, Van~de Kerckhove E, Boschmans
+  A, et~al.
+\newblock Viash: A meta-framework for building reusable workflow modules.
+\newblock Journal of Open Source Software. 2024;9(93):6089.
+
+\bibitem{luecken2025open}
+Luecken MD, Gigante S, Burkhardt DB, Cannoodt R, Strobl DC, Markov NS, et~al.
+\newblock Defining and benchmarking open problems in single-cell analysis.
+\newblock Nature Biotechnology. 2025;43(7):1035--1040.
+\newblock doi:{10.1038/s41587-025-02694-w}.
+
+\bibitem{LeboUnknown-om}
+Lebo T, Sahoo S, McGuinness D, Belhajjame K, Cheney J, Corsar D, et~al..
+  PROV-O: The PROV ontology;.
+\newblock Available from: \url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
+
+\bibitem{bechhofer2013linked}
+Bechhofer S, Buchan I, De~Roure D, Missier P, Ainsworth J, Bhagat J, et~al.
+\newblock Why linked data is not enough for scientists.
+\newblock Future Generation Computer Systems. 2013;29(2):599--611.
+
+\bibitem{soiland2022packaging}
+Soiland-Reyes S, Sefton P, Crosas M, Castro LJ, Coppens F, Fern{\'a}ndez JM,
+  et~al.
+\newblock Packaging research artefacts with RO-Crate.
+\newblock Data Science. 2022;5(2):97--138.
+
+\bibitem{Khan2018-ot}
+Khan RU, Zhang X, Kumar R, Aboagye EO.
+\newblock Evaluating the performance of {ResNet} model based on image
+  recognition.
+\newblock In: Proceedings of the 2018 International Conference on Computing and
+  Artificial Intelligence. ACM; 2018.Available from:
+  \url{https://dl.acm.org/doi/10.1145/3194452.3194461}.
+
+\bibitem{Jumper2021-oq}
+Jumper J, Evans R, Pritzel A, Green T, Figurnov M, Ronneberger O, et~al.
+\newblock Highly accurate protein structure prediction with AlphaFold.
+\newblock Nature. 2021;596:583--589.
+
+\bibitem{Lutge2021-mt}
+Lütge A, Zyprych-Walczak J, Brykczynska~Kunzmann U, Crowell HL, Calini D,
+  Malhotra D, et~al.
+\newblock CellMixS: quantifying and visualizing batch effects in single-cell
+  RNA-seq data.
+\newblock Life Sci Alliance. 2021;4.
+
+\bibitem{Reinke2024-dw}
+Reinke A, Tizabi MD, Baumgartner M, Eisenmann M, Heckmann-Nötzel D, Kavur AE,
+  et~al.
+\newblock Understanding metric-related pitfalls in image analysis validation.
+\newblock Nat Methods. 2024;21:182--194.
+
+\bibitem{funkyheatmap}
+Cannoodt R, Saelens W. type [;]Available from:
+  \url{https://cran.r-project.org/web/packages/funkyheatmap/index.html}.
+
+\bibitem{Saelens2019-jy}
+Saelens W, Cannoodt R, Todorov H, Saeys Y.
+\newblock A comparison of single-cell trajectory inference methods.
+\newblock Nat Biotechnol. 2019;37:547--554.
+
+\bibitem{Jelizarow2010-kr}
+Jelizarow M, Guillemot V, Tenenhaus A, Strimmer K, Boulesteix AL.
+\newblock Over-optimism in bioinformatics: an illustration.
+\newblock Bioinformatics. 2010;26:1990--1998.
+
+\bibitem{Boulesteix2010-zj}
+Boulesteix AL.
+\newblock Over-optimism in bioinformatics research.
+\newblock Bioinformatics. 2010;26:437--439.
+
+\bibitem{Strobl2024-lp}
+Strobl C, Leisch F.
+\newblock Against the "one method fits all data sets" philosophy for comparison
+  studies in methodological research.
+\newblock Biom J. 2024;66:e2200104.
+
+\bibitem{bettr}
+Marini F, Soneson C. type [; 2024]Available from:
+  \url{https://www.bioconductor.org/packages/release/bioc/html/bettr.html}.
+
+\bibitem{Taherdoost2023-wd}
+Taherdoost H, Madanchian M.
+\newblock Multi-criteria decision making ({MCDM}) methods and concepts.
+\newblock Encyclopedia (Basel, 2021). 2023;3:77--87.
+
+\bibitem{Niessl2022-fk}
+Nießl C, Herrmann M, Wiedemann C, Casalicchio G, Boulesteix AL.
+\newblock Over‐optimism in benchmark studies and the multiplicity of design
+  and analysis options when interpreting their results.
+\newblock Wiley Interdiscip Rev Data Min Knowl Discov. 2022;12:e1441.
+
+\bibitem{omnibenchmark}
+Mallona I, Luetge A, Carrillo B, Incicau D, Gerber R, Sonrel A, et~al.
+\newblock Omnibenchmark (alpha) for continuous and open benchmarking in
+  bioinformatics.
+\newblock arXiv preprint arXiv:240917038. 2024;.
+
+\bibitem{hatos2021apicuron}
+Hatos A, Quaglia F, Piovesan D, Tosatto SC.
+\newblock APICURON: a database to credit and acknowledge the work of
+  biocurators.
+\newblock Database. 2021;2021:baab019.
+
+\bibitem{druskat_2021_5171937}
+Druskat S, Spaaks JH, Chue~Hong N, Haines R, Baker J, Bliven S, et~al..
+  Citation File Format; 2021.
+\newblock Available from: \url{https://doi.org/10.5281/zenodo.5171937}.
+
+\bibitem{Greenstein2016-os}
+Greenstein S, Zhu F.
+\newblock Open content, Linus’ law, and neutral Point of View.
+\newblock Inf Syst Res. 2016;27:618--635.
+
+\bibitem{Sullivan2019-wh}
+Sullivan I, DeHaven A, Mellor D.
+\newblock Open and reproducible research on open science framework.
+\newblock Curr Protoc Essent Lab Tech. 2019;18:e32.
+
+\bibitem{Olevska2021-if}
+Olevska A, Bert B, Ebrahimi L, Schonfelder G, Heinl C.
+\newblock Ensuring reproducible research requires a support infrastructure: The
+  value of public registries to publishers.
+\newblock Sci Editor. 2021; p. 4--7.
+
+\bibitem{Teo2024-wk}
+Teo AYY, Squair JW, Courtine G, Skinnider MA.
+\newblock Best practices for differential accessibility analysis in single-cell
+  epigenomics.
+\newblock Nature Communications. 2024;15(1):8805.
+
+\bibitem{Siepe2023-tz}
+Siepe BS, Bartoš F, Morris TP, Boulesteix AL, Heck DW, Pawel S.
+\newblock Simulation studies for methodological research in psychology: A
+  standardized template for planning, preregistration, and reporting.
+\newblock PsyArXiv. 2023;.
+
+\bibitem{simpkin2023tertiary}
+Simpkin AJ, Mesdaghi S, S{\'a}nchez~Rodr{\'\i}guez F, Elliott L, Murphy DL,
+  Kryshtafovych A, et~al.
+\newblock Tertiary structure assessment at CASP15.
+\newblock Proteins: Structure, Function, and Bioinformatics.
+  2023;91(12):1616--1635.
+
+\bibitem{kryshtafovych2019critical}
+Kryshtafovych A, Schwede T, Topf M, Fidelis K, Moult J.
+\newblock Critical assessment of methods of protein structure prediction
+  (CASP)—Round XIII.
+\newblock Proteins: Structure, Function, and Bioinformatics.
+  2019;87(12):1011--1020.
+
+\bibitem{Howison2015-rc}
+Howison J, Deelman E, McLennan MJ, Ferreira~da Silva R, Herbsleb JD.
+\newblock Understanding the scientific software ecosystem and its impact:
+  Current and future measures.
+\newblock Res Eval. 2015;24:454--470.
+
+\bibitem{Trisovic2022-ol}
+Trisovic A, Lau MK, Pasquier T, Crosas M.
+\newblock A large-scale study on research code quality and execution.
+\newblock Sci Data. 2022;9:60.
+
+\bibitem{Pike1984-we}
+Pike R, Kernighan B.
+\newblock The {UNIX} system: Program design in the {UNIX} environment.
+\newblock AT T Bell Lab Tech J. 1984;63:1595--1605.
+
+\bibitem{Proctor2018-do}
+Proctor RW, Schneider DW.
+\newblock Hick’s law for choice reaction time: A review.
+\newblock Quarterly Journal of Experimental Psychology. 2018;71:1281--1299.
+
+\bibitem{ProcidaUnknown-vf}
+Procida D. Diátaxis;.
+\newblock Available from: \url{https://diataxis.fr/}.
+
+\bibitem{Hill2024-gf}
+Hill D, Antunes BA, Bertrand A, Nguifo EM, Yon L, Nautré-Domanski J, et~al..
+  Machine learning and reproducibility impact of random numbers; 2024.
+\newblock Available from: \url{https://hal.science/hal-04642175v1}.
+
+\bibitem{Kurtzer2017-mn}
+Kurtzer GM, Sochat V, Bauer MW.
+\newblock Singularity: Scientific containers for mobility of compute.
+\newblock PLoS One. 2017;12:e0177459.
+
+\bibitem{Hoste2012-gg}
+Hoste K, Timmerman J, Georges A, De~Weirdt S.
+\newblock EasyBuild: Building Software with Ease.
+\newblock In: 2012 SC Companion: High Performance Computing, Networking Storage
+  and Analysis. IEEE; 2012. p. 572--582.
+\newblock Available from:
+  \url{https://ieeexplore.ieee.org/abstract/document/6495863}.
+
+\bibitem{Gamblin2015-ll}
+Gamblin T, LeGendre M, Collette MR, Lee GL, Moody A, de~Supinski BR, et~al.
+\newblock The Spack package manager: bringing order to {HPC} software chaos.
+\newblock In: Proceedings of the International Conference for High Performance
+  Computing, Networking, Storage and Analysis. ACM; 2015.Available from:
+  \url{https://dl.acm.org/doi/10.1145/2807591.2807623}.
+
+\bibitem{Blomer2013-jk}
+Blomer J, Buncic P, Meusel R.
+\newblock The {CernVM} file system.
+\newblock CERN, Geneva, Switzerland, Tech Rep. 2013; p. 2--1.
+
+\bibitem{Boissonneault2019-wm}
+Boissonneault M, Oldeman BE, Taylor RP.
+\newblock Providing a unified software environment for Canada's national
+  advanced computing centers.
+\newblock In: Proceedings of the Practice and Experience in Advanced Research
+  Computing on Rise of the Machines (learning). ACM; 2019.Available from:
+  \url{https://dl.acm.org/doi/10.1145/3332186.3332210}.
+
+\bibitem{droge2023-ax}
+Dröge B, Holanda~Rusu V, Hoste K, van Leeuwen C, O'Cais A, Röblitz T.
+\newblock EESSI: A cross‐platform ready‐to‐use optimised scientific
+  software stack.
+\newblock Softw Pract Exp. 2023;53:176--210.
+
+\bibitem{Laine2007-py}
+Laine C, Goodman SN, Griswold ME, Sox HC.
+\newblock Reproducible Research: Moving toward Research the Public Can Really
+  Trust.
+\newblock Ann Intern Med. 2007;146:450.
+
+\bibitem{Kreutzer2014-ua}
+Kreutzer T.
+\newblock Open content: A practical guide to using Creative Commons licences.
+  2014;.
+
+\bibitem{Potter2015-fs}
+Potter ME, Smith T.
+\newblock Making code citable with Zenodo and GitHub.
+\newblock Software Sustainibility Institute. 2015;.
+
+\bibitem{Sicilia2017-ko}
+Sicilia MA, García-Barriocanal E, Sánchez-Alonso S.
+\newblock Community curation in open dataset repositories: Insights from
+  zenodo.
+\newblock Procedia Comput Sci. 2017;106:54--60.
+
+\bibitem{van-de-Sandt2019-wd}
+van~de Sandt S, Nielsen LH, Ioannidis A, Muench A, Henneken E, Accomazzi A,
+  et~al.
+\newblock Practice meets principle: Tracking software and data citations to
+  Zenodo DOIs.
+\newblock arXiv [csDL]. 2019;.
+
+\bibitem{Lamb2021-tb}
+Lamb C, Zacchiroli S.
+\newblock Reproducible Builds: Increasing the integrity of software supply
+  chains.
+\newblock arXiv [csSE]. 2021;.
+
+\bibitem{Huynh2023-pq}
+Huynh D. How Python Data Science Libraries Can Be Hijacked (and What You Can Do
+  About It); 2023.
+\newblock Available from:
+  \url{https://blog.mithrilsecurity.io/how-python-data-science-libraries-can-be-hijacked-and-what-you-can-do-about-it-2/}.
+
+\bibitem{Bleih2024-lv}
+Bleih A. New Vulnerability in {R}'s Deserialization Discovered; 2024-05-02.
+\newblock Available from:
+  \url{https://cyberint.com/blog/research/new-vulnerability-in-rs-deserialization-discovered/}.
+
+\end{thebibliography}
 

### 2025-07-23 2_second_review/manuscript.tex
Deduplicate the cernvmfs section (by Mark) in both highlighted and normal tex

@@
 Making the long-term replicability of a benchmark tractable implies adopting a sane software management system.
-Several solutions exist, from containerization (e.g., \texttt{apptainer} or \texttt{Singularity}) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{droge2023-ax}.
 
@@
 For these, software is the workflow manager's responsibility.
-On the other hand, some benchmarks might be designed to compare the same workflow topology run with different software backends (i.e., with conda vs compiled) or with different versions of a dependency (e.g., samtools/htslib releases).
+On the other hand, some benchmarks might be designed to compare the same workflow topology run with different software backends (i.e., with \texttt{conda} vs compiled) or with different versions of a dependency (e.g., \texttt{samtools}/\texttt{htslib} releases).
 Benchmarking systems should provide enough flexibility to make all variants possible (i.e., decouple the benchmark topology from the software backend).
@@
 Finally, scientific software management handling is known for its nuances, including abundant dependencies with known incompatibilities and complex installation recipes \cite{droge2023-ax}.
-Scientific software management is handled via traditional (Linux) package managers (RPMs, DEBs), central software stacks and environment modules (i.e., Lmod), conda, containers, or read-only network filesystems providing ready-to-use installations (CernVM-FS).
-These strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; \texttt{conda} is speeding up their deployment procedures (via \texttt{py-rattler}/\texttt{pixi}) and allowing compilation to Web assembly (to execute code within a web browser); and \texttt{apptainer} (former \texttt{Singularity}) provides rootless containerization suitable for running in scientific clusters.
+Scientific software management is handled via traditional (Linux) package managers (RPMs, DEBs), central software stacks and environment modules (i.e., Lmod), \texttt{conda}, containers (e.g., \texttt{apptainer} or \texttt{Singularity}; \cite{Kurtzer2017-mn}), or read-only network filesystems providing ready-to-use installations (CernVM-FS) \cite{Blomer2013-jk}.
+Further strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; \texttt{conda} is speeding up their deployment procedures (via \texttt{py-rattler}/\texttt{pixi}) and allowing compilation to Web assembly (to execute code within a web browser); and \texttt{apptainer} (former \texttt{Singularity}) provides rootless containerization suitable for running in scientific clusters.
 

### 2025-07-23 2_second_review/manuscript_with_tracked_changes.tex
Deduplicate the cernvmfs section (by Mark) in both highlighted and normal tex

@@
 Making the long-term replicability of a benchmark tractable implies adopting a sane software management system.
-Several solutions exist, from containerization (e.g., \texttt{apptainer} or \texttt{Singularity}) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{droge2023-ax}.
+{\color{red} \sout{ Several solutions exist, from containerization (e.g., \texttt{apptainer} or \texttt{Singularity}) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as the Digital Research Alliance of Canada (former \texttt{computecanada}) \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{droge2023-ax}.}}
 

### 2025-07-23 2_second_review/manuscript.tex
bring back Spack

@@
 Finally, scientific software management handling is known for its nuances, including abundant dependencies with known incompatibilities and complex installation recipes \cite{droge2023-ax}.
-Scientific software management is handled via traditional (Linux) package managers (RPMs, DEBs), central software stacks and environment modules (i.e., Lmod), \texttt{conda}, containers (e.g., \texttt{apptainer} or \texttt{Singularity}; \cite{Kurtzer2017-mn}), or read-only network filesystems providing ready-to-use installations (CernVM-FS) \cite{Blomer2013-jk}.
+Scientific software management is handled via traditional (Linux) package managers (RPMs, DEBs), automated and reproducible installs (such as \ttexttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}), central software stacks and environment modules (i.e., Lmod), \texttt{conda}, containers (e.g., \texttt{apptainer} or \texttt{Singularity}; \cite{Kurtzer2017-mn}), or read-only network filesystems providing ready-to-use installations (CernVM-FS) \cite{Blomer2013-jk}.
 Further strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; \texttt{conda} is speeding up their deployment procedures (via \texttt{py-rattler}/\texttt{pixi}) and allowing compilation to Web assembly (to execute code within a web browser); and \texttt{apptainer} (former \texttt{Singularity}) provides rootless containerization suitable for running in scientific clusters.

### 2025-07-23 2_second_review/manuscript.tex
Update bibitems for funkyheatmaps, bettr and diataxis; update texttts.

--- 2025-07-23 2_second_review/manuscript.tex
@@
 Finally, scientific software management handling is known for its nuances, including abundant dependencies with known incompatibilities and complex installation recipes \cite{droge2023-ax}.
-Scientific software management is handled via traditional (Linux) package managers (RPMs, DEBs), central software stacks and environment modules (i.e., Lmod), \texttt{conda}, containers (e.g., \texttt{apptainer} or \texttt{Singularity}; \cite{Kurtzer2017-mn}), or read-only network filesystems providing ready-to-use installations (CernVM-FS) \cite{Blomer2013-jk}.
-Further strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via CernVM-FS; \texttt{conda} is speeding up their deployment procedures (via \texttt{py-rattler}/\texttt{pixi}) and allowing compilation to Web assembly (to execute code within a web browser); and \texttt{apptainer} (former \texttt{Singularity}) provides rootless containerization suitable for running in scientific clusters.
+Scientific software management is handled via traditional (Linux) package managers (RPMs, DEBs), automated and reproducible installs (such as \texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}), central software stacks and environment modules (i.e., \texttt{Lmod}), \texttt{conda}, containers (e.g., \texttt{apptainer} or \texttt{Singularity}; \cite{Kurtzer2017-mn}), or read-only network filesystems providing ready-to-use installations (\texttt{CernVM-FS}) \cite{Blomer2013-jk}.
+Further strategies are under active development: EESSI \cite{droge2023-ax} and the Digital Research Alliance of Canada \cite{Boissonneault2019-wm} are effectively providing multi-purpose, reusable and efficient software installations via \texttt{CernVM-FS}; \texttt{conda} is speeding up their deployment procedures (via \texttt{py-rattler}/\texttt{pixi}) and allowing compilation to Web assembly (to execute code within a web browser); and \texttt{apptainer} (former \texttt{Singularity}) provides rootless containerization suitable for running in scientific clusters.
 
@@
 \bibitem{funkyheatmap}
-Cannoodt R, Saelens W. type [;]Available from:
-  \url{https://cran.r-project.org/web/packages/funkyheatmap/index.html}.
+Cannoodt R, Saelens W. 
+\newblock funkyheatmap: Generating Funky Heatmaps for Data Frames
+\newblock Available from \url{https://cran.r-project.org/web/packages/funkyheatmap/index.html}.
 
@@
 \bibitem{bettr}
-Marini F, Soneson C. type [; 2024]Available from:
-  \url{https://www.bioconductor.org/packages/release/bioc/html/bettr.html}.
+Marini F, Soneson C. 
+\newblock bettr: A Better Way To Explore What Is Best
+\newblock Available from: \url{https://www.bioconductor.org/packages/release/bioc/html/bettr.html}.
 
@@
 \bibitem{ProcidaUnknown-vf}
-Procida D. Diátaxis;.
+Procida D. 
+\newblock Diátaxis: A systematic approach to technical documentation authoring.
 \newblock Available from: \url{https://diataxis.fr/}.

### 2025-09-29 3_third_review/manuscript_with_tracked_changes.tex
Third round, centralization and FOSS.

--- 2025-07-23 2_second_review/manuscript.tex
@@
 
-Traditionally, benchmarks (whether BOP or MDP) are conducted in a decentralized fashion and typically by a small number of researchers (e.g., for MDP, the new method developer).
+Traditionally, benchmarks (whether BOP or MDP) are conducted in a {\color{red} solo and} decentralized fashion, {\color{red} \sout{and typically by}}, being started by a {\color{red} single or} small number of researchers (e.g., for MDP, the new method developer) {\color{red}using their local infrastructure}.
 This brings some disadvantages.
-For example, decentralization can be hugely inefficient; researchers studying the same task would collect reference datasets and code snippets to run methods, and implement a bespoke strategy (e.g., shell scripts) to orchestrate the methods run on datasets.
+For example, decentralization can be hugely inefficient; researchers studying the same task would collect reference datasets and code snippets to run methods, and implement a bespoke strategy (e.g., shell scripts) to orchestrate the methods run on datasets {\color{red}without being aware that their colleagues are simultaneously undertaking similar efforts, leading to redundant and fragmented work}.
 Working together would mean sharing not only datasets and code snippets, but also execution environments (software but perhaps also hardware), alignment on parameters to evaluate and eventually sharing, consolidating, and interpreting results.
@@
 
-Most benchmarks could begin decentralized, allowing for concept definition and preliminary results before `going public' (centralization phase).
+{\color{red}We believe } most benchmarks {\color{red}\sout{could}} begin decentralized, allowing for concept definition and preliminary results {\color{red} as conducted by solo researchers or small teams} before `going public', {\color{red}\sout{(} that is, welcoming contributions from other researchers while also agreeing on common practices, standards and a jointly accessible infrastructure for compute and data and code sharing; we call this} centralization phase {\color{red}\sout{)}}.
 Users need a system that supports building benchmarks both ways, transitioning to centralization as the design matures.
@@
 In terms of benchmark re-runnability, fully FOSS system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{droge2023-ax}.
+{\color{red} In contrast, proprietary software often restricts redistribution and re-execution due to licensing constraints, making it inherently non-rerunnable in many benchmarking contexts.
+Hence, benchmarkers should evaluate the trade-off between including all possible software, including proprietary, and its associated limitations for reproducibility and long-term accessibility.}
 
@@
 
-\bibitem{Kurtzer2017-mn}
-Kurtzer GM, Sochat V, Bauer MW.
-\newblock Singularity: Scientific containers for mobility of compute.
-\newblock PLoS One. 2017;12:e0177459.
+\bibitem{droge2023-ax}
+Dröge B, Holanda~Rusu V, Hoste K, van Leeuwen C, O'Cais A, Röblitz T.
+\newblock EESSI: A cross‐platform ready‐to‐use optimised scientific
+  software stack.
+\newblock Softw Pract Exp. 2023;53:176--210.
 
@@
 
+\bibitem{Kurtzer2017-mn}
+Kurtzer GM, Sochat V, Bauer MW.
+\newblock Singularity: Scientific containers for mobility of compute.
+\newblock PLoS One. 2017;12:e0177459.
+
 \bibitem{Blomer2013-jk}
@@
   \url{https://dl.acm.org/doi/10.1145/3332186.3332210}.
-
-\bibitem{droge2023-ax}
-Dröge B, Holanda~Rusu V, Hoste K, van Leeuwen C, O'Cais A, Röblitz T.
-\newblock EESSI: A cross‐platform ready‐to‐use optimised scientific
-  software stack.
-\newblock Softw Pract Exp. 2023;53:176--210.
 

### 2025-09-29 3_third_review/manuscript_with_tracked_changes.tex
rc, summoned the team

@@
 
-Traditionally, benchmarks (whether BOP or MDP) are conducted in a {\color{red} solo and} decentralized fashion, {\color{red} \sout{and typically by}}, being started by a {\color{red} single or} small number of researchers (e.g., for MDP, the new method developer) {\color{red}using their local infrastructure}.
+Traditionally, benchmarks (whether BOP or MDP) are conducted in a {\color{red} solo and} decentralized fashion, {\color{red} \sout{and typically by}}, being started by a {\color{red} single researcher or} small {\color{red}\sout{number of researchers}team} (e.g., for MDP, the new method developer) {\color{red}using their local infrastructure}.
 This brings some disadvantages.

### 2025-10-02 2_second_review/manuscript_with_tracked_changes.tex
Reorder eessi vs easybuild bibitems
Add Charlotte's and Mark's edits

@@
 Other ethical considerations include facilitating safe interactions and fair resource allocation and efficiency.
-Clear code of conduct statements proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) and provide explicit means to ease collaboration.
+Clear code of conduct statements and proactively flagging harassment (e.g., by sex, ethnicity or academic seniority, among others) provide explicit means to ease collaboration.
 In terms of efficiency, comparable allocation of resources (e.g., computing power, support) to all contributors, while also being aware of $CO_2$ emissions, could also be spelled out in the governance model.}

### 2025-10-02 3_third_review/manuscript_with_tracked_changes.tex
Add Charlotte's and Mark's edits

--- 2025-09-29 3_third_review/manuscript_with_tracked_changes.tex
@@
 
-Traditionally, benchmarks (whether BOP or MDP) are conducted in a {\color{red} solo and} decentralized fashion, {\color{red} \sout{and typically by}}, being started by a {\color{red} single or} small number of researchers (e.g., for MDP, the new method developer) {\color{red}using their local infrastructure}.
+Traditionally, benchmarks (whether BOP or MDP) are conducted in a {\color{red} solo and} decentralized fashion, {\color{red} \sout{and typically by}, }being started by a {\color{red} single researcher or} small {\color{red}\sout{number of researchers}team} (e.g., for MDP, the new method developer) {\color{red}using their local infrastructure}.
 This brings some disadvantages.
@@
 
-{\color{red}We believe } most benchmarks {\color{red}\sout{could}} begin decentralized, allowing for concept definition and preliminary results {\color{red} as conducted by solo researchers or small teams} before `going public', {\color{red}\sout{(} that is, welcoming contributions from other researchers while also agreeing on common practices, standards and a jointly accessible infrastructure for compute and data and code sharing; we call this} centralization phase {\color{red}\sout{)}}.
+{\color{red}We believe } most benchmarks {\color{red}\sout{could}} begin decentralized, allowing for concept definition and preliminary results {\color{red} as conducted by solo researchers or small teams} before `going public', {\color{red}\sout{(} i.e., welcoming contributions from other researchers while also agreeing on common practices, standards and a jointly accessible infrastructure for compute and data and code sharing; we call this the} centralization phase {\color{red}\sout{)}}.
 Users need a system that supports building benchmarks both ways, transitioning to centralization as the design matures.

### 2025-10-02 3_third_review/manuscript.tex
Draft colorless/untracked manuscript tex

--- 2025-09-29 3_third_review/manuscript_with_tracked_changes.tex
@@
 
-Traditionally, benchmarks (whether BOP or MDP) are conducted in a {\color{red} solo and} decentralized fashion, {\color{red} \sout{and typically by}}, being started by a {\color{red} single or} small number of researchers (e.g., for MDP, the new method developer) {\color{red}using their local infrastructure}.
+Traditionally, benchmarks (whether BOP or MDP) are conducted in a solo and decentralized fashion, being started by a single researcher or small team (e.g., for MDP, the new method developer) using their local infrastructure.
 This brings some disadvantages.
-For example, decentralization can be hugely inefficient; researchers studying the same task would collect reference datasets and code snippets to run methods, and implement a bespoke strategy (e.g., shell scripts) to orchestrate the methods run on datasets {\color{red}without being aware that their colleagues are simultaneously undertaking similar efforts, leading to redundant and fragmented work}.
+For example, decentralization can be hugely inefficient; researchers studying the same task would collect reference datasets and code snippets to run methods, and implement a bespoke strategy (e.g., shell scripts) to orchestrate the methods run on datasets without being aware that their colleagues are simultaneously undertaking similar efforts, leading to redundant and fragmented work.
 Working together would mean sharing not only datasets and code snippets, but also execution environments (software but perhaps also hardware), alignment on parameters to evaluate and eventually sharing, consolidating, and interpreting results.
@@
 
-{\color{red}We believe } most benchmarks {\color{red}\sout{could}} begin decentralized, allowing for concept definition and preliminary results {\color{red} as conducted by solo researchers or small teams} before `going public', {\color{red}\sout{(} that is, welcoming contributions from other researchers while also agreeing on common practices, standards and a jointly accessible infrastructure for compute and data and code sharing; we call this} centralization phase {\color{red}\sout{)}}.
+We believe most benchmarks begin decentralized, allowing for concept definition and preliminary results as conducted by solo researchers or small teams before `going public', i.e., welcoming contributions from other researchers while also agreeing on common practices, standards and a jointly accessible infrastructure for compute and data and code sharing; we call this the centralization phase.
 Users need a system that supports building benchmarks both ways, transitioning to centralization as the design matures.
@@
 In terms of benchmark re-runnability, fully FOSS system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{droge2023-ax}.
-{\color{red} In contrast, proprietary software often restricts redistribution and re-execution due to licensing constraints, making it inherently non-rerunnable in many benchmarking contexts.
-Hence, benchmarkers should evaluate the trade-off between including all possible software, including proprietary, and its associated limitations for reproducibility and long-term accessibility.}
+In contrast, proprietary software often restricts redistribution and re-execution due to licensing constraints, making it inherently non-rerunnable in many benchmarking contexts.
+Hence, benchmarkers should evaluate the trade-off between including all possible software, including proprietary, and its associated limitations for reproducibility and long-term accessibility.
 

### 2025-10-24 4_plos_template/plos_latex_template.tex
First templat-ization attempt

--- 2025-10-02 3_third_review/manuscript.tex
@@
-% Emerald Publishing - Construction Innovation Submission Template
-% by Oleksandr Melnyk
-% Ver 0.0.4
-% Based on: https://www.emeraldgrouppublishing.com/journal/ci#author-guidelines
+% Template for PLoS
+% Version 3.7 Aug 2025
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
 
-\documentclass[11pt]{article}
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
 
-\usepackage[english]{babel}
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
 
-% Set page size and margins
-% Replace `letterpaper' with `a4paper' for UK/EU standard size
-%\usepackage[a4paper,top=1.5cm,bottom=1.5cm,left=1.5cm,right=1.5cm,marginparwidth=1.75cm]{geometry}
-\usepackage[a4paper,top=2cm,bottom=2cm,left=2cm,right=2cm,marginparwidth=2.25cm]{geometry}
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
 
-% Useful packages
-\usepackage{amssymb}
-\usepackage{siunitx}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage[utf8]{inputenc}
-%\usepackage[right]{lineno}
-\usepackage{csquotes}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{adjustbox}
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage[nopatch=eqnum]{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
 \usepackage{array}
-\usepackage{url}
-\usepackage{titlesec}
-%\usepackage[compatibility=false]{caption}
-\usepackage{authblk}
-%\usepackage{xcolor} % Load the xcolor package for color options
-\usepackage{microtype}
-%\usepackage{url}
-\PassOptionsToPackage{hyphens}{url}
-\usepackage{hyperref}
 
-\usepackage[normalem]{ulem}
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
 
-% \usepackage[backend=biber]{biblatex}
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
 
-%\renewcommand{\thetable}{\Roman{table}}
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
 
-% Emerald Harvard Citation Style
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
 
-\usepackage[english]{babel}
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in
+\textheight 8.75in
 
-\title{Building a continuous benchmarking ecosystem in bioinformatics}
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
 
-\author[1,2,*]{Izaskun Mallona}
-\author[2,3]{Charlotte Soneson}
-\author[1,2]{Ben Carrillo}
-\author[1,2,4]{Almut Lütge}
-\author[1]{Daniel Incicau}
-\author[1,2]{Reto Gerber}
-\author[1,2]{Anthony Sonrel}
-\author[1,2,*]{Mark D. Robinson}
+% Please use the included `plos2025.bst` as your BibTeX style
+\bibliographystyle{plos2025}
 
-\affil[1]{Department of Molecular Life Sciences, University of Zurich, 8057 Zurich, Switzerland}
-\affil[2]{SIB Swiss Institute of Bioinformatics, University of Zurich, 8057 Zurich, Switzerland}
-\affil[3]{Friedrich Miescher Institute for Biomedical Research, 4056 Basel, Switzerland}
-\affil[4]{Swiss Data Science Centre, 8092 Zurich, Switzerland}
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
 
-\affil[*]{Correspondence: izaskun.mallona@gmail.com,mark.robinson@mls.uzh.ch}
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
 
-\linespread{1.4}
+%% Include all user-defined macros below
+
+\newcommand{\lorem}{\textbf{LOREM}}
+\newcommand{\ipsum}{\textbf{IPSUM}}
+
+%% END MACROS SECTION
 
 \begin{document}
-\maketitle
+\vspace*{0.2in}
 
-%\linenumbers
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf{Building a continuous benchmarking ecosystem in bioinformatics} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+%\author[1,2,*]{Izaskun Mallona}
+Izaskun Mallona\textsuperscript{1,2,*}, 
+%\author[2,3]{Charlotte Soneson}
+Charlotte Soneson\textsuperscript{2,3}
+%\author[1,2]{Ben Carrillo}
+Ben Carrillo\textsuperscript{1,2}
+%\author[1,2,4]{Almut Lütge}
+Almut Lütge\textsuperscript{1,2,4}
+%\author[1]{Daniel Incicau}
+Daniel Incicau\textsuperscript{1}
+%\author[1,2]{Reto Gerber}
+Reto Gerber\textsuperscript{1,2}
+%\author[1,2]{Anthony Sonrel}
+Anthony Sonrel\textsuperscript{1,2}
+%\author[1,2,*]{Mark D. Robinson}
+Mark D. Robinson\textsuperscript{1,2,*}
+\\
+\bigskip
+\textbf{1} {Department of Molecular Life Sciences, University of Zurich, 8057 Zurich, Switzerland}
+\\
+\textbf{2} {SIB Swiss Institute of Bioinformatics, University of Zurich, 8057 Zurich, Switzerland}
+\\
+\textbf{3} Friedrich Miescher Institute for Biomedical Research, 4056 Basel, Switzerland
+\\
+\textbf{4} Swiss Data Science Centre, 8092 Zurich, Switzerland
+\bigskip
 
-\begin{abstract}
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+%\Yinyang These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+%\ddag These authors also contributed equally to this work.
+
+% Current address notes
+%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency b Insert second current address 
+% \textcurrency c Insert third current address
+
+% Deceased author note
+%\dag Deceased
+
+% Group/Consortium Author Note
+%\textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* izaskun.mallona@gmail.com and mark.robinson@mls.uzh.ch
+
+\end{flushleft}
+% Please keep the abstract below 300 words
+
+% For PLOS Medicine research article authors, please structure your abstract
+% with "Background", "Method and Findings" and "Conclusion" sections per
+% journal requirements.
+
+% For PLOS Neglected Tropical Diseases research article authors, please
+% structure your abstract with "Background", "Methodology", "Findings", and
+% "Conclusion" sections per journal requirements.
+%
+\section*{Abstract}
 Benchmarking, which involves collecting reference datasets and demonstrating method performances, is a requirement for the development of new computational tools, but also becomes a domain of its own to achieve neutral comparisons of methods.
@@
 We discuss various ideas for organizing reproducible software environments, formally defining benchmarks, orchestrating standardized workflows, and how they interface with computing infrastructure.
-\end{abstract}
 
-%add 6 keywords
-\textbf{Keywords:} benchmarking; bioinformatics; reproducibility; software environments; licensing; workflows; infrastructure
+% Please keep the Author Summary between 150 and 200 words. Use first person.
+% PLOS ONE, PLOS Biology, PLOS Global Public Health, PLOS Mental Health, and PLOS Water authors please skip this step. Author Summary is not valid for submissions to these journals.
 
-\pagebreak
+% For PLOS Medicine authors, please structure your author summary with answers to the following questions:
+% Why was this study done?
+% What did the researchers do and find?
+% What do these findings mean?
+%
+%\section*{Author summary}
+%Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
+\linenumbers
+
+% Use "Eq" instead of "Equation" for equation citations.
 \section*{Introduction}
@@
 
+% Results and Discussion can be combined.
+
 \section*{Why a benchmarking system?}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/fig1_fig_benchmarking_aligned.pdf}
+    \includegraphics[width=\linewidth]{figures/fig1_fig_benchmarking_aligned.eps}
      \caption[Run types, properties, capabilities and features.]{Solo and community benchmarks, workflows, and benchmarking-system-enabled (\textit{bs}) benchmarks all differ in run properties and their consequences in terms of capabilities and features, including whether: inputs, truths, hidden truths, metrics or run results are available for readers; code is FOSS, software environments are managed, and execution automated; there is a clear track of (human) contributors; results are run in comparable hardware; an entrypoint to results or result interpretation are provided.
@@
 We define a benchmark as a conceptual framework to evaluate the performance of computational methods for a given task, though it can extend to assessing other aspects, like data-generating technologies.
-A benchmark thus requires a well-defined task and typically a definition of correctness, or ground-truth, to be precisely defined in advance  (Figure~\ref{fig:definitions}).
+A benchmark thus requires a well-defined task and typically a definition of correctness, or ground-truth, to be precisely defined in advance  (Fig~\ref{fig:definitions}).
 A benchmark then consists of benchmark `components', such as simulations to generate datasets, preprocessing steps, methods, and metrics.
@@
 Others, such as sharing results in an interactive manner and, particularly, result interpretation and making results reusable by others, go beyond what workflows offer.
-Similarly, collaborative benchmarking and challenges add extra layers of complexity, including collaboration tracking, hiding ground truths, or the ability to release multiple versions of the same benchmark as it develops (Figure~\ref{fig:definitions}).
+Similarly, collaborative benchmarking and challenges add extra layers of complexity, including collaboration tracking, hiding ground truths, or the ability to release multiple versions of the same benchmark as it develops (Fig~\ref{fig:definitions}).
 Moreover, scientific benchmarking borrows procedures from the scientific method and from scientific publishing, hence having expectations of fairness, reproducibility, transparency, and trust.
@@
 The benchmark definition could even be expressed as a single configuration file that specifies the scope and topology of components to include, details of the repositories (code implementations with versions), the instructions to create software environments (accessible across compute architectures), the parameters used, and what components to snapshot for a release.
-A benchmark definition in terms of layers and the corresponding challenges and opportunities within each layer is illustrated in Figure~\ref{fig:layers}.
+A benchmark definition in terms of layers and the corresponding challenges and opportunities within each layer is illustrated in Fig~\ref{fig:layers}.
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figures/fig2_bens_layers.pdf}
+    \includegraphics[width=0.8\linewidth]{figures/fig2_bens_layers.eps}
     \caption[Benchmarking conceptual layers and their challenges.]{Benchmarking conceptual layers and their challenges.
@@
 Some workflow languages can export workflow definitions to CWL, hence helping computational FAIR guidelines, i.e., bundling and processing data together with their metadata, tracking new metadata generation, recording provenance, and being accessible \cite{Goble2020-ps}.
-For benchmarking, we believe both the workflow layout and provenance can be formally defined (Figure~\ref{fig:formalization}).
+For benchmarking, we believe both the workflow layout and provenance can be formally defined (Fig~\ref{fig:formalization}).
 
@@
         %\includegraphics[width=0.75\linewidth]{../original_submission/figures/bens_formalization.pdf}
-    \includegraphics[width=0.75\linewidth]{./figures/fig3_bens_formalization_arrows.pdf} % alternative
+    \includegraphics[width=0.75\linewidth]{./figures/fig3_bens_formalization_arrows.eps} % alternative
     \caption[Benchmark formalization.]{Benchmark formalization.
@@
 
-Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack (the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments'); managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results (Figure~\ref{fig:definitions}).
+Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning appropriate hardware; handling a reproducible and efficient software stack (the advantages of decoupling environment handling from workflow execution are further discussed in the section~`Reproducible software environments'); managing storage and access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results (Fig~\ref{fig:definitions}).
 
@@
     \centering
-    \begin{adjustbox}{max width=\textwidth}
+%    \begin{adjustbox}{max width=\textwidth}
     \begin{tabular}{|l|l|l|l|l|l|l|}
@@
     \end{tabular}
-    \end{adjustbox}
+%    \end{adjustbox}
     \label{tab:licenses}
 \end{table}
-
-%https://docs.google.com/spreadsheets/d/1HvhiDHkY5BSvIJ4607krXIiymk9uGWKsEg3cwaZVfBI/edit?gid=0#gid=0
 
@@
 
-%\hfill
-
 \section*{Acknowledgments}
@@
 
-\section*{Funding}
-MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36\_14).
-CS is supported by the Novartis Research Foundation.
-The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+\nolinenumbers
 
-\section*{Conflict of interest}
+% Please compile your BiBTeX database using the "plos2025.bst" BibTeX style.
+% This file is part of the current package.
+% A sample BibTeX file is also included as "plos_bibtex_sample.bib".
+%
+% or
+%
+% Type in your references following Vancouver style and reference formatting instructions
+% available at https://journals.plos.org/plosone/s/submission-guidelines#loc-references
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
 
-We report no conflict of interest.
-
-\section*{Author contributions}
-
-Conceptualization: all. Figures: IM, BC. Original draft preparation: all. Writing, reviewing and editing: all. Supervision: IM, MDR. Funding acquisition: MDR.
-
-% the two sections below have been added to the online system now
-%\section*{Funding}
-%MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36\_14).
-CS is supported by the Novartis Research Foundation.
-
-%\section*{Competing interests}
-%No conflicts of interest are declared by the authors.
-
-\pagebreak
-\linespread{1} 
-%\printbibliography
-
-%\bibliographystyle{vancouver} % elsarticle-num} % vancouver}  
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2015}
-
-%\bibliography{src_do_not_upload/clean}
-
-\begin{thebibliography}{10}
-
-\bibitem{cao2023-jz}
-Cao Y, Yu L, Torkel M, Kim S, Lin Y, Yang P, et~al.
-\newblock The current landscape and emerging challenges of benchmarking
-  single-cell methods.
-\newblock bioRxiv. 2023; p. 2023.12.19.572303.
-
-\bibitem{Sonrel2023-te}
-Sonrel A, Luetge A, Soneson C, Mallona I, Germain PL, Knyazev S, et~al.
-\newblock Meta-analysis of (single-cell method) benchmarks reveals the need for
-  extensibility and interoperability.
-\newblock Genome Biol. 2023;24:119.
-
-\bibitem{Boulesteix2013-vy}
-Boulesteix AL, Lauer S, Eugster MJA.
-\newblock A plea for neutral comparison studies in computational sciences.
-\newblock PLoS One. 2013;8:e61562.
-
-\bibitem{Weber2019-el}
-Weber LM, Saelens W, Cannoodt R, Soneson C, Hapfelmeier A, Gardner PP, et~al.
-\newblock Essential guidelines for computational method benchmarking.
-\newblock Genome Biol. 2019;20:125.
-
-\bibitem{Brooks2024-rq}
-Brooks TG, Lahens NF, Mrčela A, Grant GR.
-\newblock Challenges and best practices in omics benchmarking.
-\newblock Nature Review Genetics. 2024;25(5):326--339.
-
-\bibitem{Rich2024-jh}
-Rich JM, Moses L, Einarsson PH, Jackson K, Luebbert L, Booeshaghi AS, et~al.
-\newblock The impact of package selection and versioning on single-cell
-  {RNA}-seq analysis.
-\newblock bioRxivorg. 2024; p. 2024.04.04.588111.
-
-\bibitem{Wilkinson2016-bh}
-Wilkinson MD, Dumontier M, Aalbersberg IJJ, Appleton G, Axton M, Baak A, et~al.
-\newblock The {FAIR} Guiding Principles for scientific data management and
-  stewardship.
-\newblock Sci Data. 2016;3:160018.
-
-\bibitem{Nevers_2022}
-Nevers Y, Jones TEM, Jyothi D, Yates B, Ferret M, Portell-Silva L, et~al.
-\newblock The Quest for Orthologs orthology benchmark service in 2022.
-\newblock Nucleic Acids Research. 2022;50(W1):W623–W632.
-\newblock doi:{10.1093/nar/gkac330}.
-
-\bibitem{Bryce_Smith_2023}
-Bryce-Smith S, Burri D, Gazzara MR, Herrmann CJ, Danecka W, Fitzsimmons CM,
-  et~al.
-\newblock Extensible benchmarking of methods that identify and quantify
-  polyadenylation sites from RNA-seq data.
-  2023;doi:{10.1101/2023.06.23.546284}.
-
-\bibitem{Pardo_Palacios_2023}
-Pardo-Palacios FJ, Wang D, Reese F, Diekhans M, Carbonell-Sala S, Williams B,
-  et~al.
-\newblock Systematic assessment of long-read RNA-seq methods for transcript
-  identification and quantification. 2023;doi:{10.1101/2023.07.25.550582}.
-
-\bibitem{meyer2021advances}
-Meyer P, Saez-Rodriguez J.
-\newblock Advances in systems biology modeling: 10 years of crowdsourcing DREAM
-  challenges.
-\newblock Cell Systems. 2021;12(6):636--653.
-
-\bibitem{zook2019open}
-Zook JM, McDaniel J, Olson ND, Wagner J, Parikh H, Heaton H, et~al.
-\newblock An open resource for accurately benchmarking small variant and
-  reference calls.
-\newblock Nature Biotechnology. 2019;37(5):561--566.
-
-\bibitem{moult2005-ne}
-Moult J.
-\newblock A decade of {CASP}: progress, bottlenecks and prognosis in protein
-  structure prediction.
-\newblock Curr Opin Struct Biol. 2005;15:285--289.
-
-\bibitem{Wratten2021-el}
-Wratten L, Wilm A, Göke J.
-\newblock Reproducible, scalable, and shareable analysis pipelines with
-  bioinformatics workflow managers.
-\newblock Nat Methods. 2021;18:1161--1168.
-
-\bibitem{Amstutz2024-qk}
-Amstutz P, Mikheev M, Crusoe MR, Tijanić N, Lampa S. Existing Workflow
-  systems; 2024.
-\newblock Available from: \url{https://s.apache.org/existing-workflow-systems}.
-
-\bibitem{Amstutz2016-vo}
-Amstutz P, Crusoe MR, Tijanić N, Chapman B, Chilton J, Heuer M, et~al.. Common
-  Workflow Language, {v1}.0; 2016.
-\newblock Available from:
-  \url{https://research.manchester.ac.uk/files/57032693/cwl_1.0_workflow.pdf}.
-
-\bibitem{Goble2020-ps}
-Goble C, Cohen-Boulakia S, Soiland-Reyes S, Garijo D, Gil Y, Crusoe MR, et~al.
-\newblock FAIR computational workflows.
-\newblock Data Intell. 2020;2:108--121.
-
-\bibitem{Hanssen2023-uf}
-Hanssen F, Gabernet G, Smith NH, Mertes C, Neogi AG, Brandhoff L, et~al.
-\newblock NCBench: providing an open, reproducible, transparent, adaptable, and
-  continuous benchmark approach for {DNA}-sequencing-based variant calling.
-\newblock F1000Research. 2023;12:1125.
-
-\bibitem{Capella-Gutierrez2017-dh}
-Capella-Gutierrez S, de~la Iglesia D, Haas J, Lourenco A, Fernández JM,
-  Repchevsky D, et~al.
-\newblock Lessons Learned: Recommendations for Establishing Critical Periodic
-  Scientific Benchmarking.
-\newblock bioRxiv. 2017; p. 181677.
-
-\bibitem{cannoodt2024viash}
-Cannoodt R, Cannoodt H, Schaumont D, Waldrant K, Van~de Kerckhove E, Boschmans
-  A, et~al.
-\newblock Viash: A meta-framework for building reusable workflow modules.
-\newblock Journal of Open Source Software. 2024;9(93):6089.
-
-\bibitem{luecken2025open}
-Luecken MD, Gigante S, Burkhardt DB, Cannoodt R, Strobl DC, Markov NS, et~al.
-\newblock Defining and benchmarking open problems in single-cell analysis.
-\newblock Nature Biotechnology. 2025;43(7):1035--1040.
-\newblock doi:{10.1038/s41587-025-02694-w}.
-
-\bibitem{LeboUnknown-om}
-Lebo T, Sahoo S, McGuinness D, Belhajjame K, Cheney J, Corsar D, et~al..
-  PROV-O: The PROV ontology;.
-\newblock Available from: \url{http://www.w3.org/TR/2013/REC-prov-o-20130430/}.
-
-\bibitem{bechhofer2013linked}
-Bechhofer S, Buchan I, De~Roure D, Missier P, Ainsworth J, Bhagat J, et~al.
-\newblock Why linked data is not enough for scientists.
-\newblock Future Generation Computer Systems. 2013;29(2):599--611.
-
-\bibitem{soiland2022packaging}
-Soiland-Reyes S, Sefton P, Crosas M, Castro LJ, Coppens F, Fern{\'a}ndez JM,
-  et~al.
-\newblock Packaging research artefacts with RO-Crate.
-\newblock Data Science. 2022;5(2):97--138.
-
-\bibitem{Khan2018-ot}
-Khan RU, Zhang X, Kumar R, Aboagye EO.
-\newblock Evaluating the performance of {ResNet} model based on image
-  recognition.
-\newblock In: Proceedings of the 2018 International Conference on Computing and
-  Artificial Intelligence. ACM; 2018.Available from:
-  \url{https://dl.acm.org/doi/10.1145/3194452.3194461}.
-
-\bibitem{Jumper2021-oq}
-Jumper J, Evans R, Pritzel A, Green T, Figurnov M, Ronneberger O, et~al.
-\newblock Highly accurate protein structure prediction with AlphaFold.
-\newblock Nature. 2021;596:583--589.
-
-\bibitem{Lutge2021-mt}
-Lütge A, Zyprych-Walczak J, Brykczynska~Kunzmann U, Crowell HL, Calini D,
-  Malhotra D, et~al.
-\newblock CellMixS: quantifying and visualizing batch effects in single-cell
-  RNA-seq data.
-\newblock Life Sci Alliance. 2021;4.
-
-\bibitem{Reinke2024-dw}
-Reinke A, Tizabi MD, Baumgartner M, Eisenmann M, Heckmann-Nötzel D, Kavur AE,
-  et~al.
-\newblock Understanding metric-related pitfalls in image analysis validation.
-\newblock Nat Methods. 2024;21:182--194.
-
-\bibitem{funkyheatmap}
-Cannoodt R, Saelens W. 
-\newblock funkyheatmap: Generating Funky Heatmaps for Data Frames
-\newblock Available from \url{https://cran.r-project.org/web/packages/funkyheatmap/index.html}.
-
-\bibitem{Saelens2019-jy}
-Saelens W, Cannoodt R, Todorov H, Saeys Y.
-\newblock A comparison of single-cell trajectory inference methods.
-\newblock Nat Biotechnol. 2019;37:547--554.
-
-\bibitem{Jelizarow2010-kr}
-Jelizarow M, Guillemot V, Tenenhaus A, Strimmer K, Boulesteix AL.
-\newblock Over-optimism in bioinformatics: an illustration.
-\newblock Bioinformatics. 2010;26:1990--1998.
-
-\bibitem{Boulesteix2010-zj}
-Boulesteix AL.
-\newblock Over-optimism in bioinformatics research.
-\newblock Bioinformatics. 2010;26:437--439.
-
-\bibitem{Strobl2024-lp}
-Strobl C, Leisch F.
-\newblock Against the "one method fits all data sets" philosophy for comparison
-  studies in methodological research.
-\newblock Biom J. 2024;66:e2200104.
-
-\bibitem{bettr}
-Marini F, Soneson C. 
-\newblock bettr: A Better Way To Explore What Is Best
-\newblock Available from: \url{https://www.bioconductor.org/packages/release/bioc/html/bettr.html}.
-
-\bibitem{Taherdoost2023-wd}
-Taherdoost H, Madanchian M.
-\newblock Multi-criteria decision making ({MCDM}) methods and concepts.
-\newblock Encyclopedia (Basel, 2021). 2023;3:77--87.
-
-\bibitem{Niessl2022-fk}
-Nießl C, Herrmann M, Wiedemann C, Casalicchio G, Boulesteix AL.
-\newblock Over‐optimism in benchmark studies and the multiplicity of design
-  and analysis options when interpreting their results.
-\newblock Wiley Interdiscip Rev Data Min Knowl Discov. 2022;12:e1441.
-
-\bibitem{omnibenchmark}
-Mallona I, Luetge A, Carrillo B, Incicau D, Gerber R, Sonrel A, et~al.
-\newblock Omnibenchmark (alpha) for continuous and open benchmarking in
-  bioinformatics.
-\newblock arXiv preprint arXiv:240917038. 2024;.
-
-\bibitem{hatos2021apicuron}
-Hatos A, Quaglia F, Piovesan D, Tosatto SC.
-\newblock APICURON: a database to credit and acknowledge the work of
-  biocurators.
-\newblock Database. 2021;2021:baab019.
-
-\bibitem{druskat_2021_5171937}
-Druskat S, Spaaks JH, Chue~Hong N, Haines R, Baker J, Bliven S, et~al..
-  Citation File Format; 2021.
-\newblock Available from: \url{https://doi.org/10.5281/zenodo.5171937}.
-
-\bibitem{Greenstein2016-os}
-Greenstein S, Zhu F.
-\newblock Open content, Linus’ law, and neutral Point of View.
-\newblock Inf Syst Res. 2016;27:618--635.
-
-\bibitem{Sullivan2019-wh}
-Sullivan I, DeHaven A, Mellor D.
-\newblock Open and reproducible research on open science framework.
-\newblock Curr Protoc Essent Lab Tech. 2019;18:e32.
-
-\bibitem{Olevska2021-if}
-Olevska A, Bert B, Ebrahimi L, Schonfelder G, Heinl C.
-\newblock Ensuring reproducible research requires a support infrastructure: The
-  value of public registries to publishers.
-\newblock Sci Editor. 2021; p. 4--7.
-
-\bibitem{Teo2024-wk}
-Teo AYY, Squair JW, Courtine G, Skinnider MA.
-\newblock Best practices for differential accessibility analysis in single-cell
-  epigenomics.
-\newblock Nature Communications. 2024;15(1):8805.
-
-\bibitem{Siepe2023-tz}
-Siepe BS, Bartoš F, Morris TP, Boulesteix AL, Heck DW, Pawel S.
-\newblock Simulation studies for methodological research in psychology: A
-  standardized template for planning, preregistration, and reporting.
-\newblock PsyArXiv. 2023;.
-
-\bibitem{simpkin2023tertiary}
-Simpkin AJ, Mesdaghi S, S{\'a}nchez~Rodr{\'\i}guez F, Elliott L, Murphy DL,
-  Kryshtafovych A, et~al.
-\newblock Tertiary structure assessment at CASP15.
-\newblock Proteins: Structure, Function, and Bioinformatics.
-  2023;91(12):1616--1635.
-
-\bibitem{kryshtafovych2019critical}
-Kryshtafovych A, Schwede T, Topf M, Fidelis K, Moult J.
-\newblock Critical assessment of methods of protein structure prediction
-  (CASP)—Round XIII.
-\newblock Proteins: Structure, Function, and Bioinformatics.
-  2019;87(12):1011--1020.
-
-\bibitem{Howison2015-rc}
-Howison J, Deelman E, McLennan MJ, Ferreira~da Silva R, Herbsleb JD.
-\newblock Understanding the scientific software ecosystem and its impact:
-  Current and future measures.
-\newblock Res Eval. 2015;24:454--470.
-
-\bibitem{Trisovic2022-ol}
-Trisovic A, Lau MK, Pasquier T, Crosas M.
-\newblock A large-scale study on research code quality and execution.
-\newblock Sci Data. 2022;9:60.
-
-\bibitem{Pike1984-we}
-Pike R, Kernighan B.
-\newblock The {UNIX} system: Program design in the {UNIX} environment.
-\newblock AT T Bell Lab Tech J. 1984;63:1595--1605.
-
-\bibitem{Proctor2018-do}
-Proctor RW, Schneider DW.
-\newblock Hick’s law for choice reaction time: A review.
-\newblock Quarterly Journal of Experimental Psychology. 2018;71:1281--1299.
-
-\bibitem{ProcidaUnknown-vf}
-Procida D. 
-\newblock Diátaxis: A systematic approach to technical documentation authoring.
-\newblock Available from: \url{https://diataxis.fr/}.
-
-\bibitem{Hill2024-gf}
-Hill D, Antunes BA, Bertrand A, Nguifo EM, Yon L, Nautré-Domanski J, et~al..
-  Machine learning and reproducibility impact of random numbers; 2024.
-\newblock Available from: \url{https://hal.science/hal-04642175v1}.
-
-\bibitem{droge2023-ax}
-Dröge B, Holanda~Rusu V, Hoste K, van Leeuwen C, O'Cais A, Röblitz T.
-\newblock EESSI: A cross‐platform ready‐to‐use optimised scientific
-  software stack.
-\newblock Softw Pract Exp. 2023;53:176--210.
-
-\bibitem{Hoste2012-gg}
-Hoste K, Timmerman J, Georges A, De~Weirdt S.
-\newblock EasyBuild: Building Software with Ease.
-\newblock In: 2012 SC Companion: High Performance Computing, Networking Storage
-  and Analysis. IEEE; 2012. p. 572--582.
-\newblock Available from:
-  \url{https://ieeexplore.ieee.org/abstract/document/6495863}.
-
-\bibitem{Gamblin2015-ll}
-Gamblin T, LeGendre M, Collette MR, Lee GL, Moody A, de~Supinski BR, et~al.
-\newblock The Spack package manager: bringing order to {HPC} software chaos.
-\newblock In: Proceedings of the International Conference for High Performance
-  Computing, Networking, Storage and Analysis. ACM; 2015.Available from:
-  \url{https://dl.acm.org/doi/10.1145/2807591.2807623}.
-
-\bibitem{Kurtzer2017-mn}
-Kurtzer GM, Sochat V, Bauer MW.
-\newblock Singularity: Scientific containers for mobility of compute.
-\newblock PLoS One. 2017;12:e0177459.
-
-\bibitem{Blomer2013-jk}
-Blomer J, Buncic P, Meusel R.
-\newblock The {CernVM} file system.
-\newblock CERN, Geneva, Switzerland, Tech Rep. 2013; p. 2--1.
-
-\bibitem{Boissonneault2019-wm}
-Boissonneault M, Oldeman BE, Taylor RP.
-\newblock Providing a unified software environment for Canada's national
-  advanced computing centers.
-\newblock In: Proceedings of the Practice and Experience in Advanced Research
-  Computing on Rise of the Machines (learning). ACM; 2019.Available from:
-  \url{https://dl.acm.org/doi/10.1145/3332186.3332210}.
-
-\bibitem{Laine2007-py}
-Laine C, Goodman SN, Griswold ME, Sox HC.
-\newblock Reproducible Research: Moving toward Research the Public Can Really
-  Trust.
-\newblock Ann Intern Med. 2007;146:450.
-
-\bibitem{Kreutzer2014-ua}
-Kreutzer T.
-\newblock Open content: A practical guide to using Creative Commons licences.
-  2014;.
-
-\bibitem{Potter2015-fs}
-Potter ME, Smith T.
-\newblock Making code citable with Zenodo and GitHub.
-\newblock Software Sustainibility Institute. 2015;.
-
-\bibitem{Sicilia2017-ko}
-Sicilia MA, García-Barriocanal E, Sánchez-Alonso S.
-\newblock Community curation in open dataset repositories: Insights from
-  zenodo.
-\newblock Procedia Comput Sci. 2017;106:54--60.
-
-\bibitem{van-de-Sandt2019-wd}
-van~de Sandt S, Nielsen LH, Ioannidis A, Muench A, Henneken E, Accomazzi A,
-  et~al.
-\newblock Practice meets principle: Tracking software and data citations to
-  Zenodo DOIs.
-\newblock arXiv [csDL]. 2019;.
-
-\bibitem{Lamb2021-tb}
-Lamb C, Zacchiroli S.
-\newblock Reproducible Builds: Increasing the integrity of software supply
-  chains.
-\newblock arXiv [csSE]. 2021;.
-
-\bibitem{Huynh2023-pq}
-Huynh D. How Python Data Science Libraries Can Be Hijacked (and What You Can Do
-  About It); 2023.
-\newblock Available from:
-  \url{https://blog.mithrilsecurity.io/how-python-data-science-libraries-can-be-hijacked-and-what-you-can-do-about-it-2/}.
-
-\bibitem{Bleih2024-lv}
-Bleih A. New Vulnerability in {R}'s Deserialization Discovered; 2024-05-02.
-\newblock Available from:
-  \url{https://cyberint.com/blog/research/new-vulnerability-in-rs-deserialization-discovered/}.
-
-\end{thebibliography}
-
+\bibliography{plos_bibtex_sample}
 \end{document}

### 2025-10-27 4_plos_template/plos_latex_template.tex
Update diataxis bibrecord; remove postcodes from affiliations

@@
 \usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% too adjust the width of a very wide table
+\usepackage{adjustbox}
 
@@
 \bigskip
-\textbf{1} {Department of Molecular Life Sciences, University of Zurich, 8057 Zurich, Switzerland}
+\textbf{1} {Department of Molecular Life Sciences, University of Zurich, Zurich, Switzerland}
 \\
-\textbf{2} {SIB Swiss Institute of Bioinformatics, University of Zurich, 8057 Zurich, Switzerland}
+\textbf{2} {SIB Swiss Institute of Bioinformatics, University of Zurich, Zurich, Switzerland}
 \\
-\textbf{3} Friedrich Miescher Institute for Biomedical Research, 4056 Basel, Switzerland
+\textbf{3} Friedrich Miescher Institute for Biomedical Research, Basel, Switzerland
 \\
-\textbf{4} Swiss Data Science Centre, 8092 Zurich, Switzerland
+\textbf{4} Swiss Data Science Centre, Zurich, Switzerland
 \bigskip
@@
     \centering
-%    \begin{adjustbox}{max width=\textwidth}
+    \begin{adjustbox}{max width=\textwidth}
     \begin{tabular}{|l|l|l|l|l|l|l|}
@@
     \end{tabular}
-%    \end{adjustbox}
+    \end{adjustbox}
     \label{tab:licenses}
